### 2013-07-31 exportfunctions/base4.tex
Added example graphs, Rd files, tex files, and codes

\documentclass[11pt]{article}
\usepackage{tikz}
\usetikzlibrary{decorations,arrows,shapes}
\begin{document}\begin{tikzpicture}[scale=1]

\node (H1) at (225bp,-125bp)[draw,circle,fill=white] {$\mu_1$};
\node (H2) at (425bp,-125bp)[draw,circle,fill=white] {$\mu_2$};
\node (H3) at (425bp,-325bp)[draw,circle,fill=white] {$\mu_3$};
\node (H4) at (225bp,-325bp)[draw,circle,fill=white] {$\mu_4$};
\draw [line width=1pt] (H1) to (325bp, -124bp) node[fill=blue!20] {$H_{1}$} to (H2);
\draw [line width=1pt] (H2) to (425bp, -224bp) node[fill=blue!20] {$H_{4}$} to (H3);
\draw [line width=1pt] (H4) to (225bp, -227bp) node[fill=blue!20] {$H_{3}$} to (H1);
\end{tikzpicture}
\end{document}

### 2013-07-31 exportfunctions/examplegraphs.R

@@ -1,0 +1,20 @@
+require(gMCP)
+
+source('~/newLife/code/gMCP/exportfunctions/functions.R')
+graphGUI()
+
+e <- environment()
+sapply(c('full4.Rd','base4.Rd','cycle4.Rd','claw4.Rd'),load,env=e)
+
+
+exportUDgraph(full4,file='full4.tex')
+exportUDgraph(cycle4,file='cycle4.tex')
+exportUDgraph(base4,file='base4.tex')
+exportUDgraph(claw4,file='claw4.tex')
+
+
+system('pdflatex full4.tex') 
+system('pdflatex cycle4.tex')
+system('pdflatex base4.tex')
+system('pdflatex claw4.tex') 
+

### 2013-07-31 exportfunctions/functions.R

@@ -1,0 +1,357 @@
+require(gMCP)
+require(MASS)
+
+exportUDgraph <- function(graph, file = NULL,package="TikZ", scale=1, alpha=0.05, pvalues,
+		fontsize=c("tiny","scriptsize", "footnotesize", "small",
+		"normalsize", "large", "Large", "LARGE", "huge", "Huge"),
+		nodeTikZ, labelTikZ="near start,above,fill=blue!20",
+		tikzEnv=TRUE, offset=c(0,0),fill=list(reject="red!80",retain="white"), nodeR=25) {
+	graph <- placeNodes(graph)
+	colors <- c("yellow","black","blue","red","green")
+	if (tikzEnv) {
+		tikz <- paste("\\begin{tikzpicture}[scale=",scale,"]\n", sep="")
+	} else {
+		tikz <- ""
+	}
+	nodes2 <- getUsableNames(getNodes(graph))
+	names(nodes2) <- getNodes(graph)
+	#tikz <- paste(tikz, "\\tikzset{help lines/.style=very thin}", paste="\n")	
+	for (node in getNodes(graph)) {
+		nodeColor <- ifelse(getRejected(graph, node),fill$reject, fill$retain)
+		x <- getXCoordinates(graph, node) + nodeR # Actually I'm not sure whether this "+ nodeR" in this function is really necessary.
+		y <- getYCoordinates(graph, node) + nodeR # It is more important to use nodeR to reduce the arc that should be drawn.
+
+                ### our graphs don't need alphas nor weights
+                
+		#alpha <- format(getWeights(graph,node), digits=3, drop0trailing=TRUE)
+		## weight <- paste(getLaTeXFraction(getWeights(graph,node)), collapse=" ")
+		## if (weight == 1) {
+		## 	weight <- "\\alpha"
+		## } else if (weight != "0") {
+		## 	weight <- paste(weight, "\\alpha", sep="")
+		## }
+
+                weight <- ""
+		double <- ""
+
+                ### we also don't use p-values
+                
+                ## if (!missing(pvalues)) {
+		## 	if (is.null(names(pvalues))) {
+		## 		names(pvalues) <- getNodes(graph)
[…]

### 2013-07-31 exportfunctions/examplegraphs.R
fixed load path

@@ -1,7 +1,5 @@
 require(gMCP)
-
-source('~/newLife/code/gMCP/exportfunctions/functions.R')
-graphGUI()
+source('functions.R')
 
 e <- environment()
 sapply(c('full4.Rd','base4.Rd','cycle4.Rd','claw4.Rd'),load,env=e)
@@ -12,7 +10,7 @@ exportUDgraph(cycle4,file='cycle4.tex')
 exportUDgraph(base4,file='base4.tex')
 exportUDgraph(claw4,file='claw4.tex')
 
-
+## tikz does not work on modesim?
 system('pdflatex full4.tex') 
 system('pdflatex cycle4.tex')
 system('pdflatex base4.tex')

### 2013-07-31 presentation/presentation.tex
added beamer presentation

% Created 2013-07-25 Thu 18:08
\documentclass[bigger]{beamer}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{fixltx2e}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{soul}
\usepackage{textcomp}
\usepackage{marvosym}
\usepackage{wasysym}
\usepackage{latexsym}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{bm}
\tolerance=1000
\usepackage{amsmath} \usepackage{algorithm} \usepackage{graphicx}
\usepackage{amssymb} \usepackage{mathrsfs} 
\newcommand{\bs}[1]{\bm{#1}}
\newcommand{\ie}{{\em i.e.},~}
\newcommand{\eg}{{\em e.g.},~}
\providecommand{\alert}[1]{\textbf{#1}}

\title{Matroids from families of null hypotheses}
\author{Willi Maurer, Florian Klinglmueller}
\date{\today}
\hypersetup{
  pdfkeywords={},
  pdfsubject={},
  pdfcreator={Emacs Org-mode version 7.8.11}}

\begin{document}

\maketitle

\begin{frame}
\frametitle{Outline}
\setcounter{tocdepth}{3}
\tableofcontents
\end{frame}

\section{Logically dependent hypotheses}

\begin{frame}
\frametitle{Pairwise mean comparisons (two-sided)}
\begin{itemize}

\item Consider $\mu_i, i \in \{1,...,m\}$

\item $H_k: \mu_i = \mu_j, k \in S = \{1,...,n\}$, where $n \leq m*(m-1)/2$

\item Can be represented by an undirected graph $\mathscr{G} = (V,E)$:
\begin{itemize}
\item vertices correspond to means, \emph{i.e.}, $V = \{\mu_1,...,\mu_m\}$,
\item edges correspond to hypotheses, \emph{i.e.}, $E =
      \left\{e_1,...,e_n\} \right\}$, $e_l = \{\mu_i,\mu_j\}$, $i \neq
      j$, and $e_l \neq e_k$ for $l \neq k$
\end{itemize}
\item Let $H_I = \bigcap_{k \in I} H_k$, $I \subseteq S$
\item $I$ induces a subgraph $\mathscr{G}_I$ on the edge set $E_I=\{e_i: i
  \in I\}$
\end{itemize} % ends low level
\end{frame}

\begin{frame}
\frametitle{Example 1: 4 Means}
\begin{itemize}

\item \textbf{/Some illustration hereabouts/}

\end{itemize} % ends low level
\end{frame}

\begin{frame}
\frametitle{General contrast tests (two-sided)}
\begin{itemize}

\item Consider $m$-variate random variable $\bs{X} \sim F(\bs{\mu})$, $\bs{\mu} = (\mu_1,...,\mu_m)^t$

\item Contrasts $\bs{c}_i = (c_{i1},...,c_{im})^t$, $\sum_{j = 1}^m
  c_{ij} = 0$
\item Contrast matrix  $\bs{C} = \left( \bs{c}_1,...,\bs{c}_n\right)$,
  with sub-matrix $\bs{C}_I = \left( \bs{c}_i \right)_{i \in
    I}$
\item Elementary hypotheses $H_i: \bs{c}_i^t\bs{\mu} = 0$,
\item Intersection hypotheses $H_I: \bs{C}_I^t\bs{\mu} = 0$

\item Elementary hypotheses are represented by column vectors,
  intersection hypotheses by sub-matrices of the contrast matrix

\end{itemize} % ends low level
\end{frame}

\begin{frame}
\frametitle{Example 2: Contrast test}
\begin{itemize}

\item \textbf{/Some illustration hereabouts/}

\end{itemize} % ends low level
\end{frame}

\begin{frame}
\frametitle{Dependent null hypotheses}

\begin{itemize}
\item $H_i$ depends on $H_I$ iff $H_{I \cup i} = H_{I}$, induces a
  dependence relation $\sim_H$ on $\mathcal{P}(S)$
\item $H_I$ independent iff for all $i \in I$, $i \nsim_H I$,
\item $I$ independent $\Leftrightarrow$ $H_{I} \subsetneq  H_{I\setminus i}$
\item For pairwise comparisons $I$ independent iff there are no cycles in $\mathscr{G}_I$
\item For general contrast tests $I$ independent iff $\bs{C_I}$ has full
  column rank
\end{itemize} % ends low level
\end{frame}

\section{Matroids}

\begin{frame}
\frametitle{Matroids}
\begin{itemize}

\item A matroid on $S$ endows $\mathcal{P}(S)$ with an algebraic
  structure 
\item Matroids generalize the structures of graphs and vector spaces,
  \eg circuits, bases, independent sets
\item Many ways to define a matroid
  \begin{itemize}
  \item From a graph perspective: Circuit axioms
  \item From a vector space perspective: Independence axioms
  \end{itemize}
\item There is a dual matroid $M^*$ which can be helpful for proofs and algorithms 
\item \textbf{/Some illustration hereabouts/}
\end{itemize} % ends low level
\end{frame}

\begin{frame}
\frametitle{Circuit axioms}

Consider $S$ and a collection of subsets $\mathscr{C}$ for which holds
\begin{enumerate}
\item $C_1 \neq C_2 \in \mathscr{C}$, then $C_1 \nsubseteq C_2$,
\item $C_1 \neq C_2 \in \mathscr{C}$, and $i \in C_1 \cap C_2$ then
   $\exists C_3 \subseteq (C_1 \cup C_2)$,
\end{enumerate}
then $\mathscr{C}$ is the set of circuits of a matroid $M$ on $S$
\begin{itemize}

\item \textbf{/Some illustration hereabouts/}

\end{itemize} % ends low level
\end{frame}

\begin{frame}
\frametitle{Independence axioms}

\begin{block}
  Consider $S$ and a collection of subsets $\mathscr{I}$ for which
  holds
  \begin{enumerate}{Independence}
  \item $\emptyset \in \mathscr{I}$,
  \item $I \in \mathscr{I}$ and $J \subseteq I$ then $J \in
    \mathscr{I}$,
  \item $I,J \in \mathscr{I}$, $|I| = |J + 1|$ then $\exists j \in I
    \setminus J$ such that $J \cup i \in \mathscr{I}$,
  \end{enumerate}
  then $\mathscr{I}$ is the set of independent sets of a matroid $M$
  on $S$. This defines an independence relation $i \nsim_M I$ iff
  $I\cap i \in \mathscr{I}$. 
\end{block}
\begin{block}{Circuit}
  $C \subset S$ is a circuit iff $C \neq \mathscr{I}$ and for all $I
  \subset C$, $I \in \mathscr{I}$ (\ie minimally dependent set)
\end{block}

\end{frame}

\begin{frame}
\frametitle{Dependence axioms}

\begin{block}{Dependence}
  Consider $S$ and a relation $\sim_M$ on elements and subsets of $S$,
  then if for any $x,y_1,...,y_n,z_1,...,z_m$ holds
  \begin{enumerate}
  \item $y_i \sim \{y_1,...,y_n\}$ for all $i \leq n$
  \item if $n\geq 1$, $x \sim_M \{y_1,...,y_n\}$ and $x \nsim_M
    \{y_1,...,y_{n-1}$ then $y_n \sim \{y_1,...,y_{n-1},x\}$
  \item if $x \sim_M \{y_1,...,y_n\}$ and for all $i \leq n$ $y_i
    \sim_M \{z_1,...,z_m\}$ then $x \sim_M \{z_1,...,z_m\}$
  \end{enumerate}
  then $\sim_M$ defines a dependence relation of a matroid on $S$.
\end{block}

\end{frame}

\begin{frame}{Maximally independent sets}
  \begin{itemize}
  \item Bases are maximally independent sets of $S$
  \item Rank $\rho I$ is the cardinality of a maximally indepedendent set in $I$
  \item Circuits are minimally dependent sets, \ie each subset of $C$
    with rank $\rho C - 1$ is a maximally independent set of $C$
  \end{itemize}
\end{frame}

\begin{frame}
\frametitle{Maximally dependent sets}
\begin{itemize}
\item The closure of a set $\sigma I = \{i \in S: i \sim I\}$
\item Closed sets {\em flats}: $\mathscr{L} = \{I \subseteq S: \sigma I = I\}$
\item Hyperplanes are flats of rank $\rho S - 1$
\item All flats are given by closing hyperplanes under intersection
\item Circuits in $M^*$ are the complements of hyperplanes in $M$
\end{itemize} % ends low level
\end{frame}
\section{Matroids from logically dependent hypotheses}

\begin{frame}
\frametitle{Matroids from logically dependent hypotheses}

  \begin{block}{Theorem}
    For a family of two-sided pairwise mean comparisons $H_i$, $i \in
    S$, an intersection hypotheses $H_I$ is independent iff the
    corresponding edge set $\{e_i:i \in I\}$ is independent in the
    cycle matroid on $\mathscr{G}_I$.

For a family of two-sided
    general contrast tests $H_i$, $i \in S$, an intersection
    hypotheses $H_I$ is independent iff the corresponding contrast
    matrix $\bs{C}_I$ is independent in the vectorial matroid on
    $\mathbb{R}^m$.

    \begin{displaymath}
      i \sim_H I \Leftrightarrow i \sim_M I
    \end{displaymath}
  \end{block}

\end{frame}

\begin{frame}
\frametitle{Closed test procedure}
\begin{itemize}

\item The flats of a matroid are closed under intersection

\item Every elementary hypothesis is a flat

\item Using local level-$\alpha$ tests for each flat we have a closed test procedure that controls the FWER in the strong sense (Closed test principle)

\end{itemize} % ends low level
\end{frame}

\begin{frame}
\frametitle{Examples}
\begin{itemize}

\item Pairwise comparisons

\item The flats of the complete graph correspond to the partitions of $\{1,...,m\}$

\item The flats of a matroid generate a lattice, procedures testing not all possible comparisons correspond to intervals on that lattice

\end{itemize} % ends low level
\end{frame}

\begin{frame}
\frametitle{Connections to existing procedures}

\textbf{Old wine in new bottles ((c) L. Hothorn)}
\begin{itemize}

\item Flats <-> Weakly Exhaustive set (Hommel, Bergmann)

\item Lemma 1.5 in (Bernhard) about weakly exhaustive sets are ``closure axioms'' of a matroid

\item Graphical procedures for pairwise comparisons (Weichert) identify dependence using cycle in the graph

\item \emph{However:}
\begin{itemize}
\item proofs are shorter (more elegant),
\item efficient algorithms exist to find the flats of matroids,
\item we may define procedures using independent sets,
\item \ldots{}
\end{itemize}

\end{itemize} % ends low level
\end{frame}
\section{Graphical approaches for logically dependent hypotheses}

\begin{frame}
\frametitle{Reminder: Graphical approaches}
\end{frame}

\begin{frame}
\frametitle{Hierarchical tests}
\end{frame}

\begin{frame}
\frametitle{General weighting strategies}
\end{frame}
\section{Matriods from general families of hypotheses}

\begin{frame}
\frametitle{Matroid conditions}
\end{frame}

\begin{frame}
\frametitle{Flats <-> Weakliy exhaustive sets}
\end{frame}

\begin{frame}
\frametitle{Flats <-> Strictly exhaustive sets}
\end{frame}
\section{Summary \& Outlook}

\begin{frame}
\frametitle{Summary}
\end{frame}

\begin{frame}
\frametitle{Outlook}

\end{frame}

\end{document}

### 2013-08-08 presentation/presentation.tex
first graph included, git on windows is a pain in the knee
included first example graphs, changed scaling of graphs to .5
Added a couple of functions (working algorithm) to the package, added the presentation, should probably consider creating a branch for the presentation

@@
 \usepackage[utf8]{inputenc}
+\usepackage{tikz}
+\usetikzlibrary{decorations,arrows,shapes}
 \usepackage[T1]{fontenc}
@@
 \frametitle{Example 1: 4 Means}
-\begin{itemize}
-
-\item \textbf{/Some illustration hereabouts/}
-
-\end{itemize} % ends low level
+\input{full4}
 \end{frame}
@@
 then $\mathscr{C}$ is the set of circuits of a matroid $M$ on $S$
-\begin{itemize}
 
-\item \textbf{/Some illustration hereabouts/}
+\input{cycle}
 
-\end{itemize} % ends low level
 \end{frame}
@@
 \frametitle{Independence axioms}
-
-\begin{block}
+\begin{block}{Independence}
   Consider $S$ and a collection of subsets $\mathscr{I}$ for which
   holds
-  \begin{enumerate}{Independence}
+  \begin{enumerate}
   \item $\emptyset \in \mathscr{I}$,
@@
 \end{block}
+
 \begin{block}{Circuit}
@@
 \end{frame}
+
+\begin{frame}{Examples}
+\framesubtitle<1>{Independent set}  
+\framesubtitle<2>{Base}  
+\framesubtitle<4>{Dependent set}  
+\framesubtitle<3>{Cycle}  
+\framesubtitle<5>{Closure}  
+\only<1>{\input{independent}}
+\only<2>{\input{base}}
+\only<4>{\input{dependent}}
+\only<3>{\input{cycle}}
+\only<5>{\input{closure}}
+\end{frame}
+
 \section{Matroids from logically dependent hypotheses}
@@
 \end{frame}
+
+\begin{frame}{Matroid procedures for pairwise mean comparisons}
+
+  \begin{itemize}
+  \item Two-sided comparisons between $m$ means
+  \item The cycle matroid of the corresponding graph representation
+    provides an equivalent dependence structure
+  \item The flats of a (simple) complete graph are
+    isomporphic to the partitions of a set of size $m$
+  \item Provides an efficient way to generate all exhaustive sets by
+    enumeration of all partitions [ER '88] or hyperplanes 
+  \item The number of exhaustive sets are given by the Bell numbers
+
+  \end{itemize}
+
+\end{frame}
+
 \section{Graphical approaches for logically dependent hypotheses}
@@
 \frametitle{General weighting strategies}
+
+\begin{enumerate}
+\item Pretest using weighted Bonferroni or conventional graphical
+  approach $R = \{i \in S: H_i \textrm{ rejected by pretest}\}$
+\item Compute all matroid hyperplanes $\mathscr{H}$ for the matroid on
+  the remaining hypotheses $I$
+\item $R \gets R \cup \{i: p_i \leq \min_J w_{i,J} \alpha, J \in \mathscr{H}\}$
+\item $A \gets \{\bigcup J: \min_j (p_j / w_{j,J}) > \alpha\}$ 
+\item $I \gets I \setminus R$
+\item Refine: $\mathscr{H} \gets \{J\in \mathscr{L}: \rho J = \rho
+  \mathscr{H} - 1\}$ 
+\item Prune: $\mathscr{H} \gets \{J \in \mathscr{H}: J \cap R = \emptyset\}$
+\item Repeat until $\rho \mathscr {H} = 0$
+\item Reject $R$ and $I \setminus A$
+\end{enumerate}
 \end{frame}

### 2013-08-08 package/R/functions.R

@@ -1,0 +1,184 @@
+ij2l <- function(ij,m){
+  ## convert mean indices to a unique number
+  ## assume i<j
+  sum(m:(m-ij[1]+1))+(ij[2]-ij[1])-m
+}
+
+l2ij <- function(l,m){
+  ## convert number to unique mean
+  ## indeces such that i<j and i,j<=m
+  i <- sum(cumsum((m-1):1) <= (l-1))
+  j <- l%%sum((m-1):(m-i))+i
+  c(i+1,j+1)
+}
+
+I2L <- function(I,m){
+  ## convert an index set to its combinadic
+  k <- length(I)
+  if(k == 1){
+    return(I)
+  }
+  ## we do not assume that I is sorted
+  I <- sort(I)
+  sum(choose(m,1:k-1)) + sum(choose(I-1,1:k)) 
+}
+
+L2I <- function(L,m){
+  ## convert combinadic to index set
+  cs <- cumsum(choose(m,1:(m-1)))
+  ## no of hypotheses
+  k <- sum(cs < L)+1
+  if(k == 1){
+    return(L)
+  }
+  L <- L - cs[k-1] - 1
+  if(L == 0){
+    return(1:k)
+  }
+  I <- integer(k)
+  ## greedy algorithm to find combination
+  for(i in k:1){
+    cs <- choose(1:m,i)
+    h <- max(c(i-1,which(cs <= L)))
+    L <- L-cs[h]
+    I[i] <- h+1
+  }
+  return(I)
+}
+
+means <- function(el){
+  ## get the vertices of an edge list
+  sort(unique(as.numeric(el)))
+}
+
+el2partition <- function(el){
+  ## get the partition representation of an edge list
+  ## uses igraph
+  part <- list()
+  elc <- matrix(as.character(el),nrow(el),ncol(el))
+  coco <- clusters(graph.edgelist(elc))
+  for(i in 1:coco$no){
+  ## beach boys yeah
+    part[[i]] <- means(el)[coco$membership == i]
+  }
+  return(part)
+}
+
+partition2closedel <- function(part){
+  ## get closure of a partition
+  el <- do.call('cbind',lapply(part,function(p){
+    if(length(p)<2) return(NULL)
+    combn(p,2)
+  }))
+  if(is.matrix(el)){
+    return(t(el))
+  } else {
+    return(NULL)
+  }
+}
+
+intersectel <- function(al,bl){
[…]

### 2013-08-08 paper/draft.tex
added paper draft to repo

\documentclass[a4paper,12pt]{article}
\usepackage{algorithmic}
\usepackage{algorithm}
\usepackage{graphicx}
%\usepackage{titlesec}
\usepackage{multirow}
\setlength\parindent{0in}
\usepackage[hdivide={2cm,*,2cm},vdivide={1in,*,1in}]{geometry}
\usepackage{setspace}
\doublespacing
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathrsfs}
\newcommand{\ud}{\, {\rm d} \kern-.015em }
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}
\newtheorem{comment}{Comment}
\newtheorem{lemma}{Lemma}
\newtheorem{definition}{Definition}
\newtheorem{proposition}{Proposition}
\newtheorem{bsp}{Example}
\usepackage{natbib}
\usepackage{color}
\newenvironment{proof}[1][Proof]{\noindent\textbf{#1:} }{}
\newcommand{\bs}[1]{\boldsymbol{#1}}
\usepackage{authblk}
\usepackage{lineno}
%\titlelabel{\thetitle.\quad}
\newcommand{\gmcp}{\mathcal{G}}

\renewcommand{\thesection}{\arabic{section}.}
\renewcommand{\thesubsection}{\arabic{section}.\arabic{subsection}}
%% Comments
%\newcommand{\fcomment}[1]{ } % Turn off
\newcommand{\fcomment}[1]{\footnote{{\color{red} #1}}} % Turn on
\newcommand{\ie}{{\em i.e.,}~}
\newcommand{\cf}[1]{\mathbf{1}_{\left\{#1\right\}}}

\title{Multiple testing with logical dependencies}

\author{Florian Klinglmueller, Willi Maurer}

%\affil[1]{Medical and Pharmaceutical Statistics Research Unit,
%Lancaster University, Lancaster, United Kingdom}
%\affil[2]{European Medicines Agency, London, United Kingdom}
%\affil[3]{Center for Medical Statistics, Informatics and Intelligent Systems, Medical University of Vienna, Vienna, Austria%}
%\affil[*]{{\small To whom correspondence should be
%addressed: E-mail: florian.klinglmueller@meduniwien.ac.at}}

\date{}

\begin{document}

\maketitle

\section{Introduction}

Closed testing procedures provide powerful multiple tests for a wide
array of general multiple comparison problems. Recently, methods based
on closed tests of (weighted) Bonferroni tests have gotten much
attention in the literature. They do not require any assumptions on or
knowledge about the correlation structure between test statistics. In
combination with weighting strategies such procedures may be tailored
to reflect contextual relations and differences between the elementary
hypotheses in practical applications. Graphical methods are available
that offer a visually accessible tool to design and communicate such
procedures.

\section{Examples}

\subsection{Pairwise comparisons of population means}
\begin{itemize}
\item All pairwise
\item Gold standard designs
\item Change point analysis
\item General treatment control - treatment treatment situations
\end{itemize}

\subsection{General linear contrast tests}
\begin{itemize}
\item Cost effectiveness (Georg)
\item SNP Studies (Hothorn-Hothorn)
\item Order restriction contrasts (Williams, ...)
\end{itemize}

\subsection{Further examples}

\begin{description}
\item[Test for the degree of a polynomial]
\item[Model selection]
\item[Compositional data]
\item[Monotonicity restrictions not arising from linear dependencies]
\end{description}

\section{Two sided hypotheses}

\subsection{Some observations on pairwise comparisons}

\begin{definition}
  Let $X$ be a random variable taking values in $\mathbb{R}^m$ with
  multivariate distribution $F(\bs{\mu})$ having mean vector $\bs{\mu}
  \in \mathbb{R}^m$. Let $R \subset \{1,...,m\}^2$, such that $i <
  j$ for all $(i,j) \in R$. For $(i,j) \in R$ $H_{ij} = \mu_i =
  \mu_j$. We call the family of elementary hypotheses $\dot{R}
  = \{H_{ij}: (i,j) \in R\}$ a {\bf two-sided pairwise comparison
    problem. (tPCP)} Let $\dot{\mathcal{T}}(m)$ denote the set of all
  possible pairwise comparison problems of $m$ means.
\end{definition}

\begin{bsp}
  Consider $X \sim N((\mu_1,\mu_2,\mu_3,\mu_4),\sigma \mathbb{I}_4)$
  let $R = \{(1,2),(1,3),(1,4)\}$ and $S =
  \{(1,4),(2,4),(3,4),(1,3),(1,2),(2,3)\}$. Then the
  corresponding tPCP's $\dot{R},\dot{S} \in \dot{\mathcal{T}}(4)$.
\end{bsp}

\begin{definition}
  A two-sided pairwise comparison problem $\dot{S}$ defines an
  undirected graph $G(\dot{S}) = \{M,T\}$ with Nodes $M = \{1,...,m\}$
  and Edges $T$. We call $G(\dot{S})$ the {\bf comparigram}. We denote
  by $V(G(\dot{S})$, $E(G(\dot{S})$ the vertex and edge sets of the
  corresponding comparigram. If clear from the context we may just use
  $E(\dot{S}),V(\dot{S})$. 
\end{definition}

\begin{bsp}
  $\dot{R}$ in above Example is represented by a claw with 3 edges,
  $\dot{S}$ is represented by the complete graph $K_4$ with 4 vertices.
\end{bsp}

\begin{definition}
  Let $h_{lk}:  M^2 \rightarrow \dot{\mathcal{T}}(m)$,
  $(l,k) \mapsto \{H_{ij}: (i,j) \in T\setminus \{(l,j)\}\}$ define the
  {\bf rejection operator}. We write $h_{lk} \circ \dot{S} =
  \dot{S}\setminus (l,k)$. Clearly   $h_{a} \circ h_{b} \circ \dot{S}
  = h_{b} \circ h_{a} \circ \dot{S}$.
\end{definition}

\begin{definition}
  Consider $\dot{R},\dot{S} \in \dot{\mathcal{T}}(m)$ with associated
  comparigrams $L = G(\dot{R}), K = G(\dot{R})$ respectively. We say
  that $\dot{R}$ {\bf isomorphic} to $\dot{S}$, $\dot{R} \cong
  \dot{S}$ if $L$ is isomorphic to $K$, that is there exists a
  bijection between nodes of the graphs $f:V(K) \rightarrow V(L)$, $a
  \mapsto u$ such that for $a,b \in V(K)$, $f(a)$ is adjacent to
  $f(b)$, if and only if $a$ and $b$ are adjacent to each other.
\end{definition}

\begin{definition}
  Consider $\dot{S} \in \dot{\mathcal{T}}(m)$ and $h_a,h_b$, $a,b \in
  E(S)$ we say that $h_a \cong h_b$ are {\bf equivalent rejections} if $h_a
  \circ \dot{S} \cong h_b \circ \dot{S}$.
\end{definition}

\begin{definition}
  Let $h = (h_1,...,h_n)$ whith $h_i \in T$ and $n \leq |T|$ define a
  {\bf fixed rejection sequence} of length $n$.
\end{definition}

\begin{definition}
  Consider PCP $\dot{S}$. Two fixed rejection sequences $h, e$ are
  called {\bf equivalent} if $h_1\circ ... \circ h_i \circ \dot{S} \cong e_1 \circ
  ... \circ e_i \circ \dot{S}$ for all   $i \in \{1,...,n\}$
\end{definition}
\begin{definition}
  Let $\dot{K}_n$ denote the $tPCP$ consisting of all pairwise
  comparisons of $n$ different means.
\end{definition}

\begin{theorem}
  Properties of $\circ$:
  \begin{enumerate}
    \item $\dot{S} \circ e_1 \circ e_2 \dot{S} \circ e_2 \circ e_1$
    \item For $e,f,u,v \in E(\dot{S})$, such that $\dot{S} \circ e
      \cong \dot{S} \circ f$, and $\dot{S} \circ e \circ u \cong
      \dot{S} \circ u \circ v$ then $\exists z \in E(\dot{S} \circ f)$
      such that $\dot{S} \circ f \circ z \cong \dot{S} \circ e \circ
      u$. 
  \end{enumerate}
\end{theorem}

\begin{theorem}
  The tPCP $\dot{K}_4$ has 6 distinct (up to isomorphism) fixed
  rejection sequences. Figure \ref{fig:k4} shows a graphical
  representation of the distinct rejection sequences for
  $K_4$. $\dot{K}_5$ has $657$ and $\dot{K}_6$ has $4836277$ it seems
  that this sequence grows in the order of $\mathcal{O}(e^{n^2})$.
\end{theorem}

\begin{proof}
  Proof by exhaustive computer search.
\end{proof}

\begin{figure}

  \caption{Rejection tree for $\dot{K}_4$}
  \label{fig:k4}
\end{figure}

\begin{theorem}
  $\dot{K}_4$ can be represented by the complete graph $K_4$ with 4
  vertices and 6 edges. The resulting null hypotheses are vertices of
  the line-graph $\mathcal{L}{K_4}$. $\mathcal{L}{K_4}$ is an
  octahedron. The exhaustive sets of $\dot{K_4}$ correspond to
  $\mathcal{L}(K_4)$, the 3 faces of the octahedron with differently
  labeled edges representing cliques of size 3 in $K_4$, the 3 edges
  in the complementary graph of $\mathcal{L}(K_4)$, and the $6$
  vertices themselves.
\end{theorem}

\subsection{Connections to matroid theory}

In this section we first show that any (finite (?)) system of null
hypotheses together with its exhaustive sets defines a matroid. The
closure operator of that matroid is the function that for a given
index set assigns the smallest containing exhaustive index set. We
then focus on the specific problem of embedding two-sided pairwise
comparisons in matroid theory. We show that the exhaustive sets of a
tPCP are flats in the cycle matroid of the comparigram, and vice
versa.

Let $X$ denote a random variable describing the outcome of an
experiment, e.g., the results of a randomized clinical trial, with
distribution $F^\theta$ from parametric family $\mathcal{F} =
{F_\vartheta|\vartheta \in \Theta}$. Let $\mathcal{H} \ \{H_i: \theta
\in \Theta_i, \Theta_i \subset \Theta, i \in S\}$ denote a family of
null hypotheses about true parameter $\theta$ with associated index
set $S = \{1,...,n\}$. Let $H_I : \theta \in \bigcap_{i \in I}
\Theta_i$ denote the intersection hypothesis associated with $I$, \ie
that assumes that all $H_i$, $i \in I$ are true simultaneously. For
completeness define $H_\emptyset : \theta \in \Theta$. To shorten
notation we will write $H_I = \bigcap_I \Theta_i$. We define a subset
$I \in S$ to be independent if the corresponding intersection
hypothesis is altered if any index is removed from $I$, that is

\begin{definition}{\bf Independent index set}
\label{def:indep}
  Let $I \subseteq S$. We say that $I$ is independent if $\bigcap_{i
    \in I} \Theta_i \subsetneq \bigcap_{i \in I\setminus \{j\}}
  \Theta_i$ for all $j \in I$. 
\end{definition}

Unfortunately this rather simple definition of independence is not
enough such that the system of intersection hypotheses is a
matroid. An additional requirement is needed.

\begin{theorem}
  Consider a family of hypotheses $\mathcal{H} = \{H_i:\theta \in
  \Theta_i,i \in S\}$. Let $\mathscr{I} = \{I \subseteq S: I \text{
    independent}\}$ denote the set of independent sets of
  $\mathcal{H}$. If for all $I,J \in \mathscr{I}$ with $|I| = |J|+1$,
  holds 
  \begin{align}
    \label{cond:matroida}
    \bigcap_J \Theta_i &\nsubseteq \bigcap_{I\setminus J} \Theta_i,
    \text{ and} \\
    \label{cond:matroidb}
    \bigcap_{J}\Theta_i &\nsupseteq     \bigcup_{I \setminus J} \Theta_i ,
  \end{align}
  then $\mathscr{I}$ defines a matroid $M(\mathcal{H})$ over $S$.
\end{theorem}

\begin{proof}
  We need to show the independence axioms:
  \begin{enumerate}
  \item $\emptyset \in \mathscr{I}$
  \item $I  \in \mathscr{I}: J \subset I \Rightarrow J \in \mathscr{I}$
  \item $I,J \in \mathscr{I}$: $|I|=|J|+1 \Rightarrow \exists j \in
    I\setminus J: J \cup j \in \mathscr{I}$ 
  \end{enumerate}
  The first axiom follows directly from $\Theta_i \neq \Theta$ and
  $H_\emptyset = \Theta$.

  To show the second axiom let, without loss of generality, $I =
  \{1,....,k\}$ $J = \{1,...,l\}$ $l \leq k$. We use induction. For $l
  = k$, $I = J$ and the result is obviously true. For $l = k-1$ assume
  there is some $j \in J$, such that $\bigcap_{J \setminus j} \Theta_i
  = \bigcap_{J} \Theta_i$. Since $I \in \mathscr{I}$, we have
  $\bigcap_{I} \Theta_i \subsetneq \bigcap_{J} \Theta_i$ and
  $\bigcap_I \Theta_i \subsetneq \bigcap_{I \setminus j} \Theta_i =
  \Theta_k \cap \bigcap_{J \setminus j} \Theta_i= \Theta_k \cap \bigcap_{J}
  \Theta_i= \bigcap_I \Theta_i$. Assume that it holds for $l = k - m$
  then using the same arguments it follows that it also holds for $l =
  k - (m + 1)$.

  To show the third axiom some additional assumptions on the sets
  $\Theta_i$ are required. We need to show that for $I,J \in
  \mathscr{I}$ there is a $j \in I \setminus J$ such that $\bigcap_{J
    \cup j}\Theta_i \subsetneq \bigcap_{(J\cup j)\setminus l}
  \Theta_i$ for all $l \in (J \cup j)$. One can write down necessary
  and sufficient conditions for the exchange axiom by enumeration of
  all possible cases of $I,J$. First observe that the axiom holds for
  $J \subset I$ with $l = I \setminus J$ otherwise consider $0 < l <
  k$ and $J = \{i_1,...,i_k\}$ and $I =
  \{i_{k+1-l},...,i_{2k-l+1}\}$. Then assuming that the exchange axiom
  does not hold for $k = 1$, $l = 1$, \ie $J = \{i_1\}$
  and $I =   \{i_2,i_3\}$, we have $\Theta_{i_1} \cap \Theta_{i_2} = \Theta_{i_1}$
  or $\Theta_{i_1} \cap \Theta_{i_2} = \Theta_{i_2}$, and $\Theta_{i_1} \cap
  \Theta_{i_3} = \Theta_{i_1}$ or $\Theta_{i_1} \cap \Theta_{i_3} = \Theta_{i_3}$. Now $\Theta_{i_1} \cap \Theta_{i_2} = \Theta_{i_1}$ that is
  $\Theta_{i_2} \subseteq \Theta_{i_1}$ and $\Theta_{i_1} \cap \Theta_{i_3} =
  \Theta_{i_3}$, \ie $\Theta_{i_1} \subseteq \Theta_{i_3}$, gives $\Theta_{i_2}
  \subseteq \Theta_{i_3}$ which contradicts the independece of $i_2$
  and $i_3$, and the same holds vice versa. Consequently either
  $\Theta_1 \subseteq \Theta_2$ and $\Theta_1 \subseteq \Theta_3$ or
  $\Theta_2 \subseteq \Theta_1$ and $\Theta_3 \subseteq \Theta_1$. Now
  a sufficient and necessary condition for the former is $\Theta_1 \in
  \Theta_2 \cap \Theta_3$ and for the latter $\Theta_2 \cup \Theta_3
  \subset \Theta_1$. As these conditions are necessary and sufficient
  negating them provides necessary and sufficient conditions for the
  exchange axiom that is:
  \begin{description}
  \item[$k=1,l=1$:] $\Theta_{i_1} \nsubseteq \Theta_{i_2} \cap
    \Theta_{i_3}$ and $\Theta_{i_2} \cup \Theta_{i_3} \nsubseteq \Theta_{i_1}$
  \item[$k=2,l=1$:] $\Theta_{i_1} \cap \Theta_{i_2} \nsubseteq
    \Theta_{i_3} \cap \Theta_{i_4}$ and $\Theta_{i_3} \cup \Theta_{i_4} \nsubseteq \Theta_{i_1} \cap \Theta_{i_2}$
  \item[$k=2,l=0$:] $\Theta_{i_1} \cap \Theta_{i_2} \nsubseteq
    \Theta_{i_3} \cap \Theta_{i_4} \cap \Theta_{i_5}$ and
    $\Theta_{i_3} \cup \Theta_{i_4} \cup \Theta_{i_5} \nsubseteq \Theta_{i_1} \cap \Theta_{i_2}$
  \item[$k=3,l=2$:] $\Theta_{i_1} \cap \Theta_{i_2} \cap \Theta_{i_3}\nsubseteq
    \Theta_{i_4} \cap \Theta_{i_5}$ and $\Theta_{i_4} \cup
    \Theta_{i_5} \nsubseteq \Theta_{i_1} \cap \Theta_{i_2} \cap \Theta_{i_3}$
  \item[....]
  \item[General $k,l$:] $\bigcap_{r \leq k-l} \Theta_{i_r} \nsubseteq
    \bigcap_{k+1 \leq s \leq 2k-l+1} \Theta_{i_s}$ and $\bigcup_{k+1
      \leq s \leq 2k-l+1} \Theta_{i_s} \nsubseteq \bigcap_{r \leq k-l} \Theta_{i_r}$.
  \end{description}
  For $m = \max \{|I|: I \in \mathscr{I}\}$ above conditions need to
  hold for all $k,l$ $k \leq m-2$ and $\max\{0,2-k\} \leq l \leq
  k-1$.
\end{proof}

For notational purposes it is practical to define a dependence
relation ``$\sim$''.

\begin{definition}
  For $j \in S$ $I \subseteq S$ define $j \sim I$ iff there is a subset
  $J \subseteq I$ such that $\bigcap_J \Theta_i \subseteq \Theta_j$.
\end{definition}

It follows immediatly that $\nsim$ is independence in the above sense.

\begin{proposition}
  $I \in \mathscr{I}$ if and only if for all $i \in I$, $i \nsim
  I\setminus i$.
\end{proposition}

\begin{comment}
  Some interesting notes on above matroid.
  \begin{itemize}
  \item We have $\emptyset \sim i$ for all $i \in S$, but $i \nsim
    \emptyset$ unless $\Theta_i = \Theta$. Essentially this means that
    there are no loops in $M(\mathcal{H})$
  \item For hypotheses $H_i,H_j$ such that $\Theta_j \subset \Theta_i$
    we have $i \sim j$ and this for example is the case for combined
    non-inferiority superiority testing.
  \item The rank function of an index set $I$ is given by the
    cardinality of its maximally independent subsets, \ie $\rho I =
    \max \{|J|:J\subseteq I, J \in \mathscr{I}\}$.
  \item To give some intuition to condition \ref{cond:matroida} one
    may see that for the simple case of $J = \{i\}$ and $I = \{j,l\}$
    the condition requires that for parallels $i \sim j$ we have $i
    \nsim l$ as $i \sim l$ would imply also $j \sim l$ which is in
    contradiction to the independence of $I$. For $|J|>1$ the
    condition does not necesarily involve parallels.
  \item As a system of null hypotheses that does not conform to
    \ref{cond:matroida} consider the following set of hypotheses:
    \begin{align*}
      H_1:& (\mu_1 \leq 0) &, (\mu_2 \leq 0) \\
      H_2:& \mu_1 &\leq 0 \\
      H_3:& \mu_2 &\leq 0,
    \end{align*}
    according to Definition \ref{def:indep} $\{1\},\{2,3\} \in
    \mathscr{I}$, however $\Theta_1 = \Theta_2 \cap \Theta_3$. From a
    testing perspective we may simple drop $H_1$ as it is simply $H_2
    \cap H_3$ which would be tested anyways.
  \item Now consider for $\delta > 0$,
    \begin{align*}
      H_1:& (\mu_1 \leq 0) &, (\mu_2 \leq 0) \\
      H_2:& \mu_1 &\leq \delta \\
      H_3:& \mu_2 &\leq \delta.
    \end{align*}
    Again, Condition \ref{cond:matroida} does not hold. However, in
    this case replacing the system by an equivalent system of null
    hypotheses is not as straight forward. A solution may be to
    replace $H_1$ by:
    \begin{align*}
      H_{1a}:& \mu_1 &\leq 0 \\
      H_{1b}:& \mu_2 &\leq 0,
    \end{align*}
  \end{itemize}
\end{comment}

Next we show an important connection between matroid theory and
multiple testing theory

\begin{theorem}
  Consider a family of null hypotheses for which conditions
  (\ref{cond:matroida}-\ref{cond:matroidb}) hold. Then an index set $I \subseteq S$ is
  weakly exhaustive if and only if $I$ is a flat in $M$.
\end{theorem}

\begin{proof}
  Not yet done!
\end{proof}

To get a result on strictly exhaustive set we need to restrict
condition (\ref{cond:matroida}) further.

\begin{theorem}
  Consider a family of null hypotheses $\mathcal{H}$ for which above
  the more stringent conditions hold:
  \begin{align}
    \label{cond:matroid2}
    \bigcap_J \Theta_i &\nsubseteq \bigcup_{I\setminus J} \Theta_i,
    \text{ and} \\
    \bigcap_{J}\Theta_i &\nsupseteq     \bigcup_{I \setminus J} \Theta_i ,
  \end{align}
  % and additionally $\bigcap_S \Theta_i^C \neq
  % \emptyset$ (practically that means that all alternatives may be true
  % at the same time).
  Then independence as defined in Definition \ref{def:indep} gives a
  matroid for   which every flat corresponds to a strictly exhaustive
  index set and vice versa.
\end{theorem}

\begin{proof}
  Not yet done!
  % We first show by contradiction that every flat is strictly
  % exhaustive: Consider $I$ is a flat but that $I$ is not strictly
  % exhaustive. As $I$ is a flat, we have for every $l \in S\setminus
  % I$, $l \sim I$. Let $J \in \mathscr{I}$ be a maximally independent
  % set in $I$. Then $J \cup l \in \mathscr{I}$ and $|J \cup l| =
  % |J|+1$. If $I$ were not strictly exhaustive, then $\bigcap_{I}
  % \Theta_i \cap \bigcap_{S\setminus I} \Theta^c_i = \emptyset$. 
\end{proof}

\begin{comment}
  Above restriction (\ref{cond:matroid2}), already for the simple case
  of $|J|=1$, is no longer a statment about the parallels of
  $M$. 
\end{comment}

On the other hand if we look at a system of hypotheses and consider
exhaustiveness as defined in [hommel, bergmann] we can define a
matroid.

\begin{theorem}
\label{the:matroid}
  Consider a (finite) family of null hypotheses $\mathcal{H} = \{H_i,
  i \in S\}$ with assiciated index set $S$. Let $\mathcal{P}(S) = \{I
  \subseteq S\}$ denote the power set of $S$ and $SE:\mathcal{P}(S)
  \rightarrow \mathcal{P}(S)$ the function that assigns to each $I
  \subseteq S$ the smallest exhaustive set $SE(I)$ such that $I \in
  SE(I)$. Taken as a closure operator $SE$ defines a matroid $M$ on
  $\mathcal{P(S)}$.
\end{theorem}

\begin{proof}
  We need to show that the closure axioms hold for $SE$:
  \begin{enumerate}
  \item $I \subseteq SE(I)$ holds by definition
  \item $J \subseteq I \Rightarrow SE(J) \subseteq SE(I)$ and,
  \item $SE(I) = SE(SE(I))$ and,
  \item $j \notin I$, $j \in SE(I \cup i) \rightarrow i \in SE(I \cup
    j)$ follow immediately from Lemma $1.5$ in [Bernhard, 91].
  \end{enumerate}
\end{proof}

\subsection{Matroid of tPCP}

\begin{definition}
  Let $\dot{S}$ be a tPCP with associated comparigram $G(\dot{S})$ we
  denote the cycle matroid of $G(\dot{S})$ by $M(\dot{S})$ or simply
  $M$ in case were clear from the context. 
\end{definition}

For convenience we reproduce here some of the properties of the cycle
matroid $M$ of a graph $G$.

\begin{proposition}{Properties of $M$ (Welsh, 2010)}
\label{the:cyclemat}
  \begin{enumerate}
  \item The cycles of $G$ are the circuits of $M$ 
  \item The spanning forests of $G$ are the bases of $M$
  \item If $G$ is connected the bases of $M$ are the spanning trees of $G$
  \item A set $X$ of edges of $G$ is independent in $M$ if and only if
    $X$ contains no cycle
  \item The rank of $M$ is $|V(G)| - k(G)$ where $k(G)$ denotes the
    number of connected components in $G$
  \item For any subset $A \subseteq E(G)$ the rank of $A$ in $M$ is
    given by
    \begin{displaymath}
      \rho_G(A) = |V(A)| - k(A)
    \end{displaymath}
    where $V(A)$ and $k(A)$ denote the set of vertices and number of
    connected componenents in the subgraph of $G$ induced by $A$
  \item Loops, and parallel edges correspond in $G$ and $M$
  \item Let $A \subset E(G)$, let $e \in E(G)\setminus A$. Then $e$
    belongs to the closure of $A$ in $M(G)$ if and only if there is a
    cycle $C$ of $G$ with
    \begin{displaymath}
      e \in C \subseteq A \cup e.
    \end{displaymath}

  \end{enumerate}
\end{proposition}

Next we give some useful notation and results on exhaustive index sets
of pairwise comparisons.

\begin{proposition}
  Let $X \subseteq S$ then there is a unique collection of mutually
  exclusive index sets $s(X) = \{X_i = \{i_1,...,i_{m_i}\} \subseteq
  V(\dot{S}), i \in \{1,...,M\}: X_i \cap X_j = \emptyset\}$, such
  that $H_X$ is the intersection of hypotheses of the form $H_{X_i}:
  \mu_{i_1} = ... = \mu_{i_{m_i}}$, \ie $H_X = \bigcap_{1:M}
  H_{X_i}$. We call $s(X)$ the {\em associated partition hypothesis}.
\end{proposition}

It follows directly that

\begin{theorem}
  Consider $I, J \in R$, $H_I = H_J$ if and only if $s(I) = s(J)$. 
\end{theorem}

The following Lemma will be helpfull for the proof of Theorem
\ref{the:compmat}.

\begin{lemma}
  Let $X \subseteq R$ then $H_{s(X)} = H_X$. 
\end{lemma}

\begin{proof}
  Consider that $H_{\sigma X} \neq H_X$. $\sigma X$ is exhaustive and
  $X \subseteq \sigma X$ follows by definition. Therefore, $H_{\sigma
    X}$ is a proper subset of $H_X$. Consequently there exist $A \in
  s(X)$ and $B \in s(\sigma X)$ such that $A$ is a proper subset of
  $B$. Now consider $\bigcup_{X_i \in (s(\sigma X)\setminus B)\cup A}
  H_{X_i}$ and $Y = \{(a,b) \in R: a,b \in X_i \in (s(\sigma
  X)\setminus B)\cup A\}$ then $X \subset Y$ and since $H_x \subset
  H_Y$ and $X$ exhaustive $Y \subset \sigma X$. However, $Y$ is
  exhaustive, which contradicts that $\sigma X$ is the smallest
  exhaustive index set in $\mathcal{P}(R)$.
\end{proof}

From this one immediately sees how the smallest containing exhaustive
index sets can be constructed.

\begin{corollary}
  $SE(X) = \{(a,b) \in R:\exists A \in s(X), a,b \in A\}$
\end{corollary}

Using these properties we show the following theorem.

\begin{theorem}
  Consider a tPCP $\dot{S}$ with comparigram $G$. Let $M$ be the
  associated cycle matroid of $G$ then an index set $I \subseteq R$ is
  exhaustive if and only if it is a flat in $M$.
\end{theorem}

\begin{proof}
  Consider $I$ exhaustive and assume that is not a flat. Let $x \in
  \sigma I \setminus I$. Then $\rho(I) = |V(I)| - k(I) = \rho(I \cup
  x) = |V(I \cup x)| - k(I \cup x)$. By adding an edge to $I$, $|V(I
  \cup x)| = |V(I)| + d$ where $d$ is either $0$, $1$ or $2$. Iince
  $G$ simple and $I$ exhaustive, any endpoints of $x$ in $V(I)$ has to
  be isolated in $I$. For $d = 0$, $x$ connects isolated vertices $a,
  b \in V(I)$ accordingly that $k(I \cup x) = k(I) - 1$ which is a
  contradiction. Iimilarly, $d = 1$ gives $k(I\cup x) = k(I)$ and $d =
  2$ leads to $k(I\cup x) = k(I)+1$ in either case the rank of $I$ is
  increased by $1$.

  Assume that $I$ is a flat in $M$ but $I$ is not
  exhaustive. Consequently there exists $I' \subseteq R$ such that $I$
  is a proper subset of $I'$ and $H_I = H_{I'}$. Therefore we have
  $s(I) = s(I')$. Let $e = (a,b) \in  I'\setminus I$ then there is $A
  \in s(I)$ such that $a,b \in A$. As $e \notin I$, there is at least
  one $c \in A$ such that $(a,c),(b,c) \in I$. Now
  $C = \{(a,b),(b,c),(a,c)\} \in \mathcal{C}$ such that $(a,b) \in
  C\setminus I$ which contradicts that $I$ is flat.
\end{proof}

We next show how the operation that for each index set identifies the
smallest containing exhaustive index set induces a matroid.

\begin{theorem}
  Consider a tPCP $\dot{S}$ for any $I \subseteq R$ let
  \begin{displaymath}
    \hat{\sigma} I = SE(I)
  \end{displaymath}
  where $SE(I)$ is the smallest exhaustive index set in
  $\mathcal{P}(R)$ cointaining $I$. Then $\hat{\sigma} I$ is the closure
  operator of a matroid $M$.
\label{the:compmat}
\end{theorem}

We now proof Theorem \ref{the:compmat}.

\begin{proof}
  Actually this follows directly from Theorem \ref{the:matroid}. I
  leave the old proof for completeness.

  We need to show
  \begin{enumerate}
  \item $X \subseteq \hat{\sigma} X$, 
  \item For $Y \subseteq X$ then $\hat{\sigma}(Y) \subseteq \hat{\sigma}(X)$, and 
  \item $\hat{\sigma} X = \hat{\sigma} \hat{\sigma} X$
  \item If $y \notin \hat{\sigma} X$, $y \in \hat{\sigma}(X \cup x)$ then $x \in
    \hat{\sigma}(X \cup y)$.
  \end{enumerate}
  (1) holds by definition.

  (2) assume $Y \subset X$ and $\hat{\sigma} Y \setminus \hat{\sigma} X \neq
  \emptyset$. Let $h \in \hat{\sigma} Y \setminus \hat{\sigma} X$ and let $h =
  (a,b)$. Then there is $A \in s(Y)$ with $a,b \in A$ but no $B \in
  s(X)$ with $a,b \in B$. Since $Y \subset X$, for all $A \in s(Y)$,
  $B \in s(X)$ such that $A \subset B$. Which leads to a
  contradiction.

  (3) follows trivially.

  (4) let $x = (a,b)$, there is no $A \in s(X)$ with $a,b \in A$ but
  there is some $B \in s(X \cup y)$ with $a,b \in B$. Let $y = (c,d)$
  we can distinguish 3 cases: (a) $\{c,d\} \cap B = \emptyset$ for all
  $B \in s(X)$ then $s(X \cup y) = s(X) \cup \{c,d\}$ such that $B =
  \{c,d\}$ and $x = y$. (b) Ther is $C \in s(X)$, $c \in C$ but
  no $D \in s(X)$ with $d \in D$. Then $s(X \cup y) = (s(X)\setminus C) \cup (C
  \cup d)$ and since this is the only set that has changed by adding
  $y$, $x \in (C\cup d)$, therefore $b = d$ and consequently
  $c,d \in C\cup b$ and also $y \in \hat{\sigma}(X \cup x)$. (c) There are
  $C,D \in s(X)$ such that $c \in C$ and $d \in D$ then $s(X \cup y) =
  (s(X)\setminus \{C,D\})\cup (C \cup D)$ and since $C\cup D$ is the
  only set changed from $s(X)$ $a \in C$ and $b \in D$ such that also
  $y \in \hat{\sigma}(X\cup x)$. 
\end{proof}

As the cycle matroid of the comparigram and the matroid defined by
$\hat{\sigma}$ have identical closed sets it follows directly that
they are the same matroid

\begin{corollary}
  The matroid $\hat{M}$ on $\dot{S}$ defined by the closure operation 
  $\hat{\sigma}I = SE(I)$, $I \subseteq \dot{S}$ is the cycle matroid
  of the comparigram $G(\dot{S})$
\end{corollary}

Accordingly all porperties given in \ref{the:cyclemat} also hold for
the matroid on exhaustive sets.

The following theorem shows an important connection between the flats
of the cycle matroid of a complete graph and the partitions of a
finite set.

\begin{theorem}{[Welsh]}
\label{the:partition.lattice}
  The lattice of flats of $M(K_n)$ is isomorphic to the distributive
  lattice of all partitions of a set with $n$ elements ordered by
  refinement. 
\end{theorem}

Translated to pairwise comparison problems this means that the system
of exhaustive sets and the associated partial order induced by the 
inclusion operator is just corresponds to the partitions of the set of
independent population means ordered by refinement. This is not
surprising as $s(X)$ associates each index set with a partition of the
set $V(R)$ and restricted to the family of flats $\mathcal{L}(R)$,
$s(X)$ is one to one. Recognizing that any partition of means
corresponds to a decomposition of $K_n$ into cliques this provides an
alternative proof to the main result of Weichert on two-sided pairwise
comparisons.

The properties of set partitions have been already extensively
studied. For example, this correspondence provides us with efficient
algorithms to generate all exhaustive sets. Further, the number of
exhaustive sets resulting from all pairwise compinations is given by
the Bell numbers.

If not all pairwise comparisons are of interest we can still apply
Theorem \ref{the:partition.lattice} using the following results on
matroid minors. Let $|$ and $.$ denote the operators for restriction
and contraction, then

\begin{theorem}{Theorem 2 [Welsh, p.66]}
  If $\mathcal{L}$ is the geometric lattice of the simple matroid M,
  the interval $[o,T]$ is the lattice of $M|T$, for any flat $T$ of
  $M$. 
\end{theorem}

\begin{theorem}
\label{the:interval}
  Consider a tPCP $\dot{S} \in \dot{\mathcal{T}}(n)$. The exhaustive
  sets of $\dot{S}$ generate a geometric lattice
  $\mathcal{L}(\dot{S})$ that is isomorphic to: 
  \begin{displaymath}
    [\emptyset,\sigma_{K_n} S] \cap S,
  \end{displaymath}
  which is the interval sublatticee of the partition lattice of rank
  $n$ where for each subset the intersected with the set of
  comparisons $S$ is taken. 
\end{theorem}

\begin{proof}
  The closure operator of $M|S$ is given by $\sigma_R(A) = \sigma A
  \cap S$. Consider $L = \mathcal{L}(M|S)$ and $K = \mathcal{L}(M|\sigma
  S)$ then it can be easily seen that for $A,B \in K$ there are unique $A',B'
  \in K \cap R$ such that $(A \land B)' = A' \land B'$ and $(A
  \lor B)' = A' \lor B'$. 
\end{proof}

Above result does show a way how to construct the exhaustive sets for
a general pairwise comparison problem. The intervall lattice
$[\emptyset,=\sigma_{K_n} S]$ can be constrcuted by refinement of
$V(\sigma_{K_n} R)$.  However, if $\sigma_{K_n} S = K_n$ the
corresponding intervall lattice is identical to the full
lattice. Nevertheless one can reduce the problem in the following
way. Perfom one refinement step of $\sigma_{K_n} S$ and intersect with
$S$ to get all hyperplanes of $M|S$. Then all consecutive flats of
$M|S$ can be found by closing the set of hyperplanes with regard to
intersection.

\subsection{Hierarchical testing of pairwise comparisons}

\begin{theorem}
  Consider a hierarchical testing procedure that tests $H_{i_r}$ at
  level $\alpha$ according to a prefixed order $(i_1,...,i_n)$. Assume
  that hypotheses with indeces $i_1$ through $i_k$ have been
  rejected. If for $k+1$ to $k+l$ there is $i_s$, $s \leq k$ such that
  \begin{equation}
    \label{eq:condh}
    \rho \{i_s,i_{k+1}\} = \rho \{i_s,i_{k+1}...,i_{k+l}\},
  \end{equation}
  hypotheses each $H_{i_{k+1}}$ through   $H_{i_{k+l}}$ can be tested
  simultaneously at full level $\alpha$. 
\end{theorem}

\begin{proof}
  Due to \ref{eq:condh} for any $k+1 \leq m' < m \leq k+l$, $i_m \in
  \sigma \{i_s,i_{m'}\}$ and consequently also $i_s \in \sigma
  \{i_m,i_{m'}\}$. Therefore $i_s \in \sigma J$ for all $J$ with
  $i_m,i_{m'} \in J$. Since $p_s \leq \alpha$ all intersection
  hypotheses corresponding to exhaustive sets containing $i_{m}$
  and $i_{m'}$ for all $k+1 \leq m' < m$ have been rejected. Such that
  $p_{i_{m}} \leq \alpha$ rejects $H_{i_{m}}$.
\end{proof}

\subsection{Graphical procedure for pairwise comparison problems}

Above properties of tPCPs can be used to derive weighted multiple
testing procedures, that
\begin{enumerate}
\item control the FWER,
\item use a weighting strategy defined by a graph,
\item exploits logical dependencies between hypotheses to yield
  improved power,
\item produces significance statements that are consistent with
  logical dependencies between hypotheses.
\end{enumerate}

Some additional notation let $\gmcp$ denote a graphical weighting
procedure, such that $\gmcp(I)$ gives the weights for intersection
hypothesis $H = \bigcup_{i \in I} H_i$. $gMCP(n)$ is the class of all
graphical procedures for $n$ hypotheses. Define $h(S)$ the set of
hyperplanes of $S$ in $M$.

\begin{algorithmic}
  \REQUIRE $\dot{S} \in \dot{\mathcal{T}(m)}$, $\gmcp \in gMCP(n)$, 
  $p \in [0,1]^n$, $\alpha$
  \ENSURE $n = |E(\dot{S})| \leq m*(m-1)/2, \alpha \in [0,1]$
  \STATE $w \gets \gmcp(R)$
  \STATE $R \gets \{i \in R: p_i \leq w_i \alpha\}$
  \STATE $S \gets S \setminus R$
  \STATE $EI \gets h(\dot(S))$
  \WHILE{$|R| < n$}
  \STATE For all $E \in EI$, $w_E \gets \gmcp(E)$
  \STATE For all $i \in S\setminus R$, $w_i^\star \gets \min_{E \in
    EI}w_{i,E}$
  \STATE $r \gets \{i:p_i \leq w_i^\star*\alpha$
  \STATE $v \gets \{i \in S\setminus R:i \notin E, \forall E \in EI\}$
  \IF{$r = v = \emptyset$}
  \STATE Reject $R$ and stop
  \ENDIF
  \STATE $R \gets R \cup r$
  \STATE $S \gets S\setminus r$
  \STATE $EI \gets h(\dot(S))$
  \ENDWHILE
  \STATE Reject $R$ and stop
\end{algorithmic}

Taking the minimum weight accross all flats in above procedure may be
improved:

\begin{algorithmic}
  \REQUIRE $\dot{S} \in \dot{\mathcal{T}(m)}$, $\gmcp \in gMCP(n)$, 
  $p \in [0,1]^n$, $\alpha$
  \ENSURE $n = |E(\dot{S})| \leq m*(m-1)/2, \alpha \in [0,1]$
  \STATE $w \gets \gmcp(R)$
  \STATE $R \gets \{i \in R: p_i \leq w_i \alpha\}$
  \STATE $S \gets S \setminus R$
  \STATE $EI \gets h(\dot(S))$
  \WHILE{$|R| < n$}
  \STATE For all $E \in EI$, $w_E \gets \gmcp(E)$
  \STATE $r \gets \{i:\forall E \in EI, i \in E,p_i \leq w_i \alpha\}$
  \STATE $v \gets \{i \in S\setminus R:i \notin E, \forall E \in EI\}$
  \IF{$r = v = \emptyset$}
  \STATE Reject $R$ and stop
  \ENDIF
  \STATE $R \gets R \cup r$
  \STATE $S \gets S\setminus r$
  \STATE $EI \gets h(\dot(S))$
  \ENDWHILE
  \STATE Reject $R$ and stop
\end{algorithmic}

Comments: Above procedures are valid for all families of hypotheses
for which exhaustive sets are identical with strictly exhaustive
sets. The crucial part of the algorithms is finding $h(S)$. For $tPCP$
this can be achieved using Theorem \ref{the:interval}.

\subsection{General contrast tests}

Realizing that each pairwise comparison may be represented by a
general contrast test we next investigate the connections between such
tests and vectorial matroids.

Some interesting observations:
\begin{itemize}
\item Pairwise comparisons (as a contrast) are represented by a vector
  having a $1$ at the component corresponding to the first and a $-1$
  at the component corresponding to the second mean in the
  comparison. For example if $m = 4$, $H_{13}: \mu_1 = \mu_3$
  corresponds to  $c_{13} = (1,0,-1,0)$ 
\item If we consider only two-sided hypotheses the $c_{ij} = -c_{ji}$ and
  $-c_{ij} = c_{ji}$  represent the same hypothesis. Consequently,
  without loss of generality we can represent $H_{ij}$ by a vector
  $e_{ij}$ over $GF(2)$ with entries $1$ at $i$ and $j$.
\item From representation theory over matroids we now that every graph,
  and consequently tPCP, may be represented over $GF(2)$ by using the
  incidence matrix.
\item Given that a representation is defined as a map that conserves
  the rank. The maximum rank of a tPCP is $m-1$.
\item It can be easily seen that the corresponding contrast matrix
  maybe reduced to a rank $\leq m-1$ upper diagonal matrix.
\item A general contrast test may, however, have up to rank $m$ and
  can not be represented over $GF(2)$.
\end{itemize}

\begin{definition}
  Consider $X \sim N(\bs{\mu},\sigma \mathbb{1}_m)$ and a contrast $A
  \in \mathbb{R}^{m \times n}$ with column vectors $a_i \in
  \mathcal{R}^m$. We call the MTP given by by the family $\mathcal{H}
  = \left\{H_i: a_i\mu = 0; i \in \{1,...,n\}\right\}$ a (finite)
  linear  contrast problem (LCP).
\end{definition}

\section{One sided problems}

So far we have only considered two sided hypotheses \ie point
hypotheses. In many practical applications however the questions of
interest are better addressed by one sided null hypotheses. For
example in a clinical trial one generally only interested whether the
a new treatment is better than some comparator.

\end{document}

### 2013-08-08 paper/draft.tex
Added section with one-sided hypotheses

@@
 \newcommand{\ie}{{\em i.e.,}~}
-\newcommand{\cf}[1]{\mathbf{1}_{\left\{#1\right\}}}
+\newcommand{\indf}[1]{\mathbf{1}_{\left\{#1\right\}}}
+%\newcommand{\text}[1]{\textrm{#1}}
+\newcommand{\R}{\mathbb{R}}
 
@@
 
+From conditions \ref{} it can be easily seen that while problems of
+one sided hypotheses may be treated by matroids, already for simple
+problems the correspondence between flats and strictly exhaustive sets
+is not one to one. Nevertheless an alternative approach to this class
+of problems can be found in the theory of oriented matroids.
+
+First let us define a one sided multiple testing problem
+
+\begin{definition}
+  Let $\bs{X}$ denote an $m$ dimensional random vector with
+  distribution function $P_{\bs{\mu}}$ from some parametric family
+  $\mathscr{P} = \{P_{\bs{\theta}}: \bs{\theta} \in \mathbb{R}^m\}$. A
+  family of {\em one sided linear contrast hypotheses} is defined by a
+  contrast matrix $\bs{A} \in \mathbb{R}^{m\times n}$ and vector
+  $\bs{b} \in \mathbb{R}^n$ that define a the family of null
+  hypotheses of the form $\mathscr{H} = \{H_i: \bs{A_i}^t\bs{\mu} \leq
+  b_i,i \in \{1,...,n\}\}$, where $\bs{A}_i$ denotes the $i$-th column
+  of $\bs{A}$ and $b_i$ the $i$-th componenth of $\bs{b}$.
+\end{definition}
+
+Now observe that each hypothesis $H_i$ corresponds to a hyperplane
+$h_i = \{\bs{x} \in \mathbb{R}^m: \bs{A}_i^t\bs{x} = b_i\}$ that divides
+$\mathbb{R}^m$ into two regions $h_i^+ = \{\bs{x} \in \mathbb{R}^m:
+\bs{A}_i^t\bs{x} < b_i\}$ and $h_i^- = \{\bs{x} \in \mathbb{R}^m: \bs{A}_i^t\bs{x} >
+b_i\}$ giving rise to a sign function:
+\begin{equation}
+  \label{eq:sign}
+  \delta_i(\bs{x}) = \left\{
+\begin{array}{ll}
+  + & \text{if } \bs{x} \in h_i^+\\
+  0 & \text{if } \bs{x} \in h_i \\
+  - & \text{if } \bs{x} \in h_i^-\\
+\end{array}\right.
+\end{equation}
+
+Further for every $\bs{x} \in \mathbb{R}^m$ we have  sign vector
+$\bs{\delta}{\bs{x}}= (\delta_1(\bs{x}),...,\delta_n(\bs{x}))$. Now
+for each $i$ and some parameter value $\bs{\theta} \in \bs{\Theta} = \R^m$,
+$\delta_i(\bs{\theta})$ tells us whether the corresponding value is under the null
+hypotheses on its border, or under the alternative. Accordingly,
+$\bs{\delta}(\bs{\theta})$ tells us in which combination of true and
+false null hypotheses $H_i$ correspond to $\bs{\theta}$. Let
+\begin{equation}
+\label{eq:faces}
+\mathcal{F}(\bs{A},\bs{b} = \{\bs{\delta}(\bs{x}): \bs{x} \in \R^m\}
+\end{equation}
+
+and for some $F \in \mathcal{F}$ define
+\begin{equation}
+  \label{eq:partition}
+  \Theta_F = \{\bs{x} \in \R^m: \bs{\delta}(\bs{x}) = F\}
+\end{equation}
+
+then we have $\bigcup_{F \in \mathcal{F}} \Theta_F = \bs{\Theta} =
+\R^m$ and for any $\Theta_F \cap \Theta_G = \emptyset$ for any $F \neq
+G$. Consequently, $\{\Theta_F:F \in \mathcal{F}$ represents a
+partition of the parameter space and each element in $\mathcal{F}$
+uniquely identifies some {\em non-empty} element of that partition. Consequently,
+following the Partitioning Principle \cite[]{finner2002partitioning}
+if we have level $\alpha$ tests $\varphi_F(X)$ for each $F \in
+\mathcal{F}$ a test that rejects $H_i$ for observed values $X$ if
+$\varphi_F(X) = 1$ for all $F$ with $F_i \geq 1$ controls the FWER in
+the strong sense. For example one may use  weighted Bonferroni tests
+\begin{equation}
+  \label{eq:bonferroni}
+  \varphi_F = \indf{\min_{i:F_i \geq 0}p_i \leq w_{F,i}}
+\end{equation}
+where the $w_{F,i}$ are can be computed using a graphical approach.
+
+Note that as $\Theta_F$, $F \in \mathcal{F}$ the $\mathscr{L} = \{I \subset \{1,...,m\}: \exists F \in
+\mathcal{F}: F_i \geq 0 \forall i \in I\}$ are all strictly exhaustive sets.
+
+The question remains how to compute $\mathcal{F}$. For this observe
+that the $h_i$ define a hyperplane configuration and $\mathcal{F}$ is
+just the face lattice of the corresponding oriented
+matroid. Consequently, $\mathcal{F}$ can be found using algorithms
+for arrangement construction from polyhedral programming
+\cite{Avis-KF-92,Avis-KF-96,Ferrez-KF-Liebling-01} which are
+polynomial in time and input size and compact in space!
+
+\begin{definition}
+  Consider a family of hypotheses $\mathscr{H} =$
+\end{definition}
+
 \end{document}

### 2013-08-08 presentation/presentation.tex
some changes on modesim

--- 2013-07-31 presentation/presentation.tex
@@
 
-\begin{frame}
-\frametitle{Circuit axioms}
+% \begin{frame}
+% \frametitle{Circuit axioms}
 
-Consider $S$ and a collection of subsets $\mathscr{C}$ for which holds
-\begin{enumerate}
-\item $C_1 \neq C_2 \in \mathscr{C}$, then $C_1 \nsubseteq C_2$,
-\item $C_1 \neq C_2 \in \mathscr{C}$, and $i \in C_1 \cap C_2$ then
-   $\exists C_3 \subseteq (C_1 \cup C_2)$,
-\end{enumerate}
-then $\mathscr{C}$ is the set of circuits of a matroid $M$ on $S$
-\begin{itemize}
+% Consider $S$ and a collection of subsets $\mathscr{C}$ for which holds
+% \begin{enumerate}
+% \item $C_1 \neq C_2 \in \mathscr{C}$, then $C_1 \nsubseteq C_2$,
+% \item $C_1 \neq C_2 \in \mathscr{C}$, and $i \in C_1 \cap C_2$ then
+%    $\exists C_3 \subseteq (C_1 \cup C_2)$,
+% \end{enumerate}
+% then $\mathscr{C}$ is the set of circuits of a matroid $M$ on $S$
 
-\item \textbf{/Some illustration hereabouts/}
+% \begin{itemize}
 
-\end{itemize} % ends low level
-\end{frame}
+% \item \textbf{/Some illustration hereabouts/}
+
+% \end{itemize} % ends low level
+% \end{frame}
 
@@
 \end{block}
-
 \end{frame}

### 2013-08-12 exportfunctions/functions.R
forgot to push no idea?

@@ -48,7 +48,8 @@ exportUDgraph <- function(graph, file = NULL,package="TikZ", scale=1, alpha=0.05
 				" at (",x+offset[1],"bp,",-y-offset[2],"bp)",
 				"[draw,circle,",ifelse(missing(nodeTikZ),"",paste(nodeTikZ,", ",sep="")),double,"fill=",nodeColor,"]",
                                 ##  " {$",node,"$ \\nodepart{lower} $",weight,"$};",sep="")
-                                  " {$",sub("H","\\\\mu_",node),"$};",sep="")
+                                ##  " {$",sub("H","\\\\mu_",node),"$};",sep="")
+                                  " {$",node,"$};",sep="")
 		tikz <- paste(tikz, nodeLine,sep="\n")			
 	}
 

### 2013-08-13 paper/draft.tex
forgot to push no idea?
put the OM result into theorem form

@@
 
+\subsection{Arrangements of hyperplanes}
+\label{sec:hyperplanes}
+
 Now observe that each hypothesis $H_i$ corresponds to a hyperplane
-$h_i = \{\bs{x} \in \mathbb{R}^m: \bs{A}_i^t\bs{x} = b_i\}$ that divides
+$h_i = \{\bs{x} \in \mathbb{R}^m: \bs{A}_i^t\bs{x} = b_i\}$. The
+family of null hypotheses $\mathscr{H}$ therefore is full determined
+by an arrangement of hyperplanes $\mathscr{A} = \{h_i:i=
+\{1,...,n\}\}$.
+
+\begin{definition}
+  An arrangement of hyperplanes is said to be regular if there is a
+  subset of $\mathscr{A}$ whose intersection is a single point
+\end{definition}
+
+Two important cases for non-regular hyperplane arrangements come to
+mind. First, all pairwise intersections of hyperplanes are empty. If
+that is the case then all hypotheses are nested. Such a case is very
+simple an appropriate testing algorithm would be a hierarchical
+approach starting with the {\em smallest} null hypotheses. Second, all
+hypotheses intersect in more than one point. In this case, however, we
+know from basic linear algebra that the intersection is an affine
+subspace. Consequently one can find a hyperplane that is orthogonal to
+this subspace onto which we can project $\mathscr{A}$ to get a central
+regular arrangement of lower dimensionality. An example of this would
+be the family of hypotheses defined by contrast matrix
+
+\begin{equation}
+  \label{eq:example:ordered}
+  \bs{A} = \left\{
+    \begin{array}{ccc}
+      1 & -1 & 0 \\
+      1 & 0 & -1 \\
+      0 & 1 & -1 
+    \end{array}
+  \right\}
+\end{equation}
+
+and vector $\bs{b} = (0,0,0)^t$. Here all points of the form $(x,x,x)$, $x
+\in \R^3$ are a solution to the corresponding system of linear
+equations. On can now project use the plane $x_1 + x_2 + x_3 = 0$ as a
+ground and represent the hyperplanes corresponding to $\bs{A}$, and
+$\bs{b}$ by the lines $l_1,l_2,l_3$
+{\tt Derive the appropriate projection onto the plane rectangular to
+  $(1,1,1)x=0$ and put here}
+
+Notice that each hyperplane $h_i$ divides
 $\mathbb{R}^m$ into two regions $h_i^+ = \{\bs{x} \in \mathbb{R}^m:
 \bs{A}_i^t\bs{x} < b_i\}$ and $h_i^- = \{\bs{x} \in \mathbb{R}^m: \bs{A}_i^t\bs{x} >
-b_i\}$ giving rise to a sign function:
+b_i\}$. For any point $x$ in the parameter space and elementary
+hypotheses $H_i$ we define sign function: 
 \begin{equation}
@@
 
-Further for every $\bs{x} \in \mathbb{R}^m$ we have  sign vector
-$\bs{\delta}{\bs{x}}= (\delta_1(\bs{x}),...,\delta_n(\bs{x}))$. Now
-for each $i$ and some parameter value $\bs{\theta} \in \bs{\Theta} = \R^m$,
-$\delta_i(\bs{\theta})$ tells us whether the corresponding value is under the null
-hypotheses on its border, or under the alternative. Accordingly,
-$\bs{\delta}(\bs{\theta})$ tells us in which combination of true and
-false null hypotheses $H_i$ correspond to $\bs{\theta}$. Let
+{\em I.e.,} $\delta_i$ is takes value $-$ if $H_i$ at $x$ is false and
+$0$ or $+$ otherwise.
+
+Now, for every $\bs{x} \in \mathbb{R}^m$ we define a sign vector
+$\bs{\delta}{\bs{x}}=
+(\delta_1(\bs{x}),...,\delta_n(\bs{x}))$. Accordingly,
+$\bs{\delta}(\bs{x})$ tells us to which 
+combination of true and false null hypotheses some $\bs{x}$
+belongs. Let
 \begin{equation}
 \label{eq:faces}
-\mathcal{F}(\bs{A},\bs{b} = \{\bs{\delta}(\bs{x}): \bs{x} \in \R^m\}
+\mathcal{F}(\mathscr{H}) = \{\bs{\delta}(\bs{x}): \bs{x} \in \R^m\}
 \end{equation}
 
-and for some $F \in \mathcal{F}$ define
+denote the set of all sign patterns defined by 
+$\mathscr{H}$.  Further it
+provides a multiple testing procedure for $\mathscr{H}$ that controls
+the FWER in the strong sense:
+
+\begin{theorem}
+  Consider a family of general contrast hypotheses $\mathscr{H} =
+  \{H_i: \bs{A_i}^t\bs{\mu} \leq b_i,i \in \{1,...,n\}\}$ and let
+  $\mathscr{F}(\mathscr{H}) = \{ \bs{delta}(\bs{x}): \bs{x} \in
+  \R^m\}$ denote the corresponding set of sign patterns. Consider a
+  set of decision rules $\varphi_F: \R^n \rightarrow \{0,1\}$ for
+  which hold $P(\varphi_F = 1 | H_i: F_i \geq 0) \leq \alpha$. Then a
+  corresponding decision rule for each elementary hypotheses $H_i$ is
+  given by 
+  \begin{equation}
+    \label{eq:signtest}
+    \psi_i = \min_{F: F_i \geq 0} \varphi_F,
+  \end{equation}
+  which controls the FWER at level $\alpha$ in the strong sense. 
+\end{theorem}
+
+\begin{proof}
+  For all $F \in \mathcal{F}$ define
   \begin{equation}
     \label{eq:partition}
-  \Theta_F = \{\bs{x} \in \R^m: \bs{\delta}(\bs{x}) = F\}
+    \Theta_F = \{\bs{x} \in \R^m: \bs{\delta}(\bs{x}) = F\},
   \end{equation}
 
-then we have $\bigcup_{F \in \mathcal{F}} \Theta_F = \bs{\Theta} =
-\R^m$ and for any $\Theta_F \cap \Theta_G = \emptyset$ for any $F \neq
-G$. Consequently, $\{\Theta_F:F \in \mathcal{F}$ represents a
-partition of the parameter space and each element in $\mathcal{F}$
-uniquely identifies some {\em non-empty} element of that partition. Consequently,
-following the Partitioning Principle \cite[]{finner2002partitioning}
-if we have level $\alpha$ tests $\varphi_F(X)$ for each $F \in
-\mathcal{F}$ a test that rejects $H_i$ for observed values $X$ if
-$\varphi_F(X) = 1$ for all $F$ with $F_i \geq 1$ controls the FWER in
-the strong sense. For example one may use  weighted Bonferroni tests
+  which is a subset of the parameter space. We have $\bigcup_{F \in
+    \mathcal{F}} \Theta_F = \bs{\Theta} = \R^m$ and for any $\Theta_F
+  \cap \Theta_G = \emptyset$ for any $F \neq G$. Consequently,
+  $\{\Theta_F:F \in \mathcal{F}$ represents a partition of the
+  parameter space and each element in $\mathcal{F}$ uniquely
+  identifies some {\em non-empty} element of that
+  partition. Consequently, following the Partitioning Principle
+  \cite[]{finner2002partitioning} if we have level $\alpha$ tests
+  $\varphi_F(X)$ for each $F \in \mathcal{F}$ a test that rejects
+  $H_i$ for observed values $X$ if $\varphi_F(X) = 1$ for all $F$ with
+  $F_i \geq 1$ controls the FWER in the strong sense.
+\end{proof}
+
+For example one may use  weighted Bonferroni tests
 \begin{equation}
@@
 
-\begin{definition}
-  Consider a family of hypotheses $\mathscr{H} =$
-\end{definition}
-
 \end{document}

### 2013-08-13 presentation/presentation.tex
meaningless changes
Willi added some slides

--- 2013-08-08 presentation/presentation.tex
@@
 \section{Logically dependent hypotheses}
+
+\begin{frame}
+\frametitle{Pairwise comparisons, some history }
+
+Testing hypotheses of no difference in pairwise comparisons of 3 treatments (categories):
+\begin{itemize}
+  \item Parameters: $\mu_1,\mu_2,\mu_3$ ; \\ Hypotheses: $H_1:\mu_1=\mu_2,H_2:\mu_2=\mu_3;H_3:\mu_1=\mu_3$
+  \item A level-$\alpha$ test of $H_ {\{1,2,3\}} = H_1\cap H_2 \cap H_3 : \mu_1=\mu_2=\mu_3$. followed by level-$\alpha$ tests of $H_1,H_2,H_3$ protects the FWER at level $\alpha$.
+  \begin{itemize}
+    \item  essentially closed test; no need to test pairwise intersection hypotheses since $H_{{i,j}} = H_{\{1,2,3\}}, i\neq j$,
+    \item  Bonferroni: if one of the hypotheses is rejected at level $\alpha/3$ the others can be tested at level alpha.
+    \item  Hierarchical: if a predetermined primary hypothesis is rejected at level $\alpha$ the other two can be tested at level $\alpha$.
+    \item  $\Rightarrow$ improves respective Bonferroni-Holm or fixed sequence tests.
+  \end{itemize}
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Pairwise comparisons, some history}
+
+Testing hypotheses of no difference in pairwise comparisons of $m>3$ treatments (categories):\\
+Shaffer (1986) : Modified Sequentially Rejective Multiple test procedures; Example: Modified Bonferroni-Holm
+\begin{itemize}
+  \item "Static":  After j rejections, test remaining hypotheses at level $\alpha/t_j$, $t_j$ = max number of true hypotheses if $j$ are false.
+  \item "Dynamic": After rejecting $H_{i_1},...,H_{i_j}$, test remaining hypotheses at level $\alpha/t^*_j$,  $t^*_j$ = number of true hypotheses if $H_{i_1},...,H_{i_j}$ are false.
+  \item Example: $m=4, n=6$ :
+    \begin{itemize}
+      \item Static: $t_0=6, t_1,t_2,t_3 = 3, t_4=2, t_5=1$
+      \item Dynamic: if, e.g., $H_{\{1,2\}}$ and $H_{\{3,4\}}$ are rejected then only $H_{\{1,3\}}$ and $H_{\{2,3\}}$, or $H_{\{1,4\}}$ and $H_{\{2,3\}}$ can be simultaneously true, hence $t^*_2 = 2 < t_2 =3$ in this case.
+       \end{itemize}
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Example 1: 4 Groups}
+\input{full4ext}
+\end{frame}
 
@@
 
+% \item \textbf{/Some illustration hereabouts/}
+
+% \end{itemize} % ends low level
+% \end{frame}
+
 \begin{frame}

### 2013-08-13 presentation/presentation.tex
set up bibtex for the presentation, entered some slides on contrast tests

@@
 \newcommand{\eg}{{\em e.g.},~}
+\newcommand{\parentd}{../}
+
 \providecommand{\alert}[1]{\textbf{#1}}
@@
 \frametitle{Example 2: Contrast test}
-\begin{itemize}
 
-\item \textbf{/Some illustration hereabouts/}
+\begin{tabular}{ll}
+  \begin{minipage}{.45\textwidth}
+    All pairwise comparisons of 4 means 
+    \begin{displaymath}
+    \bs{C}^t = \left(
+      \begin{array}{llll}
+      -1 & 1 & 0 & 0 \\
+      -1 & 0 & 1 & 0 \\
+      -1 & 0 & 0 & 1 \\
+      0 & -1 & 1 & 0 \\
+      0 & -1 & 0 & 1 \\
+      0 & 0 & -1 & 1 
+    \end{array}
+    \right)
+    \end{displaymath}
+  \end{minipage} &
+  \begin{minipage}{.55\textwidth}
+    Modified Williams Contrasts for Dose Response
+    \begin{displaymath}
+      \bs{C}^t=\left(
+      \begin{array}{llll}
+        1 & 0 & 0 & -1\\
+        1 & 0 &-.51 & -.49 \\
+        1 & -.35 & -.33 & -.32\\
+        .51 & .49 & -.51 & -.48\\
+        .51 & .49 & 0 &  -1\\
+        .35 & .33 & .32& -1
+      \end{array}\right)
+    \end{displaymath}
+  \end{minipage}
+\end{tabular}
 
-\end{itemize} % ends low level
 \end{frame}
@@
 
-\item The flats of the complete graph correspond to the partitions of $\{1,...,m\}$
+\item The flats of the complete graph correspond to the partitions of
+  $\{1,...,m\}$
+\item This provides efficient methods to generate all flats for
+  pairwise mean comparisons \cite{er1988fast,knuth2005art,kokosinski2006new}
 
@@
 \begin{itemize}
+\item \cite{westfall1997multiple} checks each intersection hypotheses
+  whether it is flat using dependence arguments.
+\item \cite{westfall2007multiple} provide a branch and bound algorithm
+  to generate all flats. They note that for pairwise comparisons all
+  exhaustive hypotheses are {\em partition hypotheses} 
+\item \cite{bergmann1988improvements} define the concepts of an
+  exhaustive sets
+\item \cite{bernhard1991computergestuetzte} shows properties of
+  exhaustive sets that are basically {\em closure axioms} of a matroid
+  and provides algorithms to check each intersection hypotheses if
+  flat using rank function.
 
+  \cite{hommel1999bonferroni} show {\em closure axioms} for exhaustive
+  sets 
 \item Flats <-> Weakly Exhaustive set (Hommel, Bergmann)
@@
 
+\bibliographystyle{alpha}
+\bibliography{\parentd paper/cleanrefs}
 \end{document}

### 2013-08-14 presentation/presentation.tex
added a slide to illustrate set partitions and bell numbers

@@
 
+\begin{tabular}{ll}
+  \begin{minipage}{.3\textwidth}
+    \includegraphics[width=1.1\textwidth]{setpartitions.pdf} \\
+     {\tiny wikipedia:Bell\_number}
+  \end{minipage} &
+  \begin{minipage}{.7\textwidth}
   \begin{itemize}
@@
   \item Provides an efficient way to generate all exhaustive sets by
-    enumeration of all partitions [ER '88] or hyperplanes 
-  \item The number of exhaustive sets are given by the Bell numbers
-
+    enumeration of all partitions or hyperplanes \cite{er1988fast,knuth2005art,kokosinski2006new} 
+  \item The number of flats is given by the Bell-numbers \ie 5, 15,
+    52, 203, 877, 4140, ... 
   \end{itemize}
+  \end{minipage}
+\end{tabular}
 

### 2013-08-15 paper/draft.tex
Added proof for hierarchical result

@@
 \begin{theorem}
+  \label{the:strictly.exhaustive}
   Consider a family of null hypotheses for which conditions
@@
 
-\begin{theorem}
-  Consider a hierarchical testing procedure that tests $H_{i_r}$ at
-  level $\alpha$ according to a prefixed order $(i_1,...,i_n)$. Assume
-  that hypotheses with indeces $i_1$ through $i_k$ have been
-  rejected. If for $k+1$ to $k+l$ there is $i_s$, $s \leq k$ such that
+% \begin{theorem}
+%   Consider a hierarchical testing procedure that tests $H_{i_r}$ at
+%   level $\alpha$ according to a prefixed order $(i_1,...,i_n)$. Assume
+%   that hypotheses with indeces $i_1$ through $i_k$ have been
+%   rejected. If for $k+1$ to $k+l$ there is $i_s$, $s \leq k$ such that
+%   \begin{equation}
+%     \label{eq:condh}
+%     \rho \{i_s,i_{k+1}\} = \rho \{i_s,i_{k+1}...,i_{k+l}\},
+%   \end{equation}
+%   hypotheses each $H_{i_{k+1}}$ through   $H_{i_{k+l}}$ can be tested
+%   simultaneously at full level $\alpha$. 
+% \end{theorem}
+
+% \begin{proof}
+%   Due to \ref{eq:condh} for any $k+1 \leq m' < m \leq k+l$, $i_m \in
+%   \sigma \{i_s,i_{m'}\}$ and consequently also $i_s \in \sigma
+%   \{i_m,i_{m'}\}$. Therefore $i_s \in \sigma J$ for all $J$ with
+%   $i_m,i_{m'} \in J$. Since $p_s \leq \alpha$ all intersection
+%   hypotheses corresponding to exhaustive sets containing $i_{m}$
+%   and $i_{m'}$ for all $k+1 \leq m' < m$ have been rejected. Such that
+%   $p_{i_{m}} \leq \alpha$ rejects $H_{i_{m}}$.
+% \end{proof}
+
+\begin{theorem}
+  Consider a system of logically dependent hypotheses $\mathscr{H}$
+  with dependence relation $\sim$ and associated matroid
+  $M(\mathscr{H})$. Assume that $M(\mathscr{H})$ neither contains
+  loops nor parallels. Consider a hierarchical testing procedure that
+  tests $H_{i}$ at level $\alpha$ according to a prefixed
+  order. Without loss of generality assume that $H_i$ are indexed
+  according to the fixed rejection sequence $(1,...,n)$. Given 
+  that hypotheses with indeces $1$ through $r$ have been
+  rejected and if for $r+1$ to $r+s$ holds
   \begin{equation}
-    \label{eq:condh}
-    \rho \{i_s,i_{k+1}\} = \rho \{i_s,i_{k+1}...,i_{k+l}\},
+    \label{eq:cond.hierarchical}
+    \rho \{1,...,r\} + 1 = \rho \{1,...,r+1\} = \rho \{1,...,r+s\},
   \end{equation}
-  hypotheses each $H_{i_{k+1}}$ through   $H_{i_{k+l}}$ can be tested
-  simultaneously at full level $\alpha$. 
+  hypotheses $H_{r+1}$ through $H_{r+s}$ can be tested simultaneously
+  at full level $\alpha$.
 \end{theorem}
@@
 \begin{proof}
-  Due to \ref{eq:condh} for any $k+1 \leq m' < m \leq k+l$, $i_m \in
-  \sigma \{i_s,i_{m'}\}$ and consequently also $i_s \in \sigma
-  \{i_m,i_{m'}\}$. Therefore $i_s \in \sigma J$ for all $J$ with
-  $i_m,i_{m'} \in J$. Since $p_s \leq \alpha$ all intersection
-  hypotheses corresponding to exhaustive sets containing $i_{m}$
-  and $i_{m'}$ for all $k+1 \leq m' < m$ have been rejected. Such that
-  $p_{i_{m}} \leq \alpha$ rejects $H_{i_{m}}$.
+  We show by that given $p_{r+k} < \alpha$ all $E \in \mathscr{L}_{r+k} =
+  \{I \in \mathscr{L}: r+k \in I\}$ are rejected by the stepwise
+  procedure. First, notice that for all $E$ such that $\min E = r+k$
+  rejection follows immediately and if $\min E \leq r$ rejection
+  follows by rejection of $H_i$ for $i \leq r$. It remains to be shown
+  that the stepwise procedure also rejects $H_E$ for $r < \min E <
+  r+k$.
+
+  First we show by contradiction that some $k \leq t$ we have $i \in
+  \sigma \{r+l,r+k\}$ for some $i \leq r$ and all $1 \leq l < k$.  We
+  need to show that for all $s,t$, $k+1 < s < t \leq k+l$, follows
+  $i_s \sim K \cup i_t$:
+
+  Assume that there is an $l$ such that there is on $i \leq r$ with $i
+  \in \sigma \{r+l,r+k\}$. Consequently for all $i \leq r$ $i \nsim
+  \{r+l,r+k\}$. Since $M(\mathscr{H})$ has no loops or parallels also
+  $r+l \nsim r+k$. Consequently, for any maximally independent set in
+  $I \subset \{1,...,r\}$, $I \cup \{r+l, r+k\}$ is independent and
+  has cardinality $\rho\{1,...,r\}+2$ which contradicts
+  (\ref{eq:cond.hierarchical}).
+
+  Above argument can be repeated for any set $J \cup r+k$ where $J
+  \subseteq \{r+1,...,r+k-1\}$. Therefore, we have $i \in E$, $i \leq
+  r$ for all exhaustive set containing $r+k$ with $r < \min E <
+  r+k$. Rejection of these sets follows from rejection of $H_i$, $i
+  \leq r$, which completes the proof. 
 \end{proof}
+
+Note that the condition that $M(\mathscr{H})$ contains neither loops
+nor parallels is not a serious restriction. A loop in terms of systems
+of hypotheses would be a set in the parameter space that contains all
+hypotheses. Thich corresponds to the alternative that at least one
+null hypotheses is true which is the domain of intersection-union
+tests. The condition that $M(\mathscr{H})$ does not contain parallels
+is more restrictive, however, considering that the absence of
+parallels is a sufficient condition such that the correspondence
+between flats of the matroid and strictly exhaustive sets (Theorem
+\ref{the:strictly.exhaustive}) this is also not serious. Once we turn
+to one-sided problems where we will use oriented matroids to describe
+the structure of systems of null hypotheses these problems will seem
+even less problematic.
 

### 2013-08-15 paper/draft.tex
clarified proof

@@
 \begin{proof}
-  We show by that given $p_{r+k} < \alpha$ all $E \in \mathscr{L}_{r+k} =
-  \{I \in \mathscr{L}: r+k \in I\}$ are rejected by the stepwise
-  procedure. First, notice that for all $E$ such that $\min E = r+k$
-  rejection follows immediately and if $\min E \leq r$ rejection
-  follows by rejection of $H_i$ for $i \leq r$. It remains to be shown
-  that the stepwise procedure also rejects $H_E$ for $r < \min E <
-  r+k$.
-
-  First we show by contradiction that some $k \leq t$ we have $i \in
-  \sigma \{r+l,r+k\}$ for some $i \leq r$ and all $1 \leq l < k$.  We
-  need to show that for all $s,t$, $k+1 < s < t \leq k+l$, follows
-  $i_s \sim K \cup i_t$:
-
-  Assume that there is an $l$ such that there is on $i \leq r$ with $i
-  \in \sigma \{r+l,r+k\}$. Consequently for all $i \leq r$ $i \nsim
+  We show that given $p_{r+k} \leq \alpha$ for some $1 \leq k \leq s$
+  all exhaustive sets $E \in \mathscr{L}_{r+k} = \{E \in \mathscr{L}:
+  r+k \in E\}$ are rejected by the stepwise procedure. It remains to
+  be shown that the stepwise procedure also rejects $H_E$ for all
+  exhaustive sets with $\min E < r+k$. Note that $\{E \in \mathscr{L}:
+  \min E < r+k\} = \overline{\underline{E}}_1^r \cup
+  \overline{\underline{E}}^{r+k-1}_{r+1}$ where
+  $\overline{\underline{E}}^b_a := \{\sigma I: \text{for all } I
+  \subseteq S, a \leq \min I \leq b\}$. For $ E \in
+  \overline{\underline{E}}_1^r$ follows since $p_i \leq \alpha$ for $i
+  \leq r$.
+
+  We show by contradiction that we have $i \in \sigma \{r+l,r+k\}$ for
+  some $i \leq r$ and all $1 \leq l < k$.  Assume that there is an $l$
+  such that there is no $i \leq r$ with $i \in \sigma
+  \{r+l,r+k\}$. Consequently for all $i \leq r$, we have $i \nsim
   \{r+l,r+k\}$. Since $M(\mathscr{H})$ has no loops or parallels also
-  $r+l \nsim r+k$. Consequently, for any maximally independent set in
+  $(r+l) \nsim (r+k)$. Consequently, for any maximally independent set
   $I \subset \{1,...,r\}$, $I \cup \{r+l, r+k\}$ is independent and
@@
 
-  Above argument can be repeated for any set $J \cup r+k$ where $J
-  \subseteq \{r+1,...,r+k-1\}$. Therefore, we have $i \in E$, $i \leq
-  r$ for all exhaustive set containing $r+k$ with $r < \min E <
-  r+k$. Rejection of these sets follows from rejection of $H_i$, $i
-  \leq r$, which completes the proof. 
+  Above argument can be repeated for any set $I$ with $r+1 \leq \min I
+  < k$. Therefore, we have $i \in E$ for all $E \in
+  \overline{\underline{E}}^{r+k-1}_{r+1}$. Rejection of these sets
+  follows from rejection of $H_i$, $i \leq r$, which completes the
+  proof. 
 \end{proof}

### 2013-08-15 paper/draft.tex
fixed theorem after yet another false proof

@@
     \label{eq:cond.hierarchical}
-    \rho \{1,...,r\} + 1 = \rho \{1,...,r+1\} = \rho \{1,...,r+s\},
+    \forall 1 \leq l < k \leq s: \; \exists i \leq r: i \sim \{l,k\},
   \end{equation}
@@
   all exhaustive sets $E \in \mathscr{L}_{r+k} = \{E \in \mathscr{L}:
-  r+k \in E\}$ are rejected by the stepwise procedure. It remains to
-  be shown that the stepwise procedure also rejects $H_E$ for all
-  exhaustive sets with $\min E < r+k$. Note that $\{E \in \mathscr{L}:
-  \min E < r+k\} = \overline{\underline{E}}_1^r \cup
+  r+k \in E\}$ are rejected by the stepwise procedure. Since $p_{r+k}
+  \leq \alpha$ implies rejection of all $E$ such that $\min E \geq
+  r+k$, it remains to be shown that the stepwise procedure rejects
+  $H_E$ for all exhaustive sets with $\min E < r+k$. Note that $\{E
+  \in \mathscr{L}: \min E < r+k\} = \overline{\underline{E}}_1^r \cup
   \overline{\underline{E}}^{r+k-1}_{r+1}$ where
@@
 
-  We show by contradiction that we have $i \in \sigma \{r+l,r+k\}$ for
-  some $i \leq r$ and all $1 \leq l < k$.  Assume that there is an $l$
-  such that there is no $i \leq r$ with $i \in \sigma
-  \{r+l,r+k\}$. Consequently for all $i \leq r$, we have $i \nsim
-  \{r+l,r+k\}$. Since $M(\mathscr{H})$ has no loops or parallels also
-  $(r+l) \nsim (r+k)$. Consequently, for any maximally independent set
-  $I \subset \{1,...,r\}$, $I \cup \{r+l, r+k\}$ is independent and
-  has cardinality $\rho\{1,...,r\}+2$ which contradicts
-  (\ref{eq:cond.hierarchical}).
-
-  Above argument can be repeated for any set $I$ with $r+1 \leq \min I
-  < k$. Therefore, we have $i \in E$ for all $E \in
-  \overline{\underline{E}}^{r+k-1}_{r+1}$. Rejection of these sets
+  Consider any set $I$ with $r < \min I < k$ then there is an $i \sim
+  \{\min I,k\}$ with $i \leq r$. Consequently, $i \in \sigma \{\min
+  I,k\}$ and also $\sigma I$. Therefore, we have $i \in E$ for all $E
+  \in \overline{\underline{E}}^{r+k-1}_{r+1}$. Rejection of these sets
   follows from rejection of $H_i$, $i \leq r$, which completes the

### 2013-08-15 presentation/presentation.tex
Illustrated and fixed hierarchical theorem

@@
 \providecommand{\alert}[1]{\textbf{#1}}
+\usetheme{Novartis}
+\doNotUseNovartisLogoOnAllSlides
+\doNotUseNovartisLogoOnTitlePage
 
 \title{Matroids from families of null hypotheses}
-\author{Willi Maurer, Florian Klinglmueller}
+\author{Florian Klinglmueller, Willi Maurer}
 \date{\today}
@@
 Testing hypotheses of no difference in pairwise comparisons of $m>3$ treatments (categories):\\
-Shaffer (1986) : Modified Sequentially Rejective Multiple test procedures; Example: Modified Bonferroni-Holm
+\cite{shaffer1986modified}: Modified Sequentially Rejective Multiple test procedures; Example: Modified Bonferroni-Holm
 \begin{itemize}
@@
 
+\begin{tabular}{ll}
+  \begin{minipage}{.4\textwidth}
+    \input{cycle}
+  \end{minipage} &
+  \begin{minipage}{.6\textwidth}
     Consider $S$ and a collection of subsets $\mathscr{C}$ for which holds
@@
     then $\mathscr{C}$ is the set of circuits of a matroid $M$ on $S$
-
-\input{cycle}
+  \end{minipage}
+\end{tabular}
 
@@
 \frametitle{Independence axioms}
-\begin{block}{Independence}
+
+\begin{tabular}{ll}
+  \begin{minipage}{.4\textwidth}
+  \only<1>{\input{independent}}
+  \only<2>{\input{cycle}}
+  \invisible<1>{\alert{\footnotesize minimally dependent set!}}
+  \end{minipage} &
+  \begin{minipage}{.6\textwidth}
   Consider $S$ and a collection of subsets $\mathscr{I}$ for which
@@
   $I\cap i \in \mathscr{I}$.
-\end{block}
 
-\begin{block}{Circuit}
-  $C \subset S$ is a circuit iff $C \neq \mathscr{I}$ and for all $I
-  \subset C$, $I \in \mathscr{I}$ (\ie minimally dependent set)
-\end{block}
+  \end{minipage}
+\end{tabular}
+
 \end{frame}
@@
 
-\begin{block}{Dependence}
-  Consider $S$ and a relation $\sim_M$ on elements and subsets of $S$,
-  then if for any $x,y_1,...,y_n,z_1,...,z_m$ holds
+\begin{tabular}{ll}
+  \begin{minipage}{.4\textwidth}
+    \input{dependent}
+  \end{minipage} &
+  \begin{minipage}{.6\textwidth}
+  Relation $\sim_M$ on elements and subsets of $S$,
+  if for any $x \in S$, and sets $Y,Z \subseteq S$ holds
   \begin{enumerate}
-  \item $y_i \sim \{y_1,...,y_n\}$ for all $i \leq n$
-  \item if $n\geq 1$, $x \sim_M \{y_1,...,y_n\}$ and $x \nsim_M
-    \{y_1,...,y_{n-1}$ then $y_n \sim \{y_1,...,y_{n-1},x\}$
-  \item if $x \sim_M \{y_1,...,y_n\}$ and for all $i \leq n$ $y_i
-    \sim_M \{z_1,...,z_m\}$ then $x \sim_M \{z_1,...,z_m\}$
+  \item $y \sim Y$ for all $y \in Y$
+  \item if $x \sim_M Y$ and $x \nsim_M Y\setminus y$
+    then $y \sim x \cup (Y \setminus y)$
+  \item if $x \sim_M Y$ and for all $y \in Y$, $y \sim_M Z$ then $x \sim_M Z$
   \end{enumerate}
-  then $\sim_M$ defines a dependence relation of a matroid on $S$.
-\end{block}
+  then $\sim_M$ defines a matroid on $S$.
+  \end{minipage}
+\end{tabular}
 
@@
 \frametitle{Reminder: Graphical approaches}
+
+\begin{tabular}{ll}
+  \begin{minipage}{.4\textwidth}
+    \begin{tikzpicture}[remember picture,overlay]
+      \node[yshift=-2.5cm,xshift=-1.4cm] at (current page.north west){\input{qvatest}};
+    \end{tikzpicture}
+
+  \end{minipage} &
+  \begin{minipage}{.6\textwidth}
+    \begin{itemize}
+    \item Provides weights $w_{i,I}$ for each elementary hypotheses
+      $H_i$ and each non-empty subset $I \subset S$
+    \item Weights are used to define a closed test procedure of
+      weighted intersection hypotheses tests
+    \item We have $\sum_{i \in S} w_{i,I} \leq 1$ (conservative),
+    \item and $w_{i,I} \leq w_{i,J}$ for $J \subseteq I$ (consonance).
+    \end{itemize}
+  \end{minipage}
+\end{tabular}
 \end{frame}
@@
 \begin{frame}
-\frametitle{Hierarchical tests}
+  \frametitle{Example: Hierarchical LSD}
+\begin{tabular}{ll}
+  \begin{minipage}{.4\textwidth}
+    \input{lsd2.tex}
+  \end{minipage} &
+  \begin{minipage}{.6\textwidth}
+    \begin{itemize}
+    \item 3 Hypotheses:
+      \begin{enumerate}
+      \item Placebo vs. High Dose
+      \item Placebo vs. Low Dose
+      \item High Dose vs. Low Dose
+      \end{enumerate}
+    \item Prefixed sequence: $H_1 \rightarrow H_2 \rightarrow H_3$
+    \item Test $H_1$ at level $\alpha$
+    \item If we reject $H_1$ test $H_2$ and $H_3$ at level $\alpha$
+      simultaneously
+    \item Can we generalize this?
+    \end{itemize}
+  \end{minipage}
+\end{tabular}
+
+\end{frame}
+
+\begin{frame}
+  \frametitle{Fixed sequence test of pairwise mean comparisons}
+\begin{itemize}
+\item {\bf Picture of hierarchical test hereabouts}
+\end{itemize}
+
+\begin{block}{Theorem}
+  Assume that $M(\mathscr{H})$ neither contains loops nor
+  parallels. Assume that hypotheses $H_1$ through $H_r$ have been
+  rejected. Consider that hypotheses $H_{r+1}$ through $H_{r+t}$ are to
+  be tested using at level $\alpha_r$ according to a prefixed order.
+  If for some $1 \leq s \leq t$ holds
+  \begin{equation}
+    \label{eq:cond.hierarchical}
+    \forall 1 \leq l < k \leq s: \; \exists i \leq r: i \sim \{l,k\},
+  \end{equation}
+  hypotheses $H_{r+1}$ through $H_{r+s}$ can be tested simultaneously
+  at full level $\alpha_ra$.
+\end{block}
+
+\end{frame}
+
+\begin{frame}
+  \frametitle{Example: Multiple doses}
+
+\begin{tabular}{ll}
+  \begin{minipage}{.4\textwidth}
+    \begin{tikzpicture}[overlay]
+      \node[yshift=70bp,xshift=-40bp] at (current page.south west){\input{lsd4.tex}};
+    \end{tikzpicture}
+  \end{minipage} &
+  \begin{minipage}{.6\textwidth}
+    \begin{tikzpicture}[remember picture,overlay]
+      \node[yshift=100bp,xshift=160bp] at (current page.north west){\input{lsd4t.tex}};
+    \end{tikzpicture}
+  \end{minipage}
+\end{tabular}
 \end{frame}

### 2013-08-15 presentation/full4ext.tex
added a couple of graph files, set up the pictures for qva example

--- 2013-07-31 exportfunctions/base4.tex
@@
-\documentclass[11pt]{article}
-\usepackage{tikz}
-\usetikzlibrary{decorations,arrows,shapes}
-\begin{document}\begin{tikzpicture}[scale=1]
+\begin{tikzpicture}[scale=.5]
 
@@
 \draw [line width=1pt] (H1) to (325bp, -124bp) node[fill=blue!20] {$H_{1}$} to (H2);
+\draw [line width=1pt] (H1) to (275bp, -175bp) node[fill=blue!20] {$H_{2}$} to (H3);
 \draw [line width=1pt] (H2) to (425bp, -224bp) node[fill=blue!20] {$H_{4}$} to (H3);
+\draw [line width=1pt] (H3) to (326bp, -324bp) node[fill=blue!20] {$H_{6}$} to (H4);
 \draw [line width=1pt] (H4) to (225bp, -227bp) node[fill=blue!20] {$H_{3}$} to (H1);
+\draw [line width=1pt] (H4) to (374bp, -176bp) node[fill=blue!20] {$H_{5}$} to (H2);
+
+\node (H1) at (525bp,-125bp)[draw,circle,fill=white] {$\mu_1$};
+\node (H2) at (725bp,-125bp)[draw,circle,fill=white] {$\mu_2$};
+\node (H3) at (725bp,-325bp)[draw,circle,fill=white] {$\mu_3$};
+\node (H4) at (525bp,-325bp)[draw,circle,fill=white] {$\mu_4$};
+\draw [line width=1pt] (H1) to (625bp, -124bp) node[fill=red!20] {$H_{1}$} to (H2);
+\draw [line width=1pt] (H1) to (575bp, -175bp) node[fill=blue!20] {$H_{2}$} to (H3);
+\draw [line width=1pt] (H2) to (725bp, -224bp) node[fill=blue!20] {$H_{4}$} to (H3);
+\draw [line width=1pt] (H3) to (626bp, -324bp) node[fill=red!20] {$H_{6}$} to (H4);
+\draw [line width=1pt] (H4) to (525bp, -227bp) node[fill=blue!20] {$H_{3}$} to (H1);
+\draw [line width=1pt] (H4) to (674bp, -176bp) node[fill=blue!20] {$H_{5}$} to (H2);
+
 \end{tikzpicture}
-\end{document}

### 2013-08-15 presentation/lsd2.tex
added a couple of graph files, set up the pictures for qva example

\begin{tikzpicture}[scale=.5]
\node (Pl) at (25bp,-225bp)[draw,circle,fill=white] {$Pl$};
\node (Dl) at (175bp,-175bp)[draw,circle,fill=white] {$Dl$};
\node (Dh) at (225bp,-25bp)[draw,circle,fill=white] {$Dh$};
\draw [line width=1pt] (Pl.363) arc(273:290:209bp) node[fill=blue!20] {$H_{2}$} arc(290:302:209bp) to (Dl);
\draw [line width=1pt] (Pl.70) arc(160:145:282bp) node[fill=blue!20] {$H_{1}$} arc(145:112:282bp) to (Dh);
\draw [line width=1pt] (Dl.416) arc(326:340:212bp) node[fill=blue!20] {$H_{3}$} arc(340:355:212bp) to (Dh);
\end{tikzpicture}

### 2013-08-15 presentation/lsd4.tex
added a couple of graph files, set up the pictures for qva example

\begin{tikzpicture}[transform canvas={scale=.8},scale=.6]

\node (H1) at (75bp,-225bp)[draw,circle,fill=white] {$H1$};
\node (D1) at (225bp,-375bp)[draw,circle,fill=white] {$D1$};
\node (D2) at (375bp,-325bp)[draw,circle,fill=white] {$D2$};
\node (D3) at (375bp,-125bp)[draw,circle,fill=white] {$D3$};
\node (D4) at (225bp,-75bp)[draw,circle,fill=white] {$D4$};
\draw [line width=1pt] (H1) to (112bp, -262bp) node[fill=blue!20] {$H_{5}$} to (D1);
\draw [line width=1pt] (H1) to (150bp, -250bp) node[fill=blue!20] {$H_{3}$} to (D2);
\draw [line width=1pt] (H1) to (150bp, -200bp) node[fill=blue!20] {$H_{2}$} to (D3);
\draw [line width=1pt] (H1) to (112bp, -188bp) node[fill=blue!20] {$H_{1}$} to (D4);
\draw [line width=1pt] (D2) to (338bp, -337bp) node[fill=blue!20] {$H_{9}$} to (D1);
\draw [line width=1pt] (D3) to (338bp, -187bp) node[fill=blue!20] {$H_{7}$} to (D1);
\draw [line width=1pt] (D3) to (375bp, -175bp) node[fill=blue!20] {$H_{10}$} to (D2);
\draw [line width=1pt] (D4) to (225bp, -150bp) node[fill=blue!20] {$H_{6}$} to (D1);
\draw [line width=1pt] (D4) to (262bp, -137bp) node[fill=blue!20] {$H_{8}$} to (D2);
\draw [line width=1pt] (D4) to (262bp, -87bp) node[fill=blue!20] {$H_{4}$} to (D3);
\end{tikzpicture}

### 2013-08-15 presentation/lsd4t.tex
added a couple of graph files, set up the pictures for qva example

\begin{tikzpicture}[transform canvas={scale=.5}]

\node (H1) at (50bp,-350bp)[draw,circle split,fill=green!80] {$H1$ \nodepart{lower} $0$};
\node (H2) at (141bp,-350bp)[draw,circle split,fill=green!80] {$H2$ \nodepart{lower} $0$};
\node (H3) at (234bp,-350bp)[draw,circle split,fill=green!80] {$H3$ \nodepart{lower} $0$};
\node (H4) at (325bp,-350bp)[draw,circle split,fill=green!80] {$H4$ \nodepart{lower} $0$};
\node (H5) at (125bp,-475bp)[draw,circle split,fill=green!80] {$H5$ \nodepart{lower} $0$};
\node (H6) at (175bp,-475bp)[draw,circle split,fill=green!80] {$H6$ \nodepart{lower} $0$};
\node (H7) at (225bp,-475bp)[draw,circle split,fill=green!80] {$H7$ \nodepart{lower} $0$};
\node (H8) at (175bp,-625bp)[draw,circle split,fill=green!80] {$H8$ \nodepart{lower} $0$};
\node (H9) at (125bp,-625bp)[draw,circle split,fill=green!80] {$H9$ \nodepart{lower} $0$};
\node (H10) at (225bp,-625bp)[draw,circle split,fill=green!80] {$H10$ \nodepart{lower} $0$};

\draw [line width=1pt] (0bp,-315bp) -- (370bp,-315bp) -- (370bp,-385bp) -- (0bp,-385bp) -- (0bp,-315bp);
\node[anchor=north east] (text) at (370bp,-385bp) {reject all!};
\draw [line width=1pt] (80bp,-440bp) -- (270bp,-440bp) -- (270bp,-510bp) -- (80bp,-510bp) -- (80bp,-440bp);
\node[anchor=north east] (text) at (270bp,-510bp) {reject all!};
\draw [->,line width=1pt] (H1) to (96bp, -350bp) node[fill=blue!20] {$1$} to (H2);
\draw [->,line width=1pt] (H2) to (187bp, -350bp) node[fill=blue!20] {$1$} to (H3);
\draw [->,line width=1pt] (H3) to (278bp, -350bp) node[fill=blue!20] {$1$} to (H4);
\draw [->,line width=1pt] (175bp,-385bp) to (157bp, -420bp) node[fill=blue!20] {$1$} to (H5);
\draw [->,line width=1pt] (175bp,-385bp) to (175bp, -420bp) node[fill=blue!20] {$1$} to (H6);
\draw [->,line width=1pt] (175bp,-385bp) to (193bp, -420bp) node[fill=blue!20] {$1$} to (H7);
\draw [->,line width=1pt] (175bp,-510bp) to (175bp, -550bp) node[fill=blue!20] {$1$} to (H8);
\draw [->,line width=1pt] (175bp,-510bp) to (157bp, -550bp) node[fill=blue!20] {$1$} to (H9);
\draw [->,line width=1pt] (175bp,-510bp) to (193bp, -550bp) node[fill=blue!20] {$1$} to (H10);
\end{tikzpicture}

### 2013-08-15 presentation/qvameans.tex
added a couple of graph files, set up the pictures for qva example

\begin{tikzpicture}[transform canvas={scale=0.6},scale=.7]

\node (QVA) at (75bp,-275bp)[draw,circle,fill=white] {$QVA$};
\node (QAB) at (225bp,-225bp)[draw,circle,fill=white] {$QAB$};
\node (NVA) at (375bp,-275bp)[draw,circle,fill=white] {$NVA$};
\node (Pl) at (225bp,-75bp)[draw,circle,fill=white] {$Pl$};
\node (Tio) at (225bp,-375bp)[draw,circle,fill=white] {$Tio$};
\node (SE) at (75bp,-425bp)[draw,circle,fill=white] {$SE$};
\draw [line width=1pt] (QVA) to (112bp, -263bp) node[fill=blue!20] {$H_{4}$} to (QAB);
\draw [line width=1pt] (QVA) to (150bp, -275bp) node[fill=blue!20] {$H_{5}$} to (NVA);
\draw [line width=1pt] (QVA) to (112bp, -300bp) node[fill=blue!20] {$H_{6}$} to (Tio);
\draw [line width=1pt] (QVA) to (75bp, -350bp) node[fill=blue!20] {$H_{7}$} to (SE);
\draw [line width=1pt] (Pl) to (188bp, -125bp) node[fill=blue!20] {$H_{1}$} to (QVA);
\draw [line width=1pt] (Pl) to (225bp, -150bp) node[fill=blue!20] {$H_{2}$} to (QAB);
\draw [line width=1pt] (Pl) to (262bp, -125bp) node[fill=blue!20] {$H_{3}$} to (NVA);
\end{tikzpicture}

### 2013-08-15 presentation/presentation.tex
added a couple of graph files, set up the pictures for qva example
fixed typo in hierarchical theorem

@@
   parallels. Assume that hypotheses $H_1$ through $H_r$ have been
-  rejected. Consider that hypotheses $H_{r+1}$ through $H_{r+t}$ are to
-  be tested using at level $\alpha_r$ according to a prefixed order.
-  If for some $1 \leq s \leq t$ holds
+  rejected. Consider that hypotheses $H_{r+1}$ through $H_{r+t}$ are
+  to be tested at level $\alpha_r$ according to a prefixed order.  If
+  for some $1 \leq s \leq t$ holds
   \begin{equation}
     \label{eq:cond.hierarchical}
-    \forall 1 \leq l < k \leq s: \; \exists i \leq r: i \sim \{l,k\},
+    \forall 1 \leq l < k \leq s: \; \exists i \leq r: i \sim \{(r+l),(r+k)\},
   \end{equation}
@@
       \node[yshift=100bp,xshift=160bp] at (current page.north west){\input{lsd4t.tex}};
+    \end{tikzpicture}
+  \end{minipage}
+\end{tabular}
+\end{frame}
+
+\begin{frame}
+  \frametitle{General weighting strategies}
+  \framesubtitle{QVA Example}
+
+\begin{tabular}{ll}
+  \begin{minipage}{.4\textwidth}
+    \begin{tikzpicture}[remember picture,overlay]
+      \node[yshift=-2.5cm,xshift=-1.4cm] at (current page.north west){\input{qvatest}};
+    \end{tikzpicture}
+
+  \end{minipage} &
+  \begin{minipage}{.6\textwidth}
+    \begin{tikzpicture}[remember picture,overlay]
+      \node[yshift=-50bp,xshift=-180bp] at (current page.north east){\input{qvameans}};
   \end{tikzpicture}

### 2013-08-15 paper/draft.tex

@@
 \documentclass[a4paper,12pt]{article}
+\usepackage{amsthm}
 \usepackage{algorithmic}
@@
 \usepackage{color}
-\newenvironment{proof}[1][Proof]{\noindent\textbf{#1:} }{}
+
+%\newenvironment{proof}[1][Proof]{\noindent\textbf{#1:} }{}
 \newcommand{\bs}[1]{\boldsymbol{#1}}
@@
 \begin{proposition}{Properties of $M$ (Welsh, 2010)}
-\label{the:cyclemat}
+  The following properties hold for the cycle matroid of a graph $G$:
   \begin{enumerate}
@@
   \end{enumerate}
+\label{the:cyclemat}
 \end{proposition}
@@
 
-Accordingly all porperties given in \ref{the:cyclemat} also hold for
+Accordingly all porperties given in Proposition \ref{the:cyclemat} also hold for
 the matroid on exhaustive sets.
@@
   with dependence relation $\sim$ and associated matroid
-  $M(\mathscr{H})$. Assume that $M(\mathscr{H})$ neither contains
-  loops nor parallels. Consider a hierarchical testing procedure that
-  tests $H_{i}$ at level $\alpha$ according to a prefixed
-  order. Without loss of generality assume that $H_i$ are indexed
-  according to the fixed rejection sequence $(1,...,n)$. Given 
-  that hypotheses with indeces $1$ through $r$ have been
-  rejected and if for $r+1$ to $r+s$ holds
+  $M(\mathscr{H})$.   Assume that $M(\mathscr{H})$ neither contains loops nor
+  parallels. Assume that hypotheses $H_1$ through $H_r$ have been
+  rejected. Consider that hypotheses $H_{r+1}$ through $H_{r+t}$ are to
+  be tested at level $\alpha$ according to a prefixed order.
+  If for some $1 \leq s \leq t$ holds
   \begin{equation}
     \label{eq:cond.hierarchical}
-    \forall 1 \leq l < k \leq s: \; \exists i \leq r: i \sim \{l,k\},
+    \forall 1 \leq l < k \leq s: \; \exists i \leq r: i \sim \{(r+l),(r+k)\},
   \end{equation}
@@
   at full level $\alpha$.
+  \begin{equation}
+    \label{eq:cond.hierarchical}
+    \forall 1 \leq l < k \leq s: \; \exists i \leq r: i \sim \{r+l,k\},
+  \end{equation}
+  hypotheses $H_{r+1}$ through $H_{r+s}$ can be tested simultaneously
+  at full level $\alpha$.
 \end{theorem}
@@
   all exhaustive sets $E \in \mathscr{L}_{r+k} = \{E \in \mathscr{L}:
-  r+k \in E\}$ are rejected by the stepwise procedure. Since $p_{r+k}
-  \leq \alpha$ implies rejection of all $E$ such that $\min E \geq
-  r+k$, it remains to be shown that the stepwise procedure rejects
-  $H_E$ for all exhaustive sets with $\min E < r+k$. Note that $\{E
-  \in \mathscr{L}: \min E < r+k\} = \overline{\underline{E}}_1^r \cup
-  \overline{\underline{E}}^{r+k-1}_{r+1}$ where
-  $\overline{\underline{E}}^b_a := \{\sigma I: \text{for all } I
-  \subseteq S, a \leq \min I \leq b\}$. For $ E \in
-  \overline{\underline{E}}_1^r$ follows since $p_i \leq \alpha$ for $i
-  \leq r$.
-
-  Consider any set $I$ with $r < \min I < k$ then there is an $i \sim
-  \{\min I,k\}$ with $i \leq r$. Consequently, $i \in \sigma \{\min
-  I,k\}$ and also $\sigma I$. Therefore, we have $i \in E$ for all $E
-  \in \overline{\underline{E}}^{r+k-1}_{r+1}$. Rejection of these sets
-  follows from rejection of $H_i$, $i \leq r$, which completes the
-  proof.
+  r+k \in E\}$ are rejected by the stepwise procedure. We have
+  $\mathscr{L}_{r+k} = \{\sigma I:I \subseteq S, (r+k) \in
+  I\}$. $p_{r+k} \leq \alpha$ implies rejection of all $I \subseteq S$ such that
+  $\min I = r+k$. It remains to be shown that the stepwise procedure
+  rejects $H_E$ for all exhaustive sets with $\min E < r+k$. We have 
+  $\{E \in \mathscr{L}_{r+k}: \min E < r+k\} = \{\sigma I: I \subseteq
+  S, \min I \leq r\} \cup \{\sigma I: I \subseteq S, r < \min I <
+  r+k\}$. For all $E \in \{\sigma I: I \subseteq
+  S, \min I \leq r\}$ rejection of $H_E$ follows from rejection of
+  $H_i$, $i \leq r$.
+
+  Consider any set $I$ with $r < \min I < r+k$ then there is an $i
+  \leq r$ such that $i \sim \{\min I,k\}$. Consequently, $i \in \sigma
+  \{\min I,k\} \subseteq \sigma I$. Therefore, we have $i \in E$ for
+  all $E \in \{\sigma I: I \subseteq S, r < \min I < r+k\}$. Rejection
+  of the corresponding intersection hypotheses follows from rejection
+  of $H_i$, $i \leq r$, which completes the proof.
 \end{proof}

### 2013-08-16 presentation/cycle.tex
included first example graphs, changed scaling of graphs to .5
added cycles, rearranged slides, added lattice

--- 2013-07-31 exportfunctions/base4.tex
@@
-\documentclass[11pt]{article}
-\usepackage{tikz}
-\usetikzlibrary{decorations,arrows,shapes}
-\begin{document}\begin{tikzpicture}[scale=1]
-
+\begin{tikzpicture}[scale=.5]
 \node (H1) at (225bp,-125bp)[draw,circle,fill=white] {$\mu_1$};
@@
 \node (H4) at (225bp,-325bp)[draw,circle,fill=white] {$\mu_4$};
-\draw [line width=1pt] (H1) to (325bp, -124bp) node[fill=blue!20] {$H_{1}$} to (H2);
-\draw [line width=1pt] (H2) to (425bp, -224bp) node[fill=blue!20] {$H_{4}$} to (H3);
-\draw [line width=1pt] (H4) to (225bp, -227bp) node[fill=blue!20] {$H_{3}$} to (H1);
+\only<2,5,6,8>{\draw [line width=1pt] (H1) to (325bp, -124bp) node[fill=blue!20] {$H_{1}$} to (H2);}
+\only<2,4,7,8>{\draw [line width=1pt] (H1) to (275bp, -175bp) node[fill=blue!20] {$H_{2}$} to (H3);}
+\only<2,3,6,7,9>{\draw [line width=1pt] (H2) to (425bp, -224bp) node[fill=blue!20] {$H_{4}$} to (H3);}
+\only<3,4,6,8,9,10>{\draw [line width=1pt] (H3) to (326bp, -324bp) node[fill=blue!20] {$H_{6}$} to (H4);}
+\only<4,5,6,7>{\draw [line width=1pt] (H4) to (225bp, -227bp) node[fill=blue!20] {$H_{3}$} to (H1);}
+\only<3,5,7,8,9,10>{\draw [line width=1pt] (H4) to (374bp, -176bp) node[fill=blue!20] {$H_{5}$} to (H2);}
+\only<9,10>{\draw [line width=2pt,color=red,dashed] (H1) to (325bp, -124bp) node[fill=blue!20] {$H_{1}$} to (H2);}
+\only<9,10>{\draw [line width=2pt,color=red,dashed] (H1) to (275bp, -175bp) node[fill=blue!20] {$H_{2}$} to (H3);}
+\only<9>{\draw [line width=2pt,color=red,dashed] (H2) to (425bp, -224bp) node[fill=blue!20] {$H_{4}$} to (H3);}
+
 \end{tikzpicture}
-\end{document}

### 2013-08-16 presentation/presentation.tex
added cycles, rearranged slides, added lattice

@@
 \begin{document}
+\frametitle{Matroids}
 
@@
 \begin{frame}
-\frametitle{Example 1: 4 Groups}
-\input{full4ext}
-\end{frame}
-
-\begin{frame}
 \frametitle{Pairwise mean comparisons (two-sided)}
@@
 \begin{frame}
-\frametitle{Example 1: 4 Means}
-\input{full4}
+\frametitle{Example 1: 4 Groups}
+\input{full4ext}
 \end{frame}
+
+% \begin{frame}
+% \frametitle{Example 1: 4 Means}
+% \input{full4}
+% \end{frame}
 
@@
 \item $H_I$ independent iff for all $i \in I$, $i \nsim_H I$,
-\item $I$ independent $\Leftrightarrow$ $H_{I} \subsetneq  H_{I\setminus i}$
+\item $I$ independent $\Leftrightarrow$ $H_{I} \subsetneq
+  H_{I\setminus i}$ for all $i \in I$
+\pause
 \item For pairwise comparisons $I$ independent iff there are no cycles in $\mathscr{G}_I$
@@
 
-\item A matroid on $S$ endows $\mathcal{P}(S)$ with an algebraic
+\item Matroids generalize the structures of graphs and vector spaces,
+  \eg cycles, bases, independent sets
+\item A matroid $M$ on $S$ endows $\mathcal{P}(S)$ with an algebraic
   structure 
-\item Matroids generalize the structures of graphs and vector spaces,
-  \eg circuits, bases, independent sets
-\item Many ways to define a matroid
+\item Cryptomorphism: many ways to define a matroid: 
   \begin{itemize}
-  \item From a graph perspective: Circuit axioms
-  \item From a vector space perspective: Independence axioms
+  \item From a graph perspective: Circuit axioms.
+  \item From a vector space perspective: Independence axioms.
   \end{itemize}
 \item There is a dual matroid $M^*$ which can be helpful for proofs and algorithms 
-\item \textbf{/Some illustration hereabouts/}
 \end{itemize} % ends low level
@@
   \begin{minipage}{.4\textwidth}
-    \input{cycle}
+    \only<2->{\input{cycle}}
+.
   \end{minipage} &
@@
 
-\begin{frame}{Maximally independent sets}
+\section{Matroids from logically dependent hypotheses}
+\begin{frame}
+  \frametitle{Dependent hypotheses}
+  \begin{block}{Equivalence of dependence relations $\sim_H = \sim_M$}
+    \begin{itemize}
+    \item<1-> $H_i$ depends on $H_I$ iff $H_{I \cup i} = H_{I}$
+    \item<2-> Does this mean $\mathscr{H}$ defines a matroid $M$, such that
+      $\sim_H = \sim_M$?
+    \end{itemize}
+  \end{block}
+\end{frame}
+
+\begin{frame}
+\frametitle{Matroids from logically dependent hypotheses}
+
+  \begin{block}{Two-sided pariwise comparisons}
+    For a family of two-sided pairwise mean comparisons $H_i$, $i \in
+    S$, an elementary hypotheses $H_j$ depends on intersection
+    hypotheses $H_I$ iff the corresponding edge $e_j$ is dependent on
+    the edge set $\{e_i:i \in I\}$ of the cycle matroid defined by 
+    $\mathscr{G}_I$.   
+  \end{block}
+
+  \begin{block}{Two-sided linear contrast tests}
+    For a family of two-sided general contrast tests $H_i$, $i \in S$,
+    an elementary hypothesis $H_j$ depends on intersection hypotheses
+    $H_I$ iff the corresponding contrast vector $\bs{c_j}$ is
+    dependent on the column vectors of the matrix $\bs{C}_I$ in the
+    vectorial matroid defined by $\bs{C}$ on $\mathbb{R}^m$.
+  \end{block}
+
+\end{frame}
+
+\begin{frame}
+  \frametitle{Matroids from null hypotheses}
+  {\Large What is it good for?\\
+  \invisible<1>{Structure}}
+\end{frame}
+
+\begin{frame}
+\frametitle{Maximally independent sets}
   \begin{itemize}
@@
 \item The closure of a set $\sigma I = \{i \in S: i \sim I\}$
-\item Closed sets {\em flats}: $\mathscr{L} = \{I \subseteq S: \sigma I = I\}$
+\item Closed sets {\em flats}: $\mathscr{L} = \{\sigma I: I \subseteq S\}$
 \item Hyperplanes are flats of rank $\rho S - 1$
@@
 
-\section{Matroids from logically dependent hypotheses}
+\begin{frame}
+\frametitle{Flats}
+\begin{itemize}
+
+\item The flats of a matroid are closed under intersection
+
+\item Every elementary hypothesis is a flat
+\item For each $I \subseteq S$, $H_I = H_{\sigma I}$
+\item Define a stepwise procedure on the flats of $M$
+
+\end{itemize} % ends low level
+\end{frame}
 
 \begin{frame}
-\frametitle{Matroids from logically dependent hypotheses}
-
-  \begin{block}{Theorem}
-    For a family of two-sided pairwise mean comparisons $H_i$, $i \in
-    S$, an intersection hypotheses $H_I$ is independent iff the
-    corresponding edge set $\{e_i:i \in I\}$ is independent in the
-    cycle matroid on $\mathscr{G}_I$.
-
-For a family of two-sided
-    general contrast tests $H_i$, $i \in S$, an intersection
-    hypotheses $H_I$ is independent iff the corresponding contrast
-    matrix $\bs{C}_I$ is independent in the vectorial matroid on
-    $\mathbb{R}^m$.
-
-    \begin{displaymath}
-      i \sim_H I \Leftrightarrow i \sim_M I
-    \end{displaymath}
+  \frametitle{Stepwise test procedure}
+  \begin{block}{Multiple testing procedure for logically related hypotheses}
+    Assume that $\mathscr{H}$ with dependence relation $\sim_H$
+    defines a matroid on $M$ on $S$. Then a test procedure that
+    rejects elementary hypotheses $H_i$ if all $H_E$, $i \in E$, $E
+    \in \mathscr{L}$ are rejected using a local level-$\alpha$ test
+    controls the FWER in the strong sense.
   \end{block}
@@
 \begin{frame}
-\frametitle{Closed test procedure}
+  \frametitle{Stepwise multiple testing procedure on the lattice of
+    flats}
+
+\begin{tabular}{ll}
+  \begin{minipage}{.3\textwidth}
+    \includegraphics<1>[width=1.2\textwidth]{lattice4.pdf}
+    \includegraphics<2>[width=1.2\textwidth]{lattice4accept.pdf}
+    \includegraphics<3>[width=1.2\textwidth]{lattice4reject.pdf}
+    \includegraphics<4>[width=1.2\textwidth]{lattice4rejacc.pdf}\\
+    {\tiny wikipedia:Partition\_of\_a\_Set}
+  \end{minipage} &
+  \begin{minipage}{.7\textwidth}
   \begin{itemize}
+  \item All flats are given by intersections of hyperpanes
+    $\rightarrow$ knowing all hyperplanes remaining flats are given by
+    intersection of index sets
+  \item Hyperplanes can by found by algorithms that enumerate the
+    circuits of the dual matroid \cite{boros2003algorithms}
+  \item The partial order of flats defined by inclusion defines a
+    geometric lattice
+  \item<2-> Testing only a subset of elementary hypotheses corresponds
+    to intervals
+  \item<3-> Rejection of elementary hypotheses removes intervals
+  \end{itemize}
 
-\item The flats of a matroid are closed under intersection
+  \end{minipage}
+\end{tabular}
 
-\item Every elementary hypothesis is a flat
-
-\item Using local level-$\alpha$ tests for each flat we have a closed test procedure that controls the FWER in the strong sense (Closed test principle)
-
-\end{itemize} % ends low level
 \end{frame}
 
-\begin{frame}
-\frametitle{Examples}
+\begin{frame}{Matroid procedures for pairwise mean comparisons}
+\begin{tabular}{ll}
+  \begin{minipage}{.3\textwidth}
+    \includegraphics[width=1.1\textwidth]{setpartitions.pdf} \\
+     {\tiny wikipedia:Bell\_number}
+  \end{minipage} &
+  \begin{minipage}{.7\textwidth}
   \begin{itemize}
+  \item Two-sided comparisons of $m$ means
+  \item The cycle matroid of the graph representation
+    provides equivalent dependence structure
+  \item The flats of a (simple) complete graph are
+    isomporphic to the partitions of a set of size $m$
+  \item Efficient algorithms to generate all exhaustive sets by
+    enumeration of set partitions \cite{er1988fast,knuth2005art,kokosinski2006new} 
+  \item The number of flats are $m$-th Bellnumber $-1$ \ie 4, 14,
+    51, 202, 876, 4139, ... 
+  \end{itemize}
+  \end{minipage}
+\end{tabular}
 
-\item Pairwise comparisons
-
-\item The flats of the complete graph correspond to the partitions of
-  $\{1,...,m\}$
-\item This provides efficient methods to generate all flats for
-  pairwise mean comparisons \cite{er1988fast,knuth2005art,kokosinski2006new}
-
-\item The flats of a matroid generate a lattice, procedures testing not all possible comparisons correspond to intervals on that lattice
-
-\end{itemize} % ends low level
 \end{frame}
@@
 \begin{itemize}
-\item \cite{westfall1997multiple} checks each intersection hypotheses
-  whether it is flat using dependence arguments.
-\item \cite{westfall2007multiple} provide a branch and bound algorithm
-  to generate all flats. They note that for pairwise comparisons all
-  exhaustive hypotheses are {\em partition hypotheses} 
+\item \cite{shaffer1986modified} represents pairwise comparisons
+  by set partitions 
 \item \cite{bergmann1988improvements} define the concepts of an
@@
   flat using rank function.
+\item \cite{westfall1997multiple} checks each intersection hypotheses
+  whether it is flat using dependence arguments.
+\item \cite{weichert2000robuste} suggest an MTP based on clique
+  decomposition of the graph
+\item \cite{westfall2007multiple} provide a branch and bound algorithm
+  to generate all flats. They note that for pairwise comparisons all
+  exhaustive hypotheses are {\em partition hypotheses} 
+\end{itemize}
+\end{frame}
 
-  \cite{hommel1999bonferroni} show {\em closure axioms} for exhaustive
-  sets 
-\item Flats <-> Weakly Exhaustive set (Hommel, Bergmann)
-
-\item Lemma 1.5 in (Bernhard) about weakly exhaustive sets are ``closure axioms'' of a matroid
-
-\item Graphical procedures for pairwise comparisons (Weichert) identify dependence using cycle in the graph
-
-\item \emph{However:}
+\begin{frame}
+\frametitle{Matroid representation}
   \begin{itemize}
@@
   \item efficient algorithms exist to find the flats of matroids,
+  \item {\em Scumm principle}\cite{welsh2010matroid} everything
+    important happens at the top (hyperplanes)
   \item we may define procedures using independent sets,
@@
   \end{itemize}
-
-\end{itemize} % ends low level
-\end{frame}
-
-\begin{frame}{Matroid procedures for pairwise mean comparisons}
-
-\begin{tabular}{ll}
-  \begin{minipage}{.3\textwidth}
-    \includegraphics[width=1.1\textwidth]{setpartitions.pdf} \\
-     {\tiny wikipedia:Bell\_number}
-  \end{minipage} &
-  \begin{minipage}{.7\textwidth}
-  \begin{itemize}
-  \item Two-sided comparisons between $m$ means
-  \item The cycle matroid of the corresponding graph representation
-    provides an equivalent dependence structure
-  \item The flats of a (simple) complete graph are
-    isomporphic to the partitions of a set of size $m$
-  \item Provides an efficient way to generate all exhaustive sets by
-    enumeration of all partitions or hyperplanes \cite{er1988fast,knuth2005art,kokosinski2006new} 
-  \item The number of flats is given by the Bell-numbers \ie 5, 15,
-    52, 203, 877, 4140, ... 
-  \end{itemize}
-  \end{minipage}
-\end{tabular}
-
 \end{frame}
@@
     \begin{itemize}
-    \item Provides weights $w_{i,I}$ for each elementary hypotheses
-      $H_i$ and each non-empty subset $I \subset S$
+    \item Provides weights $w_{i,I}$ for each index $i$ and each
+      non-empty subset $I \subset S$
     \item Weights are used to define a closed test procedure of

### 2013-08-19 presentation/presentation.tex
fixed a couple of typos, added some slides towards one-sided hypotheses

@@
 \usepackage{bm}
+\usepackage{ulem}
 \tolerance=1000
@@
 \newcommand{\parentd}{../}
-
+\newcommand{\R}{\mathbb{R}}
+\newcommand{\indf}[1]{\mathbf{1}_{\left\{#1\right\}}}
 \providecommand{\alert}[1]{\textbf{#1}}
@@
   \begin{itemize}
-  \item All flats are given by intersections of hyperpanes
+  \item All flats are given by intersections of hyperplanes
     $\rightarrow$ knowing all hyperplanes remaining flats are given by
@@
     enumeration of set partitions \cite{er1988fast,knuth2005art,kokosinski2006new} 
-  \item The number of flats are $m$-th Bellnumber $-1$ \ie 4, 14,
+  \item The number of flats are $m$-th Bell-Number $-1$ \ie 4, 14,
     51, 202, 876, 4139, ... 
@@
   exhaustive sets that are basically {\em closure axioms} of a matroid
-  and provides algorithms to check each intersection hypotheses if
-  flat using rank function.
-\item \cite{westfall1997multiple} checks each intersection hypotheses
-  whether it is flat using dependence arguments.
+  and provides algorithms to check each intersection hypotheses if is
+  a flat using rank function.
+\item \cite{westfall1997multiple} generates flats using dependence
+  arguments in vector spaces.
 \item \cite{weichert2000robuste} suggest an MTP based on clique
-  decomposition of the graph
+  decomposition of the graph, cliques in the graph are flats
 \item \cite{westfall2007multiple} provide a branch and bound algorithm
@@
   \item efficient algorithms exist to find the flats of matroids,
-  \item {\em Scumm principle}\cite{welsh2010matroid} everything
+  \item {\em Scum principle}\cite{welsh2010matroid} everything
     important happens at the top (hyperplanes)
@@
 \begin{frame}
-\frametitle{Matroid conditions}
+\frametitle{General Results}
+\begin{itemize}
+\item We give sufficient (and probably necessary) conditions for a
+  general family of hypotheses $\mathscr{H}$ to define a matroid
+\item We give sufficient (and probably necessary) conditions on
+  $\mathscr{H}$ for $\mathscr{L}$ to be the set of strictly exhaustive
+  sets
+\item Systems with one sided hypotheses, and parallel hypotheses (\eg
+  combined non-inferiority/superiority) do not give matroids with
+  $\mathscr{L}$ all strictly exhaustive 
+\end{itemize}
+
 \end{frame}
@@
 \begin{frame}
-\frametitle{Flats <-> Weakliy exhaustive sets}
+  \frametitle{One sided hypotheses}
+  \begin{block}{One-sided linear contrast hypotheses}
+    A family of {\em one sided linear contrast hypotheses} is defined
+    by a contrast matrix $\bs{A} \in \mathbb{R}^{m\times n}$ and
+    vector $\bs{b} \in \mathbb{R}^n$ that define a  family of null
+    hypotheses of the form $\mathscr{H} = \{H_i: \bs{A_i}^t\bs{\mu}
+    \leq b_i,i \in \{1,...,n\}\}$, where $\bs{A}_i$ denotes the $i$-th
+    column of $\bs{A}$ and $b_i$ the $i$-th componenth of $\bs{b}$.
+  \end{block}
 \end{frame}
@@
 \begin{frame}
-\frametitle{Flats <-> Strictly exhaustive sets}
+  \frametitle{Hyperplane arrangement}
+  \begin{itemize}
+  \item Each hypothesis $H_i$ corresponds to a hyperplane $h_i =
+    \{\bs{x} \in \mathbb{R}^m: \bs{A}_i^t\bs{x} = b_i\}$
+  \item Each hyperplane $h_i$ divides $\mathbb{R}^m$ into two regions
+    \begin{itemize}
+    \item $h_i^+ = \{\bs{x} \in \mathbb{R}^m: \bs{A}_i^t\bs{x} <
+      b_i\}$ 
+    \item $h_i^- = \{\bs{x} \in \mathbb{R}^m: \bs{A}_i^t\bs{x}
+      > b_i\}$
+    \end{itemize}
+  \item Defines a sign pattern 
+    \begin{equation}
+      \label{eq:sign}
+      \delta_i(\bs{x}) = \left\{
+        \begin{array}{ll}
+          + & \text{if } \bs{x} \in h_i^+\\
+          0 & \text{if } \bs{x} \in h_i \\
+          - & \text{if } \bs{x} \in h_i^-\\
+        \end{array}\right.
+    \end{equation}
+  \item $\delta_i$ is $-$ if $H_i$ false at $x$, or $0$ or $+$
+    otherwise. $(\delta_1,...,\delta_n)$ tells us to which combination
+    of true and false null hypotheses some $\bs{x}$ belongs.
+
+  \item Set of all sign patterns defined by $\mathscr{H}$:
+    \begin{equation}
+      \label{eq:faces}
+      \mathcal{F}(\mathscr{H}) = \{\bs{\delta}(\bs{x}): \bs{x} \in \R^m\}
+    \end{equation}
+  \end{itemize}
 \end{frame}
+
+\begin{frame}
+  \frametitle{Example}
+
+\begin{tabular}{ll}
+  \begin{minipage}{.4\textwidth}
+  \includegraphics[height=1\textwidth]{spheresfukuda.eps}
+  \end{minipage} &
+  \begin{minipage}{.6\textwidth}
+    \begin{itemize}
+    \item 4 contrast hypotheses about 3 means
+    \item Cut the hyperplanes with the sphere in $\R^3$ (makes
+      visualisation easier)
+    \item Cells of the sphere arrangment correspond to sign patterns
+      of $\mathscr{H}$
+    \item The abstraction of such an arrangement is an \alert{Oriented Matroid}
+    \end{itemize}
+  \end{minipage}
+\end{tabular}
+
+\end{frame}
+
+\begin{frame}
+  \frametitle{Stepwise multiple testing procedure for one-sided linear
+    contrast hypotheses}
+
+  Consider a family of general contrast hypotheses $\mathscr{H} =
+  \{H_i: \bs{A_i}^t\bs{\mu} \leq b_i,i \in \{1,...,n\}\}$ and let
+  $\mathscr{F}(\mathscr{H}) = \{ \bs{\delta}(\bs{x}): \bs{x} \in
+  \R^m\}$ denote the corresponding set of sign patterns. Consider a
+  set of decision rules $\varphi_F: \R^n \rightarrow \{0,1\}$ for
+  which hold $P(\varphi_F = 1 | H_i: F_i \geq 0) \leq \alpha$. Then a
+  corresponding decision rule for each elementary hypotheses $H_i$ is
+  given by 
+  \begin{equation}
+    \label{eq:signtest}
+    \psi_i = \min_{F: F_i \geq 0} \varphi_F,
+  \end{equation}
+  which controls the FWER at level $\alpha$ in the strong sense.
+
+\end{frame}
+
+\begin{frame}
+  \frametitle{Computation of $\mathscr{F}$}
+  The question remains how to compute $\mathcal{F}$. For this observe
+that the $h_i$ define a hyperplane configuration and $\mathcal{F}$ is
+just the face lattice of the corresponding oriented
+matroid. Consequently, $\mathcal{F}$ can be found using algorithms
+for arrangement construction from polyhedral programming
+\cite{Avis-KF-92,Avis-KF-96,Ferrez-KF-Liebling-01} which are
+polynomial in time and input size and compact in space!
+
+\end{frame}
+
+\begin{frame}
+  \frametitle{Example}
+  3 ordered means 
+\end{frame}
+
 \section{Summary \& Outlook}
@@
 \frametitle{Summary}
+\begin{itemize}
+\item Matroids provide usefull abstraction for systems of algebraically dependent hypotheses
+\end{itemize}
 \end{frame}
@@
 \frametitle{Outlook}
-
+\begin{itemize}
+\item MCPmod on \sout{steroids} matroids
+\end{itemize}
 \end{frame}

### 2013-08-19 presentation/presentation.tex
incorporated willi's changes

--- 2013-08-15 presentation/presentation.tex
@@
   \item "Static":  After j rejections, test remaining hypotheses at level $\alpha/t_j$, $t_j$ = max number of true hypotheses if $j$ are false.
-  \item "Dynamic": After rejecting $H_{i_1},...,H_{i_j}$, test remaining hypotheses at level $\alpha/t^*_j$,  $t^*_j$ = number of true hypotheses if $H_{i_1},...,H_{i_j}$ are false.
+  \item "Dynamic": After rejecting $H_{i_1},...,H_{i_j}$, test remaining hypotheses at level $\alpha/t^*_j$,  $t^*_j$ = max. number of true hypotheses if $H_{i_1},...,H_{i_j}$ are false.
   \item Example: $m=4, n=6$ :
@@
       \item Static: $t_0=6, t_1,t_2,t_3 = 3, t_4=2, t_5=1$
-      \item Dynamic: if, e.g., $H_{\{1,2\}}$ and $H_{\{3,4\}}$ are rejected then only $H_{\{1,3\}}$ and $H_{\{2,3\}}$, or $H_{\{1,4\}}$ and $H_{\{2,3\}}$ can be simultaneously true, hence $t^*_2 = 2 < t_2 =3$ in this case.
+      \item Dynamic: $t^*_j \le t_j$, hence more powerful than "static" 
            \end{itemize}
@@
 \begin{frame}
-\frametitle{Example 1: 4 Groups}
-\input{full4ext}
-\end{frame}
-
-\begin{frame}
 \frametitle{Pairwise mean comparisons (two-sided)}
+\begin{tabular}{ll}
+  \begin{minipage}{.3\textwidth}
+    \only<2->{\input{full4}}
+.
+  \end{minipage} &
+  \begin{minipage}{.7\textwidth}
   \begin{itemize}
-
 \item Consider $\mu_i, i \in \{1,...,m\}$
@@
 \end{itemize} % ends low level
-\end{frame}
-
-\begin{frame}
-\frametitle{Example 1: 4 Means}
-\input{full4}
-\end{frame}
-
-\begin{frame}
-\frametitle{General contrast tests (two-sided)}
-\begin{itemize}
-
-\item Consider $m$-variate random variable $\bs{X} \sim F(\bs{\mu})$, $\bs{\mu} = (\mu_1,...,\mu_m)^t$
-
-\item Contrasts $\bs{c}_i = (c_{i1},...,c_{im})^t$, $\sum_{j = 1}^m
-  c_{ij} = 0$
-\item Contrast matrix  $\bs{C} = \left( \bs{c}_1,...,\bs{c}_n\right)$,
-  with sub-matrix $\bs{C}_I = \left( \bs{c}_i \right)_{i \in
-    I}$
-\item Elementary hypotheses $H_i: \bs{c}_i^t\bs{\mu} = 0$,
-\item Intersection hypotheses $H_I: \bs{C}_I^t\bs{\mu} = 0$
-
-\item Elementary hypotheses are represented by column vectors,
-  intersection hypotheses by sub-matrices of the contrast matrix
-
-\end{itemize} % ends low level
-\end{frame}
-
-\begin{frame}
-\frametitle{Example 2: Contrast test}
-
-\begin{tabular}{ll}
-  \begin{minipage}{.45\textwidth}
-    All pairwise comparisons of 4 means 
-    \begin{displaymath}
-    \bs{C}^t = \left(
-      \begin{array}{llll}
-      -1 & 1 & 0 & 0 \\
-      -1 & 0 & 1 & 0 \\
-      -1 & 0 & 0 & 1 \\
-      0 & -1 & 1 & 0 \\
-      0 & -1 & 0 & 1 \\
-      0 & 0 & -1 & 1 
-    \end{array}
-    \right)
-    \end{displaymath}
-  \end{minipage} &
-  \begin{minipage}{.55\textwidth}
-    Modified Williams Contrasts for Dose Response
-    \begin{displaymath}
-      \bs{C}^t=\left(
-      \begin{array}{llll}
-        1 & 0 & 0 & -1\\
-        1 & 0 &-.51 & -.49 \\
-        1 & -.35 & -.33 & -.32\\
-        .51 & .49 & -.51 & -.48\\
-        .51 & .49 & 0 &  -1\\
-        .35 & .33 & .32& -1
-      \end{array}\right)
-    \end{displaymath}
-  \end{minipage}
-\end{tabular}
-
-\end{frame}
-
-\begin{frame}
-\frametitle{Dependent null hypotheses}
-
-\begin{itemize}
-\item $H_i$ depends on $H_I$ iff $H_{I \cup i} = H_{I}$, induces a
-  dependence relation $\sim_H$ on $\mathcal{P}(S)$
-\item $H_I$ independent iff for all $i \in I$, $i \nsim_H I$,
-\item $I$ independent $\Leftrightarrow$ $H_{I} \subsetneq  H_{I\setminus i}$
-\item For pairwise comparisons $I$ independent iff there are no cycles in $\mathscr{G}_I$
-\item For general contrast tests $I$ independent iff $\bs{C_I}$ has full
-  column rank
-\end{itemize} % ends low level
-\end{frame}
-
-\section{Matroids}
-
-\begin{frame}
-\frametitle{Matroids}
-\begin{itemize}
-
-\item A matroid on $S$ endows $\mathcal{P}(S)$ with an algebraic
-  structure 
-\item Matroids generalize the structures of graphs and vector spaces,
-  \eg circuits, bases, independent sets
-\item Many ways to define a matroid
-  \begin{itemize}
-  \item From a graph perspective: Circuit axioms
-  \item From a vector space perspective: Independence axioms
-  \end{itemize}
-\item There is a dual matroid $M^*$ which can be helpful for proofs and algorithms 
-\item \textbf{/Some illustration hereabouts/}
-\end{itemize} % ends low level
-\end{frame}
-
-\begin{frame}
-\frametitle{Circuit axioms}
-
-\begin{tabular}{ll}
-  \begin{minipage}{.4\textwidth}
-    \input{cycle}
-  \end{minipage} &
-  \begin{minipage}{.6\textwidth}
-    Consider $S$ and a collection of subsets $\mathscr{C}$ for which holds
-    \begin{enumerate}
-    \item $C_1 \neq C_2 \in \mathscr{C}$, then $C_1 \nsubseteq C_2$,
-    \item $C_1 \neq C_2 \in \mathscr{C}$, and $i \in C_1 \cap C_2$ then
-      $\exists C_3 \subseteq (C_1 \cup C_2)$,
-    \end{enumerate}
-    then $\mathscr{C}$ is the set of circuits of a matroid $M$ on $S$
-  \end{minipage}
-\end{tabular}
-
-\end{frame}
-
-% \item \textbf{/Some illustration hereabouts/}
-
-% \end{itemize} % ends low level
-% \end{frame}
-
-\begin{frame}
-\frametitle{Independence axioms}
-
-\begin{tabular}{ll}
-  \begin{minipage}{.4\textwidth}
-  \only<1>{\input{independent}}
-  \only<2>{\input{cycle}}
-  \invisible<1>{\alert{\footnotesize minimally dependent set!}}
-  \end{minipage} &
-  \begin{minipage}{.6\textwidth}
-  Consider $S$ and a collection of subsets $\mathscr{I}$ for which
-  holds
-  \begin{enumerate}
-  \item $\emptyset \in \mathscr{I}$,
-  \item $I \in \mathscr{I}$ and $J \subseteq I$ then $J \in
-    \mathscr{I}$,
-  \item $I,J \in \mathscr{I}$, $|I| = |J + 1|$ then $\exists j \in I
-    \setminus J$ such that $J \cup i \in \mathscr{I}$,
-  \end{enumerate}
-  then $\mathscr{I}$ is the set of independent sets of a matroid $M$
-  on $S$. This defines an independence relation $i \nsim_M I$ iff
-  $I\cap i \in \mathscr{I}$.
-
-  \end{minipage}
-\end{tabular}
-
-\end{frame}
-
-\begin{frame}
-\frametitle{Dependence axioms}
-
-\begin{tabular}{ll}
-  \begin{minipage}{.4\textwidth}
-    \input{dependent}
-  \end{minipage} &
-  \begin{minipage}{.6\textwidth}
-  Relation $\sim_M$ on elements and subsets of $S$,
-  if for any $x \in S$, and sets $Y,Z \subseteq S$ holds
-  \begin{enumerate}
-  \item $y \sim Y$ for all $y \in Y$
-  \item if $x \sim_M Y$ and $x \nsim_M Y\setminus y$
-    then $y \sim x \cup (Y \setminus y)$
-  \item if $x \sim_M Y$ and for all $y \in Y$, $y \sim_M Z$ then $x \sim_M Z$
-  \end{enumerate}
-  then $\sim_M$ defines a matroid on $S$.
-  \end{minipage}
-\end{tabular}
-
-\end{frame}
-
-\begin{frame}{Maximally independent sets}
-  \begin{itemize}
-  \item Bases are maximally independent sets of $S$
-  \item Rank $\rho I$ is the cardinality of a maximally indepedendent set in $I$
-  \item Circuits are minimally dependent sets, \ie each subset of $C$
-    with rank $\rho C - 1$ is a maximally independent set of $C$
-  \end{itemize}
-\end{frame}
-
-\begin{frame}
-\frametitle{Maximally dependent sets}
-\begin{itemize}
-\item The closure of a set $\sigma I = \{i \in S: i \sim I\}$
-\item Closed sets {\em flats}: $\mathscr{L} = \{I \subseteq S: \sigma I = I\}$
-\item Hyperplanes are flats of rank $\rho S - 1$
-\item All flats are given by closing hyperplanes under intersection
-\item Circuits in $M^*$ are the complements of hyperplanes in $M$
-\end{itemize} % ends low level
-\end{frame}
-
-\begin{frame}{Examples}
-\framesubtitle<1>{Independent set}  
-\framesubtitle<2>{Base}  
-\framesubtitle<4>{Dependent set}  
-\framesubtitle<3>{Cycle}  
-\framesubtitle<5>{Closure}  
-\only<1>{\input{independent}}
-\only<2>{\input{base}}
-\only<4>{\input{dependent}}
-\only<3>{\input{cycle}}
-\only<5>{\input{closure}}
-\end{frame}
-
-\section{Matroids from logically dependent hypotheses}
-
-\begin{frame}
-\frametitle{Matroids from logically dependent hypotheses}
-
-  \begin{block}{Theorem}
-    For a family of two-sided pairwise mean comparisons $H_i$, $i \in
-    S$, an intersection hypotheses $H_I$ is independent iff the
-    corresponding edge set $\{e_i:i \in I\}$ is independent in the
-    cycle matroid on $\mathscr{G}_I$.
-
-For a family of two-sided
-    general contrast tests $H_i$, $i \in S$, an intersection
-    hypotheses $H_I$ is independent iff the corresponding contrast
-    matrix $\bs{C}_I$ is independent in the vectorial matroid on
-    $\mathbb{R}^m$.
-
-    \begin{displaymath}
-      i \sim_H I \Leftrightarrow i \sim_M I
-    \end{displaymath}
-  \end{block}
-
-\end{frame}
-
-\begin{frame}
-\frametitle{Closed test procedure}
-\begin{itemize}
-
-\item The flats of a matroid are closed under intersection
-
-\item Every elementary hypothesis is a flat
-
-\item Using local level-$\alpha$ tests for each flat we have a closed test procedure that controls the FWER in the strong sense (Closed test principle)
-
-\end{itemize} % ends low level
-\end{frame}
-
-\begin{frame}
-\frametitle{Examples}
-\begin{itemize}
-
-\item Pairwise comparisons
-
-\item The flats of the complete graph correspond to the partitions of
-  $\{1,...,m\}$
-\item This provides efficient methods to generate all flats for
-  pairwise mean comparisons \cite{er1988fast,knuth2005art,kokosinski2006new}
-
-\item The flats of a matroid generate a lattice, procedures testing not all possible comparisons correspond to intervals on that lattice
-
-\end{itemize} % ends low level
-\end{frame}
-
-\begin{frame}
-\frametitle{Connections to existing procedures}
-
-\textbf{Old wine in new bottles ((c) L. Hothorn)}
-\begin{itemize}
-\item \cite{westfall1997multiple} checks each intersection hypotheses
-  whether it is flat using dependence arguments.
-\item \cite{westfall2007multiple} provide a branch and bound algorithm
-  to generate all flats. They note that for pairwise comparisons all
-  exhaustive hypotheses are {\em partition hypotheses} 
-\item \cite{bergmann1988improvements} define the concepts of an
-  exhaustive sets
-\item \cite{bernhard1991computergestuetzte} shows properties of
-  exhaustive sets that are basically {\em closure axioms} of a matroid
-  and provides algorithms to check each intersection hypotheses if
-  flat using rank function.
-
-  \cite{hommel1999bonferroni} show {\em closure axioms} for exhaustive
-  sets 
-\item Flats <-> Weakly Exhaustive set (Hommel, Bergmann)
-
-\item Lemma 1.5 in (Bernhard) about weakly exhaustive sets are ``closure axioms'' of a matroid
-
-\item Graphical procedures for pairwise comparisons (Weichert) identify dependence using cycle in the graph
-
-\item \emph{However:}
-\begin{itemize}
-\item proofs are shorter (more elegant),
-\item efficient algorithms exist to find the flats of matroids,
-\item we may define procedures using independent sets,
-\item \ldots{}
-\end{itemize}
-
-\end{itemize} % ends low level
-\end{frame}
-
-\begin{frame}{Matroid procedures for pairwise mean comparisons}
-
-\begin{tabular}{ll}
-  \begin{minipage}{.3\textwidth}
-    \includegraphics[width=1.1\textwidth]{setpartitions.pdf} \\
-     {\tiny wikipedia:Bell\_number}
-  \end{minipage} &
-  \begin{minipage}{.7\textwidth}
-  \begin{itemize}
-  \item Two-sided comparisons between $m$ means
-  \item The cycle matroid of the corresponding graph representation
-    provides an equivalent dependence structure
-  \item The flats of a (simple) complete graph are
-    isomporphic to the partitions of a set of size $m$
-  \item Provides an efficient way to generate all exhaustive sets by
-    enumeration of all partitions or hyperplanes \cite{er1988fast,knuth2005art,kokosinski2006new} 
-  \item The number of flats is given by the Bell-numbers \ie 5, 15,
-    52, 203, 877, 4140, ... 
-  \end{itemize}
-  \end{minipage}
-\end{tabular}
-
-\end{frame}
-
-\section{Graphical approaches for logically dependent hypotheses}
-
-\begin{frame}
-\frametitle{Reminder: Graphical approaches}
-
-\begin{tabular}{ll}
-  \begin{minipage}{.4\textwidth}
-    \begin{tikzpicture}[remember picture,overlay]
-      \node[yshift=-2.5cm,xshift=-1.4cm] at (current page.north west){\input{qvatest}};
-    \end{tikzpicture}
-
-  \end{minipage} &
-  \begin{minipage}{.6\textwidth}
-    \begin{itemize}
-    \item Provides weights $w_{i,I}$ for each elementary hypotheses
-      $H_i$ and each non-empty subset $I \subset S$
-    \item Weights are used to define a closed test procedure of
-      weighted intersection hypotheses tests
-    \item We have $\sum_{i \in S} w_{i,I} \leq 1$ (conservative),
-    \item and $w_{i,I} \leq w_{i,J}$ for $J \subseteq I$ (consonance).
-    \end{itemize}
   \end{minipage}
@@
 \begin{frame}
-  \frametitle{Example: Hierarchical LSD}
+\frametitle{Max number of true hypotheses}
 \begin{tabular}{ll}
   \begin{minipage}{.4\textwidth}
-    \input{lsd2.tex}
+   \input{shaffer}
+.
   \end{minipage} &
@@
   \begin{itemize}
-    \item 3 Hypotheses:
-      \begin{enumerate}
-      \item Placebo vs. High Dose
-      \item Placebo vs. Low Dose
-      \item High Dose vs. Low Dose
-      \end{enumerate}
-    \item Prefixed sequence: $H_1 \rightarrow H_2 \rightarrow H_3$
-    \item Test $H_1$ at level $\alpha$
-    \item If we reject $H_1$ test $H_2$ and $H_3$ at level $\alpha$
-      simultaneously
-    \item Can we generalize this?
-    \end{itemize}
-  \end{minipage}
-\end{tabular}
-
-\end{frame}
-
-\begin{frame}
-  \frametitle{Fixed sequence test of pairwise mean comparisons}
+  \item "Static":  $t_j$ = max number of true hypotheses if $j$ are false.
+  \item "Dynamic": $t^*_j $ = max. number of true hypotheses if $H_{i_1},...,H_{i_j}$ are false.
+  \item Example: $m=4, n=6, j = 2$ :
     \begin{itemize}
-\item {\bf Picture of hierarchical test hereabouts}
+      \item  {Static: if, e.g., $H_1$ and $H_2$ are false then at most $H_4, H_5 $ and $H_6$ can be simultaneously true, hence $t_2 = 3$}
+      \item  {Dynamic: if, e.g., $H_1$ and $H_6$ are false then only $H_3$ and $H_4$, or $H_3$ and $H_5$ can be simultaneously true, hence $t^*_2 = 2 < t_2 =3$.}
       \end{itemize}
 
-\begin{block}{Theorem}
-  Assume that $M(\mathscr{H})$ neither contains loops nor
-  parallels. Assume that hypotheses $H_1$ through $H_r$ have been
-  rejected. Consider that hypotheses $H_{r+1}$ through $H_{r+t}$ are
-  to be tested at level $\alpha_r$ according to a prefixed order.  If
-  for some $1 \leq s \leq t$ holds
-  \begin{equation}
-    \label{eq:cond.hierarchical}
-    \forall 1 \leq l < k \leq s: \; \exists i \leq r: i \sim \{(r+l),(r+k)\},
-  \end{equation}
-  hypotheses $H_{r+1}$ through $H_{r+s}$ can be tested simultaneously
-  at full level $\alpha_ra$.
-\end{block}
-
-\end{frame}
-
-\begin{frame}
-  \frametitle{Example: Multiple doses}
-
-\begin{tabular}{ll}
-  \begin{minipage}{.4\textwidth}
-    \begin{tikzpicture}[overlay]
-      \node[yshift=70bp,xshift=-40bp] at (current page.south west){\input{lsd4.tex}};
-    \end{tikzpicture}
-  \end{minipage} &
-  \begin{minipage}{.6\textwidth}
-    \begin{tikzpicture}[remember picture,overlay]
-      \node[yshift=100bp,xshift=160bp] at (current page.north west){\input{lsd4t.tex}};
-    \end{tikzpicture}
+\end{itemize} % ends low level
   \end{minipage}
@@
 \end{frame}
+%
+%
+%\begin{frame}
+%\frametitle{Pairwise mean comparisons (two-sided)}
+%\begin{itemize}
+%
+%\item Consider $\mu_i, i \in \{1,...,m\}$
+%
+%\item $H_k: \mu_i = \mu_j, k \in S = \{1,...,n\}$, where $n \leq m*(m-1)/2$
+%
+%\item Can be represented by an undirected graph $\mathscr{G} = (V,E)$:
+%\begin{itemize}
+%\item vertices correspond to means, \emph{i.e.}, $V = \{\mu_1,...,\mu_m\}$,
+%\item edges correspond to hypotheses, \emph{i.e.}, $E =
+%      \left\{e_1,...,e_n\} \right\}$, $e_l = \{\mu_i,\mu_j\}$, $i \neq
+%      j$, and $e_l \neq e_k$ for $l \neq k$
+%\end{itemize}
+%\item Let $H_I = \bigcap_{k \in I} H_k$, $I \subseteq S$
+%\item $I$ induces a subgraph $\mathscr{G}_I$ on the edge set $E_I=\{e_i: i
+%  \in I\}$
+%\end{itemize} % ends low level
+%\end{frame}
+%
+%\begin{frame}
+%\frametitle{Example 1: 4 Groups}
+%\input{full4ext}
+%\end{frame}
+%
 
-\begin{frame}
-  \frametitle{General weighting strategies}
-  \framesubtitle{QVA Example}
+% \begin{frame}
+% \frametitle{Example 1: 4 Means}
+% \input{full4}
+% \end{frame}
 
-\begin{tabular}{ll}
-  \begin{minipage}{.4\textwidth}
-    \begin{tikzpicture}[remember picture,overlay]
-      \node[yshift=-2.5cm,xshift=-1.4cm] at (current page.north west){\input{qvatest}};
-    \end{tikzpicture}
+%
+%\begin{frame}
+%\frametitle{General contrast tests (two-sided)}
+%\begin{itemize}
+%
+%\item Consider $m$-variate random variable $\bs{X} \sim F(\bs{\mu})$, $\bs{\mu} = (\mu_1,...,\mu_m)^t$
+%
+%\item Contrasts $\bs{c}_i = (c_{i1},...,c_{im})^t$, $\sum_{j = 1}^m
+%  c_{ij} = 0$
+%\item Contrast matrix  $\bs{C} = \left( \bs{c}_1,...,\bs{c}_n\right)$,
+%  with sub-matrix $\bs{C}_I = \left( \bs{c}_i \right)_{i \in
+%    I}$
+%\item Elementary hypotheses $H_i: \bs{c}_i^t\bs{\mu} = 0$,
+%\item Intersection hypotheses $H_I: \bs{C}_I^t\bs{\mu} = 0$
+%
+%\item Elementary hypotheses are represented by column vectors,
+%  intersection hypotheses by sub-matrices of the contrast matrix
+%
+%\end{itemize} % ends low level
+%\end{frame}
+%
 
-  \end{minipage} &
-  \begin{minipage}{.6\textwidth}
-    \begin{tikzpicture}[remember picture,overlay]
-      \node[yshift=-50bp,xshift=-180bp] at (current page.north east){\input{qvameans}};
-  \end{tikzpicture}
-  \end{minipage}
-\end{tabular}
-\end{frame}
-
-\begin{frame}
-\frametitle{General weighting strategies}
-
-\begin{enumerate}
-\item Pretest using weighted Bonferroni or conventional graphical
-  approach $R = \{i \in S: H_i \textrm{ rejected by pretest}\}$
-\item Compute all matroid hyperplanes $\mathscr{H}$ for the matroid on
-  the remaining hypotheses $I$
-\item $R \gets R \cup \{i: p_i \leq \min_J w_{i,J} \alpha, J \in \mathscr{H}\}$
-\item $A \gets \{\bigcup J: \min_j (p_j / w_{j,J}) > \alpha\}$ 
-\item $I \gets I \setminus R$
-\item Refine: $\mathscr{H} \gets \{J\in \mathscr{L}: \rho J = \rho
-  \mathscr{H} - 1\}$ 
-\item Prune: $\mathscr{H} \gets \{J \in \mathscr{H}: J \cap R = \emptyset\}$
-\item Repeat until $\rho \mathscr {H} = 0$
-\item Reject $R$ and $I \setminus A$
-\end{enumerate}
-\end{frame}
-\section{Matriods from general families of hypotheses}
-
-\begin{frame}
-\frametitle{Matroid conditions}
-\end{frame}
-
-\begin{frame}
-\frametitle{Flats <-> Weakliy exhaustive sets}
-\end{frame}
-
-\begin{frame}
-\frametitle{Flats <-> Strictly exhaustive sets}
-\end{frame}
-\section{Summary \& Outlook}
-
-\begin{frame}
-\frametitle{Summary}
-\end{frame}
-
-\begin{frame}
-\frametitle{Outlook}
-
-\end{frame}
-
-\bibliographystyle{alpha}
-\bibliography{\parentd paper/cleanrefs}
+%
+%\begin{frame}
+%\frametitle{Example 2: Contrast test}
+%
+%
+%\begin{tabular}{ll}
+%  \begin{minipage}{.45\textwidth}
+%    All pairwise comparisons of 4 means
+%    \begin{displaymath}
+%    \bs{C}^t = \left(
+%      \begin{array}{llll}
+%      -1 & 1 & 0 & 0 \\
+%      -1 & 0 & 1 & 0 \\
+%      -1 & 0 & 0 & 1 \\
+%      0 & -1 & 1 & 0 \\
+%      0 & -1 & 0 & 1 \\
+%      0 & 0 & -1 & 1
+%    \end{array}
+%    \right)
+%    \end{displaymath}
+%  \end{minipage} &
+%  \begin{minipage}{.55\textwidth}
+%    Modified Williams Contrasts for Dose Response
+%    \begin{displaymath}
+%      \bs{C}^t=\left(
+%      \begin{array}{llll}
+%        1 & 0 & 0 & -1\\
+%        1 & 0 &-.51 & -.49 \\
+%        1 & -.35 & -.33 & -.32\\
+%        .51 & .49 & -.51 & -.48\\
+%        .51 & .49 & 0 &  -1\\
+%        .35 & .33 & .32& -1
+%      \end{array}\right)
+%    \end{displaymath}
+%  \end{minipage}
+%\end{tabular}
+%
+%\end{frame}
+%
+%
+%
+%\begin{frame}
+%\frametitle{Dependent null hypotheses}
+%
+%\begin{itemize}
+%\item $H_i$ depends on $H_I$ iff $H_{I \cup i} = H_{I}$, induces a
+%  dependence relation $\sim_H$ on $\mathcal{P}(S)$
+%\item $H_I$ independent iff for all $i \in I$, $i \nsim_H I$,
+%\item $I$ independent $\Leftrightarrow$ $H_{I} \subsetneq
+%  H_{I\setminus i}$ for all $i \in I$
+%\pause
+%\item For pairwise comparisons $I$ independent iff there are no cycles in $\mathscr{G}_I$
+%\item For general contrast tests $I$ independent iff $\bs{C_I}$ has full
+%  column rank
+%\end{itemize} % ends low level
+%\end{frame}
+%
+%
+%
+%\section{Matroids}
+%
+%
+%
+%\begin{frame}
+%\frametitle{Matroids}
+%\begin{itemize}
+%
+%\item Matroids generalize the structures of graphs and vector spaces,
+%  \eg cycles, bases, independent sets
+%\item A matroid $M$ on $S$ endows $\mathcal{P}(S)$ with an algebraic
+%  structure
+%\item Cryptomorphism: many ways to define a matroid:
+%  \begin{itemize}
+%  \item From a graph perspective: Circuit axioms.
+%  \item From a vector space perspective: Independence axioms.
+%  \end{itemize}
+%\item There is a dual matroid $M^*$ which can be helpful for proofs and algorithms
+%\end{itemize} % ends low level
+%\end{frame}
+%
+%
+%
+%\begin{frame}
+%\frametitle{Circuit axioms}
+%
+%
+%\begin{tabular}{ll}
+%  \begin{minipage}{.4\textwidth}
+%    \only<2->{\input{cycle}}
+%.
+%  \end{minipage} &
+%  \begin{minipage}{.6\textwidth}
+%    Consider $S$ and a collection of subsets $\mathscr{C}$ for which holds
+%    \begin{enumerate}
+%    \item $C_1 \neq C_2 \in \mathscr{C}$, then $C_1 \nsubseteq C_2$,
+%    \item $C_1 \neq C_2 \in \mathscr{C}$, and $i \in C_1 \cap C_2$ then
+%      $\exists C_3 \subseteq (C_1 \cup C_2)\setminus i$,
+%    \end{enumerate}
+%    then $\mathscr{C}$ is the set of circuits of a matroid $M$ on $S$
+%  \end{minipage}
+%\end{tabular}
+%
+%\end{frame}
+%
+%% \item \textbf{/Some illustration hereabouts/}
+%
+%% \end{itemize} % ends low level
+%% \end{frame}
+%
+%
+%\begin{frame}
+%\frametitle{Independence axioms}
+%
+%\begin{tabular}{ll}
+%  \begin{minipage}{.4\textwidth}
+%  \only<1>{\input{independent}}
+%  \only<2>{\input{cycle}}
+%  \invisible<1>{\alert{\footnotesize minimally dependent set!}}
+%  \end{minipage} &
+%  \begin{minipage}{.6\textwidth}
+%  Consider $S$ and a collection of subsets $\mathscr{I}$ for which
+%  holds
+%  \begin{enumerate}
+%  \item $\emptyset \in \mathscr{I}$,
+%  \item $I \in \mathscr{I}$ and $J \subseteq I$ then $J \in
+%    \mathscr{I}$,
+%  \item $I,J \in \mathscr{I}$, $|I| = |J + 1|$ then $\exists j \in I
+%    \setminus J$ such that $J \cup i \in \mathscr{I}$,
+%  \end{enumerate}
+%  then $\mathscr{I}$ is the set of independent sets of a matroid $M$
+%  on $S$. This defines an independence relation $i \nsim_M I$ iff
+%  $I\cap i \in \mathscr{I}$.
+%
+%  \end{minipage}
+%\end{tabular}
+%
+%
+%
+%
+%\end{frame}
+%
+%
+%\begin{frame}
+%\frametitle{Dependence axioms}
+%
+%\begin{tabular}{ll}
+%  \begin{minipage}{.4\textwidth}
+%    \input{dependent}
+%  \end{minipage} &
+%  \begin{minipage}{.6\textwidth}
+%  Relation $\sim_M$ on elements and subsets of $S$,
+%  if for any $x \in S$, and sets $Y,Z \subseteq S$ holds
+%  \begin{enumerate}
+%  \item $y \sim Y$ for all $y \in Y$
+%  \item if $x \sim_M Y$ and $x \nsim_M Y\setminus y$
+%    then $y \sim x \cup (Y \setminus y)$
+%  \item if $x \sim_M Y$ and for all $y \in Y$, $y \sim_M Z$ then $x \sim_M Z$
+%  \end{enumerate}
+%  then $\sim_M$ defines a matroid on $S$.
+%  \end{minipage}
+%\end{tabular}
+%
+%
+%
+%\end{frame}
+%
+%\section{Matroids from logically dependent hypotheses}
+%\begin{frame}
+%  \frametitle{Dependent hypotheses}
+%  \begin{block}{Equivalence of dependence relations $\sim_H = \sim_M$}
+%    \begin{itemize}
+%    \item<1-> $H_i$ depends on $H_I$ iff $H_{I \cup i} = H_{I}$
+%    \item<2-> Does this mean $\mathscr{H}$ defines a matroid $M$, such that
+%      $\sim_H = \sim_M$?
+%    \end{itemize}
+%  \end{block}
+%\end{frame}
+%
+%
+%
+%\begin{frame}
+%\frametitle{Matroids from logically dependent hypotheses}
+%
+%  \begin{block}{Two-sided pariwise comparisons}
+%    For a family of two-sided pairwise mean comparisons $H_i$, $i \in
+%    S$, an elementary hypotheses $H_j$ depends on intersection
+%    hypotheses $H_I$ iff the corresponding edge $e_j$ is dependent on
+%    the edge set $\{e_i:i \in I\}$ of the cycle matroid defined by
+%    $\mathscr{G}_I$.
+%  \end{block}
+%
+%  \begin{block}{Two-sided linear contrast tests}
+%    For a family of two-sided general contrast tests $H_i$, $i \in S$,
+%    an elementary hypothesis $H_j$ depends on intersection hypotheses
+%    $H_I$ iff the corresponding contrast vector $\bs{c_j}$ is
+%    dependent on the column vectors of the matrix $\bs{C}_I$ in the
+%    vectorial matroid defined by $\bs{C}$ on $\mathbb{R}^m$.
+%  \end{block}
+%
+%\end{frame}
+%
+%
+%\begin{frame}
+%  \frametitle{Matroids from null hypotheses}
+%  {\Large What is it good for?\\
+%  \invisible<1>{Structure}}
+%\end{frame}
+%
+%\begin{frame}
+%\frametitle{Maximally independent sets}
+%  \begin{itemize}
+%  \item Bases are maximally independent sets of $S$
+%  \item Rank $\rho I$ is the cardinality of a maximally indepedendent set in $I$
+%  \item Circuits are minimally dependent sets, \ie each subset of $C$
+%    with rank $\rho C - 1$ is a maximally independent set of $C$
+%  \end{itemize}
+%\end{frame}
+%
+%
+%
+%\begin{frame}
+%\frametitle{Maximally dependent sets}
+%\begin{itemize}
+%\item The closure of a set $\sigma I = \{i \in S: i \sim I\}$
+%\item Closed sets {\em flats}: $\mathscr{L} = \{\sigma I: I \subseteq S\}$
+%\item Hyperplanes are flats of rank $\rho S - 1$
+%\item All flats are given by closing hyperplanes under intersection
+%\item Circuits in $M^*$ are the complements of hyperplanes in $M$
+%\end{itemize} % ends low level
+%\end{frame}
+%
+%
+%
+%\begin{frame}{Examples}
+%\framesubtitle<1>{Independent set}
+%\framesubtitle<2>{Base}
+%\framesubtitle<4>{Dependent set}
+%\framesubtitle<3>{Cycle}
+%\framesubtitle<5>{Closure}
+%\only<1>{\input{independent}}
+%\only<2>{\input{base}}
+%\only<4>{\input{dependent}}
+%\only<3>{\input{cycle}}
+%\only<5>{\input{closure}}
+%\end{frame}
+%
+%
+%
+%
+%
+%
+%
+%
+%\begin{frame}
+%\frametitle{Flats}
+%\begin{itemize}
+%
+%\item The flats of a matroid are closed under intersection
+%
+%\item Every elementary hypothesis is a flat
+%\item For each $I \subseteq S$, $H_I = H_{\sigma I}$
+%\item Define a stepwise procedure on the flats of $M$
+%
+%
+%\end{itemize} % ends low level
+%\end{frame}
+%
+%\begin{frame}
+%  \frametitle{Stepwise test procedure}
+%  \begin{block}{Multiple testing procedure for logically related hypotheses}
+%    Assume that $\mathscr{H}$ with dependence relation $\sim_H$
+%    defines a matroid on $M$ on $S$. Then a test procedure that
+%    rejects elementary hypotheses $H_i$ if all $H_E$, $i \in E$, $E
+%    \in \mathscr{L}$ are rejected using a local level-$\alpha$ test
+%    controls the FWER in the strong sense.
+%  \end{block}
+%
+%\end{frame}
+%
+%\begin{frame}
+%  \frametitle{Stepwise multiple testing procedure on the lattice of
+%    flats}
+%%
+%\begin{tabular}{ll}
+%  \begin{minipage}{.3\textwidth}
+%    \includegraphics<1>[width=1.2\textwidth]{lattice4.pdf}
+%    \includegraphics<2>[width=1.2\textwidth]{lattice4accept.pdf}
+%    \includegraphics<3>[width=1.2\textwidth]{lattice4reject.pdf}
+%    \includegraphics<4>[width=1.2\textwidth]{lattice4rejacc.pdf}\\
+%    {\tiny wikipedia:Partition\_of\_a\_Set}
+%  \end{minipage} &
+%  \begin{minipage}{.7\textwidth}
+%  \begin{itemize}
+%  \item All flats are given by intersections of hyperplanes
+%    $\rightarrow$ knowing all hyperplanes remaining flats are given by
+%    intersection of index sets
+%  \item Hyperplanes can by found by algorithms that enumerate the
+%    circuits of the dual matroid \cite{boros2003algorithms}
+%  \item The partial order of flats defined by inclusion defines a
+%    geometric lattice
+%  \item<2-> Testing only a subset of elementary hypotheses corresponds
+%    to intervals
+%  \item<3-> Rejection of elementary hypotheses removes intervals
+%  \end{itemize}
+%
+%  \end{minipage}
+%\end{tabular}
+%
+%\end{frame}
+%
+%\begin{frame}{Matroid procedures for pairwise mean comparisons}
+%\begin{tabular}{ll}
+%  \begin{minipage}{.3\textwidth}
+%    \includegraphics[width=1.1\textwidth]{setpartitions.pdf} \\
+%     {\tiny wikipedia:Bell\_number}
+%  \end{minipage} &
+%  \begin{minipage}{.7\textwidth}
+%  \begin{itemize}
+%  \item Two-sided comparisons of $m$ means
+%  \item The cycle matroid of the graph representation
+%    provides equivalent dependence structure
+%  \item The flats of a (simple) complete graph are
+%    isomporphic to the partitions of a set of size $m$
+%  \item Efficient algorithms to generate all exhaustive sets by
+%    enumeration of set partitions \cite{er1988fast,knuth2005art,kokosinski2006new}
+%  \item The number of flats are $m$-th Bell-Number $-1$ \ie 4, 14,
+%    51, 202, 876, 4139, ...
+%  \end{itemize}
+%  \end{minipage}
+%\end{tabular}
+%
+%\end{frame}
+%
+%
+%
+%
+%\begin{frame}
+%\frametitle{Connections to existing procedures}
+%
+%\textbf{Old wine in new bottles ((c) L. Hothorn)}
+%\begin{itemize}
+%\item \cite{shaffer1986modified} represents pairwise comparisons
+%  by set partitions
+%\item \cite{bergmann1988improvements} define the concepts of an
+%  exhaustive sets
+%\item \cite{bernhard1991computergestuetzte} shows properties of
+%  exhaustive sets that are basically {\em closure axioms} of a matroid
+%  and provides algorithms to check each intersection hypotheses if is
+%  a flat using rank function.
+%\item \cite{westfall1997multiple} generates flats using dependence
+%  arguments in vector spaces.
+%\item \cite{weichert2000robuste} suggest an MTP based on clique
+%  decomposition of the graph, cliques in the graph are flats
+%\item \cite{westfall2007multiple} provide a branch and bound algorithm
+%  to generate all flats. They note that for pairwise comparisons all
+%  exhaustive hypotheses are {\em partition hypotheses}
+%\end{itemize}
+%\end{frame}
+%
+%
+%\begin{frame}
+%\frametitle{Matroid representation}
+%  \begin{itemize}
+%  \item proofs are shorter (more elegant),
+%  \item efficient algorithms exist to find the flats of matroids,
+%  \item {\em Scum principle}\cite{welsh2010matroid} everything
+%    important happens at the top (hyperplanes)
+%  \item we may define procedures using independent sets,
+%  \item \ldots{}
+%  \end{itemize}
+%\end{frame}
+%
+%\section{Graphical approaches for logically dependent hypotheses}
+%
+%
+%
+%\begin{frame}
+%\frametitle{Reminder: Graphical approaches}
+%
+%\begin{tabular}{ll}
+%  \begin{minipage}{.4\textwidth}
+%    \begin{tikzpicture}[remember picture,overlay]
+%      \node[yshift=-2.5cm,xshift=-1.4cm] at (current page.north west){\input{qvatest}};
+%    \end{tikzpicture}
+%
+%  \end{minipage} &
+%  \begin{minipage}{.6\textwidth}
+%    \begin{itemize}
+%    \item Provides weights $w_{i,I}$ for each index $i$ and each
+%      non-empty subset $I \subset S$
+%    \item Weights are used to define a closed test procedure of
+%      weighted intersection hypotheses tests
+%    \item We have $\sum_{i \in S} w_{i,I} \leq 1$ (conservative),
+%    \item and $w_{i,I} \leq w_{i,J}$ for $J \subseteq I$ (consonance).
+%    \end{itemize}
+%  \end{minipage}
+%\end{tabular}
+%\end{frame}
+%
+%
+%\begin{frame}
+%  \frametitle{Example: Hierarchical LSD}
+%\begin{tabular}{ll}
+%  \begin{minipage}{.4\textwidth}
+%    \input{lsd2.tex}
+%  \end{minipage} &
+%  \begin{minipage}{.6\textwidth}
+%    \begin{itemize}
+%    \item 3 Hypotheses:
+%      \begin{enumerate}
+%      \item Placebo vs. High Dose
+%      \item Placebo vs. Low Dose
+%      \item High Dose vs. Low Dose
+%      \end{enumerate}
+%    \item Prefixed sequence: $H_1 \rightarrow H_2 \rightarrow H_3$
+%    \item Test $H_1$ at level $\alpha$
+%    \item If we reject $H_1$ test $H_2$ and $H_3$ at level $\alpha$
+%      simultaneously
+%    \item Can we generalize this?
+%    \end{itemize}
+%  \end{minipage}
+%\end{tabular}
+%
+%\end{frame}
+%
+%\begin{frame}
+%  \frametitle{Fixed sequence test of pairwise mean comparisons}
+%\begin{itemize}
+%\item {\bf Picture of hierarchical test hereabouts}
+%\end{itemize}
+%
+%\begin{block}{Theorem}
+%  Assume that $M(\mathscr{H})$ neither contains loops nor
+%  parallels. Assume that hypotheses $H_1$ through $H_r$ have been
+%  rejected. Consider that hypotheses $H_{r+1}$ through $H_{r+t}$ are
+%  to be tested at level $\alpha_r$ according to a prefixed order.  If
+%  for some $1 \leq s \leq t$ holds
+%  \begin{equation}
+%    \label{eq:cond.hierarchical}
+%    \forall 1 \leq l < k \leq s: \; \exists i \leq r: i \sim \{(r+l),(r+k)\},
+%  \end{equation}
+%  hypotheses $H_{r+1}$ through $H_{r+s}$ can be tested simultaneously
+%  at full level $\alpha_ra$.
+%\end{block}
+%
+%\end{frame}
+%
+%\begin{frame}
+%  \frametitle{Example: Multiple doses}
+%
+%
+%
+%\begin{tabular}{ll}
+%  \begin{minipage}{.4\textwidth}
+%    \begin{tikzpicture}[overlay]
+%      \node[yshift=70bp,xshift=-40bp] at (current page.south west){\input{lsd4.tex}};
+%    \end{tikzpicture}
+%  \end{minipage} &
+%  \begin{minipage}{.6\textwidth}
+%    \begin{tikzpicture}[remember picture,overlay]
+%      \node[yshift=100bp,xshift=160bp] at (current page.north west){\input{lsd4t.tex}};
+%    \end{tikzpicture}
+%  \end{minipage}
+%\end{tabular}
+%\end{frame}
+%
+%\begin{frame}
+%  \frametitle{General weighting strategies}
+%  \framesubtitle{QVA Example}
+%
+%
+%\begin{tabular}{ll}
+%  \begin{minipage}{.4\textwidth}
+%    \begin{tikzpicture}[remember picture,overlay]
+%      \node[yshift=-2.5cm,xshift=-1.4cm] at (current page.north west){\input{qvatest}};
+%    \end{tikzpicture}
+%
+%  \end{minipage} &
+%  \begin{minipage}{.6\textwidth}
+%    \begin{tikzpicture}[remember picture,overlay]
+%      \node[yshift=-50bp,xshift=-180bp] at (current page.north east){\input{qvameans}};
+%  \end{tikzpicture}
+%  \end{minipage}
+%\end{tabular}
+%\end{frame}
+%
+%\begin{frame}
+%\frametitle{General weighting strategies}
+%
+%\begin{enumerate}
+%\item Pretest using weighted Bonferroni or conventional graphical
+%  approach $R = \{i \in S: H_i \textrm{ rejected by pretest}\}$
+%\item Compute all matroid hyperplanes $\mathscr{H}$ for the matroid on
+%  the remaining hypotheses $I$
+%\item $R \gets R \cup \{i: p_i \leq \min_J w_{i,J} \alpha, J \in \mathscr{H}\}$
+%\item $A \gets \{\bigcup J: \min_j (p_j / w_{j,J}) > \alpha\}$
+%\item $I \gets I \setminus R$
+%\item Refine: $\mathscr{H} \gets \{J\in \mathscr{L}: \rho J = \rho
+%  \mathscr{H} - 1\}$
+%\item Prune: $\mathscr{H} \gets \{J \in \mathscr{H}: J \cap R = \emptyset\}$
+%\item Repeat until $\rho \mathscr {H} = 0$
+%\item Reject $R$ and $I \setminus A$
+%\end{enumerate}
+%\end{frame}
+%\section{Matriods from general families of hypotheses}
+%
+%
+%
+%\begin{frame}
+%\frametitle{General Results}
+%\begin{itemize}
+%\item We give sufficient (and probably necessary) conditions for a
+%  general family of hypotheses $\mathscr{H}$ to define a matroid
+%\item We give sufficient (and probably necessary) conditions on
+%  $\mathscr{H}$ for $\mathscr{L}$ to be the set of strictly exhaustive
+%  sets
+%\item Systems with one sided hypotheses, and parallel hypotheses (\eg
+%  combined non-inferiority/superiority) do not give matroids with
+%  $\mathscr{L}$ all strictly exhaustive
+%\end{itemize}
+%
+%
+%\end{frame}
+%
+%\begin{frame}
+%  \frametitle{One sided hypotheses}
+%  \begin{block}{One-sided linear contrast hypotheses}
+%    A family of {\em one sided linear contrast hypotheses} is defined
+%    by a contrast matrix $\bs{A} \in \mathbb{R}^{m\times n}$ and
+%    vector $\bs{b} \in \mathbb{R}^n$ that define a  family of null
+%    hypotheses of the form $\mathscr{H} = \{H_i: \bs{A_i}^t\bs{\mu}
+%    \leq b_i,i \in \{1,...,n\}\}$, where $\bs{A}_i$ denotes the $i$-th
+%    column of $\bs{A}$ and $b_i$ the $i$-th componenth of $\bs{b}$.
+%  \end{block}
+%\end{frame}
+%
+%\begin{frame}
+%  \frametitle{Hyperplane arrangement}
+%  \begin{itemize}
+%  \item Each hypothesis $H_i$ corresponds to a hyperplane $h_i =
+%    \{\bs{x} \in \mathbb{R}^m: \bs{A}_i^t\bs{x} = b_i\}$
+%  \item Each hyperplane $h_i$ divides $\mathbb{R}^m$ into two regions
+%    \begin{itemize}
+%    \item $h_i^+ = \{\bs{x} \in \mathbb{R}^m: \bs{A}_i^t\bs{x} <
+%      b_i\}$
+%    \item $h_i^- = \{\bs{x} \in \mathbb{R}^m: \bs{A}_i^t\bs{x}
+%      > b_i\}$
+%    \end{itemize}
+%  \item Defines a sign pattern
+%    \begin{equation}
+%      \label{eq:sign}
+%      \delta_i(\bs{x}) = \left\{
+%        \begin{array}{ll}
+%          + & \text{if } \bs{x} \in h_i^+\\
+%          0 & \text{if } \bs{x} \in h_i \\
+%          - & \text{if } \bs{x} \in h_i^-\\
+%        \end{array}\right.
+%    \end{equation}
+%  \item $\delta_i$ is $-$ if $H_i$ false at $x$, or $0$ or $+$
+%    otherwise. $(\delta_1,...,\delta_n)$ tells us to which combination
+%    of true and false null hypotheses some $\bs{x}$ belongs.
+%
+%  \item Set of all sign patterns defined by $\mathscr{H}$:
+%    \begin{equation}
+%      \label{eq:faces}
+%      \mathcal{F}(\mathscr{H}) = \{\bs{\delta}(\bs{x}): \bs{x} \in \R^m\}
+%    \end{equation}
+%  \end{itemize}
+%\end{frame}
+%
+%\begin{frame}
+%  \frametitle{Example}
+%
+%\begin{tabular}{ll}
+%  \begin{minipage}{.4\textwidth}
+%  \includegraphics[height=1\textwidth]{spheresfukuda.eps}
+%  \end{minipage} &
+%  \begin{minipage}{.6\textwidth}
+%    \begin{itemize}
+%    \item 4 contrast hypotheses about 3 means
+%    \item Cut the hyperplanes with the sphere in $\R^3$ (makes
+%      visualisation easier)
+%    \item Cells of the sphere arrangment correspond to sign patterns
+%      of $\mathscr{H}$
+%    \item The abstraction of such an arrangement is an \alert{Oriented Matroid}
+%    \end{itemize}
+%  \end{minipage}
+%\end{tabular}
+%
+%\end{frame}
+%
+%\begin{frame}
+%  \frametitle{Stepwise multiple testing procedure for one-sided linear
+%    contrast hypotheses}
+%
+%  Consider a family of general contrast hypotheses $\mathscr{H} =
+%  \{H_i: \bs{A_i}^t\bs{\mu} \leq b_i,i \in \{1,...,n\}\}$ and let
+%  $\mathscr{F}(\mathscr{H}) = \{ \bs{\delta}(\bs{x}): \bs{x} \in
+%  \R^m\}$ denote the corresponding set of sign patterns. Consider a
+%  set of decision rules $\varphi_F: \R^n \rightarrow \{0,1\}$ for
+%  which hold $P(\varphi_F = 1 | H_i: F_i \geq 0) \leq \alpha$. Then a
+%  corresponding decision rule for each elementary hypotheses $H_i$ is
+%  given by
+%  \begin{equation}
+%    \label{eq:signtest}
+%    \psi_i = \min_{F: F_i \geq 0} \varphi_F,
+%  \end{equation}
+%  which controls the FWER at level $\alpha$ in the strong sense.
+%
+%
+%\end{frame}
+%
+%
+%\begin{frame}
+%  \frametitle{Computation of $\mathscr{F}$}
+%  The question remains how to compute $\mathcal{F}$. For this observe
+%that the $h_i$ define a hyperplane configuration and $\mathcal{F}$ is
+%just the face lattice of the corresponding oriented
+%matroid. Consequently, $\mathcal{F}$ can be found using algorithms
+%for arrangement construction from polyhedral programming
+%\cite{Avis-KF-92,Avis-KF-96,Ferrez-KF-Liebling-01} which are
+%polynomial in time and input size and compact in space!
+%
+%\end{frame}
+%
+%\begin{frame}
+%  \frametitle{Example}
+%  3 ordered means
+%\end{frame}
+%
+%\section{Summary \& Outlook}
+%
+%
+%
+%\begin{frame}
+%\frametitle{Summary}
+%\begin{itemize}
+%\item Matroids provide usefull abstraction for systems of algebraically dependent hypotheses
+%\end{itemize}
+%\end{frame}
+%
+%
+%
+%\begin{frame}
+%\frametitle{Outlook}
+%\begin{itemize}
+%\item MCPmod on \sout{steroids} matroids
+%\end{itemize}
+%\end{frame}
+%
+%\bibliographystyle{alpha}
+%\bibliography{\parentd paper/cleanrefs}
 \end{document}

### 2013-08-19 presentation/presentation.tex
cleaned up the first couple of slides

--- 2013-08-19 presentation/presentation.tex
@@
 \newcommand{\bs}[1]{\bm{#1}}
-\newcommand{\ie}{{\em i.e.},~}
-\newcommand{\eg}{{\em e.g.},~}
+\newcommand{\ie}{{\sl i.e.},~}
+\newcommand{\eg}{{\sl e.g.},~}
 \newcommand{\parentd}{../}
@@
 \doNotUseNovartisLogoOnTitlePage
+\setbeamertemplate{blocks}[rounded][shadow=true]
 
-\title{Matroids from families of null hypotheses}
+\title{Matroids \& testing in families of dependent hypotheses}
 \author{Florian Klinglmueller, Willi Maurer}
@@
 
-\section{Logically dependent hypotheses}
+\section{Algebraically dependent hypotheses}
 
 \begin{frame}
-\frametitle{Pairwise comparisons, some history }
+\frametitle{Some history:}
+\framesubtitle{Testing hypotheses of no differences in pairwise
+  comparisons of 3 treatments}
 
-Testing hypotheses of no difference in pairwise comparisons of 3 treatments (categories):
+% Testing hypotheses of no difference in pairwise comparisons of 3 treatments (categories):
+
+\begin{block}{$H_1:\mu_1=\mu_2,\; H_2:\mu_2=\mu_3,\;H_3:\mu_1=\mu_3$}
+  \begin{enumerate}
+  \item Test $H_ {\{1,2,3\}} = H_1\cap H_2 \cap H_3 :
+    \mu_1=\mu_2=\mu_3$ at level $\alpha$,
+  \item test each $H_1,H_2,H_3$ at level $\alpha$,
+  \end{enumerate}
+  protects the FWER at level $\alpha$.
+\end{block}
+\pause
   \begin{itemize}
-  \item Parameters: $\mu_1,\mu_2,\mu_3$ ; \\ Hypotheses: $H_1:\mu_1=\mu_2,H_2:\mu_2=\mu_3;H_3:\mu_1=\mu_3$
-  \item A level-$\alpha$ test of $H_ {\{1,2,3\}} = H_1\cap H_2 \cap H_3 : \mu_1=\mu_2=\mu_3$. followed by level-$\alpha$ tests of $H_1,H_2,H_3$ protects the FWER at level $\alpha$.
-  \begin{itemize}
-    \item  essentially closed test; no need to test pairwise intersection hypotheses since $H_{{i,j}} = H_{\{1,2,3\}}, i\neq j$,
+    \item  essentially closed test, since %; no need to test pairwise
+                                %intersection hypotheses since 
+        $H_{{i,j}} = H_{\{1,2,3\}}, i\neq j,$
+
     \item  Bonferroni: if one of the hypotheses is rejected at level $\alpha/3$ the others can be tested at level alpha.
@@
   \end{itemize}
-\end{itemize}
+
 \end{frame}
@@
 \begin{frame}
-\frametitle{Pairwise comparisons, some history}
+\frametitle{Some history}
+\framesubtitle{Testing hypotheses of no difference in pairwise
+    comparisons of $m>3$ treatments:}
 
-Testing hypotheses of no difference in pairwise comparisons of $m>3$ treatments (categories):\\
-\cite{shaffer1986modified}: Modified Sequentially Rejective Multiple test procedures; Example: Modified Bonferroni-Holm
+  Modified Sequentially Rejective
+  Multiple test procedures \cite{shaffer1986modified} % Example: Modified Bonferroni-Holm
       \begin{itemize}
       \item "Static":  After j rejections, test remaining hypotheses at level $\alpha/t_j$, $t_j$ = max number of true hypotheses if $j$ are false.
-  \item "Dynamic": After rejecting $H_{i_1},...,H_{i_j}$, test remaining hypotheses at level $\alpha/t^*_j$,  $t^*_j$ = number of true hypotheses if $H_{i_1},...,H_{i_j}$ are false.
-  \item Example: $m=4, n=6$ :
-    \begin{itemize}
-      \item Static: $t_0=6, t_1,t_2,t_3 = 3, t_4=2, t_5=1$
-      \item Dynamic: if, e.g., $H_{\{1,2\}}$ and $H_{\{3,4\}}$ are rejected then only $H_{\{1,3\}}$ and $H_{\{2,3\}}$, or $H_{\{1,4\}}$ and $H_{\{2,3\}}$ can be simultaneously true, hence $t^*_2 = 2 < t_2 =3$ in this case.
+      \item "Dynamic": After rejecting $H_{i_1},...,H_{i_j}$, test
+        remaining hypotheses at level $\alpha/t^*_j$,  $t^*_j$ =
+        max. number of true hypotheses if $H_{i_1},...,H_{i_j}$ are
+        false.
       \end{itemize}
-\end{itemize}
+
 \end{frame}
@@
 \frametitle{Pairwise mean comparisons (two-sided)}
+\begin{tabular}{ll}
+
+  \begin{minipage}{.5\textwidth}
+    \input{full4}
+  \end{minipage} 
+  \begin{minipage}{.5\textwidth}
+    \begin{block}{Graphical representation}
       \begin{itemize}
+      \item $H_k: \mu_i = \mu_j, k \in S = \{1,...,n\}$, where $n \leq
+        m*(m-1)/2$
 
-\item Consider $\mu_i, i \in \{1,...,m\}$
-
-\item $H_k: \mu_i = \mu_j, k \in S = \{1,...,n\}$, where $n \leq m*(m-1)/2$
-
-\item Can be represented by an undirected graph $\mathscr{G} = (V,E)$:
+      \item Undirected graph %$\mathscr{G} = (V,E)$:
         \begin{itemize}
-\item vertices correspond to means, \emph{i.e.}, $V = \{\mu_1,...,\mu_m\}$,
-\item edges correspond to hypotheses, \emph{i.e.}, $E =
-      \left\{e_1,...,e_n\} \right\}$, $e_l = \{\mu_i,\mu_j\}$, $i \neq
-      j$, and $e_l \neq e_k$ for $l \neq k$
+        \item vertices correspond to means, 
+          %\emph{i.e.}, $V = \{\mu_1,...,\mu_m\}$,
+        \item edges correspond to hypotheses, 
+          % \emph{i.e.}, $E = \left\{e_1,...,e_n\} \right\}$, $e_\ell = \{\mu_i,\mu_j\}$, $i
+          % \neq j$, and $e_\ell \neq e_k$ for $\ell \neq k$
         \end{itemize}
       \item Let $H_I = \bigcap_{k \in I} H_k$, $I \subseteq S$
-\item $I$ induces a subgraph $\mathscr{G}_I$ on the edge set $E_I=\{e_i: i
-  \in I\}$
+      \item $I$ induces a subgraph %$\mathscr{G}_I$ 
+        on the edge set
+        %$E_I=\{e_i: i \in I\}$
       \end{itemize} % ends low level
+    \end{block}
+  \end{minipage}
+
+\end{tabular}
 \end{frame}
 
+% \begin{block}{Example: $m=4, n=6$ :}
+%       \begin{itemize}
+%       \item Static: $t_0=6, t_1,t_2,t_3 = 3, t_4=2, t_5=1$
+%       \item Dynamic: $t^*_j \le t_j$, hence more powerful than
+%         "static"
+%       \end{itemize}
+%   \end{block}
+
 \begin{frame}
-\frametitle{Example 1: 4 Groups}
-\input{full4ext}
+\frametitle{Example: Modified Bonferroni Holm \cite{shaffer1986modified}}
+\begin{tabular}{ll}
+  \begin{minipage}{.4\textwidth}
+
+    \input{shaffer}
+\end{minipage} &
+  \begin{minipage}{.6\textwidth}
+    \begin{block}{$m=4, n=6, \only<2->{j=2}$}
+      % \begin{itemize}
+      % \item "Static": $t_j$ = max number of true hypotheses if $j$ are
+      %   false.
+      % \item "Dynamic": $t^*_j $ = max. number of true hypotheses if
+      %   $H_{i_1},...,H_{i_j}$ are false.
+%      \item Example:  \only<3->{$ j = 2$ :}
+        \begin{description}
+        \item<3->[Static:]  $H_1$ and $H_2$ are false
+            $\Rightarrow$ $t_2 = 3$ (at most $H_4, H_5 $ and $H_6$ can be
+            simultaneously true)
+        \item<7->[Dynamic:] $H_1$ and $H_6$ are false $\Rightarrow$
+          $t^*_2 = 2 < t_2 =3$ (only $H_3$, $H_4$, or $H_2$, $H_5$ can be
+            simultaneously true)
+        \end{description}
+
+ %     \end{itemize} % ends low level
+    \end{block}
+  \end{minipage}
+\end{tabular}
 \end{frame}
+
+% \begin{frame}
+% \frametitle{Pairwise mean comparisons (two-sided)}
+% \begin{itemize}
+
+% \item Consider $\mu_i, i \in \{1,...,m\}$
+
+% \item $H_k: \mu_i = \mu_j, k \in S = \{1,...,n\}$, where $n \leq m*(m-1)/2$
+
+% \item Can be represented by an undirected graph $\mathscr{G} = (V,E)$:
+% \begin{itemize}
+% \item vertices correspond to means, \emph{i.e.}, $V = \{\mu_1,...,\mu_m\}$,
+% \item edges correspond to hypotheses, \emph{i.e.}, $E =
+%      \left\{e_1,...,e_n\} \right\}$, $e_l = \{\mu_i,\mu_j\}$, $i \neq
+%      j$, and $e_l \neq e_k$ for $l \neq k$
+% \end{itemize}
+% \item Let $H_I = \bigcap_{k \in I} H_k$, $I \subseteq S$
+% \item $I$ induces a subgraph $\mathscr{G}_I$ on the edge set $E_I=\{e_i: i
+%  \in I\}$
+% \end{itemize} % ends low level
+% \end{frame}
+
+% \begin{frame}
+% \frametitle{Example 1: 4 Groups}
+% \input{full4ext}
+% \end{frame}
 
@@
 \begin{tabular}{ll}
-  \begin{minipage}{.45\textwidth}
-    All pairwise comparisons of 4 means 
+ \begin{minipage}{.44\textwidth}
+   \begin{block}{\footnotesize Pairwise comparisons of 4 means}
+     {\footnotesize
        \begin{displaymath}
          \bs{C}^t = \left(
-      \begin{array}{llll}
+           \begin{array}{rrrr}
              -1 & 1 & 0 & 0 \\
@@
          \right)
-    \end{displaymath}
+       \end{displaymath}}
+   \end{block}
  \end{minipage} &
-  \begin{minipage}{.55\textwidth}
-    Modified Williams Contrasts for Dose Response
+ \begin{minipage}{.56\textwidth}
+\pause
+   \begin{block}{\footnotesize Modified Williams Contrasts for Dose Response \cite{bretz2010multiple}}
+      {\footnotesize
        \begin{displaymath}
          \bs{C}^t=\left(
-      \begin{array}{llll}
+           \begin{array}{rrrr}
              1 & 0 & 0 & -1\\
@@
            \end{array}\right)
-    \end{displaymath}
+       \end{displaymath}}
+   \end{block}
  \end{minipage}
@@
 \begin{itemize}
-\item $H_i$ depends on $H_I$ iff $H_{I \cup i} = H_{I}$, induces a
-  dependence relation $\sim_H$ on $\mathcal{P}(S)$
-\item $H_I$ independent iff for all $i \in I$, $i \nsim_H I$,
-\item $I$ independent $\Leftrightarrow$ $H_{I} \subsetneq
-  H_{I\setminus i}$ for all $i \in I$
+\item $H_i$ depends on $H_I$: $H_{I \cup i} = H_{I}$
+\item induces a dependence relation $i \sim_H I$ on index sets $I \in \mathcal{P}(S)$
+\item $I$ independent iff for all $i
+  \in I$ $i \nsim_H I\setminus i$ (\ie $H_{I} \subsetneq H_{I\setminus i}$)
 \pause
@@
  structure
-\item Cryptomorphism: many ways to define a matroid: 
+\item {\em Cryptomorphism} \ie many ways to define a matroid:
  \begin{itemize}
@@
  \end{itemize}
-\item There is a dual matroid $M^*$ which can be helpful for proofs and algorithms 
+%\item There is a dual matroid $M^*$ which can be helpful for proofs and algorithms
 \end{itemize} % ends low level
@@
 \begin{tabular}{ll}
-  \begin{minipage}{.4\textwidth}
-    \only<2->{\input{cycle}}
-.
-  \end{minipage} &
-  \begin{minipage}{.6\textwidth}
-    Consider $S$ and a collection of subsets $\mathscr{C}$ for which holds
+ \begin{minipage}{.5\textwidth}
+   \begin{block}{Circuits of a matroid $M$ on $S$:}
+       A collection $\mathscr{C}$ of subsets of $S$
+       for which holds
      \begin{enumerate}
      \item $C_1 \neq C_2 \in \mathscr{C}$, then $C_1 \nsubseteq C_2$,
-    \item $C_1 \neq C_2 \in \mathscr{C}$, and $i \in C_1 \cap C_2$ then
-      $\exists C_3 \subseteq (C_1 \cup C_2)$,
+     \item $C_1 \neq C_2 \in \mathscr{C}$, and $i \in C_1 \cap C_2$
+       then $\exists C_3 \subseteq (C_1 \cup C_2)\setminus i$,
      \end{enumerate}
-    then $\mathscr{C}$ is the set of circuits of a matroid $M$ on $S$
+   \end{block}
+
+ \end{minipage} &
+
+ \begin{minipage}{.5\textwidth}
+   \only<2->{\input{cycle}}
  \end{minipage} 
@@
 
-% \item \textbf{/Some illustration hereabouts/}
-
-% \end{itemize} % ends low level
-% \end{frame}
-
 \begin{frame}
@@
 \begin{tabular}{ll}
-  \begin{minipage}{.4\textwidth}
-  \only<1>{\input{independent}}
-  \only<2>{\input{cycle}}
-  \invisible<1>{\alert{\footnotesize minimally dependent set!}}
-  \end{minipage} &
-  \begin{minipage}{.6\textwidth}
-  Consider $S$ and a collection of subsets $\mathscr{I}$ for which
+ \begin{minipage}{.5\textwidth}
+   \begin{block}{Independent sets of a matroid $M$ on $S$}
+     A a collection $\mathscr{I}$ of subsets of $S$ for which
      holds
@@
        \mathscr{I}$,
-  \item $I,J \in \mathscr{I}$, $|I| = |J + 1|$ then $\exists j \in I
-    \setminus J$ such that $J \cup i \in \mathscr{I}$,
+     \item $I,J \in \mathscr{I}$, $|I| = |J + 1|$ then $\exists j \in
+       I \setminus J$ such that $J \cup i \in \mathscr{I}$,
      \end{enumerate}
-  then $\mathscr{I}$ is the set of independent sets of a matroid $M$
-  on $S$. This defines an independence relation $i \nsim_M I$ iff
-  $I\cap i \in \mathscr{I}$.
+   \end{block}
+ \end{minipage} &
 
+ \begin{minipage}{.5\textwidth}
+ \only<1>{\input{independent}}
+ \only<2>{\input{cycle}}
+ \invisible<1>{\alert{\footnotesize minimally dependent set!}}
  \end{minipage} 
@@
 
+\end{frame}
+
+\begin{frame}
+  \frametitle{}
+  \begin{itemize}
+  \item Circuits are minimally dependent sets (remove any element to
+    get an independent set)
+  \item Independent sets are sets for which no subset of elements is a
+    circuit
+  \item Maximally independent sets are bases (Base axioms)
+  \item Cardinality of largest independent subset is the rank of a set
+    (Rank axioms)
+  \item Maximally dependent sets are called flats (Closure axioms)
+  \end{itemize}
 \end{frame}
@@
 
-\section{Matroids from logically dependent hypotheses}
+\section{Matroids from algebraically dependent hypotheses}
 \begin{frame}
@@
 \begin{frame}
-\frametitle{Matroids from logically dependent hypotheses}
+\frametitle{Matroids from algebraically dependent hypotheses}
 
@@
  \frametitle{Stepwise test procedure}
-  \begin{block}{Multiple testing procedure for logically related hypotheses}
+ \begin{block}{Multiple testing procedure for algebraically related hypotheses}
    Assume that $\mathscr{H}$ with dependence relation $\sim_H$
@@
 
-\section{Graphical approaches for logically dependent hypotheses}
+\section{Graphical approaches for algebraically dependent and
+  contextually related hypotheses}
 
@@
  \begin{minipage}{.4\textwidth}
-  \includegraphics[height=1\textwidth]{spheresfukuda.eps}
+ \includegraphics[height=1\textwidth]{spheresfukuda.pdf}
  \end{minipage} &
@@
 \item Matroids provide usefull abstraction for systems of algebraically dependent hypotheses
+\item Efficient algorithms from the realms of combinatorics and algebraic geometry
+\item {\em Scum principle} everything important happens at the top
+\item Stepwise procedure for algebraically dependent and contextually
+  related hypotheses (graphical approach)
+\item Promising results also for one-sided hypotheses
 \end{itemize}
@@
 \begin{itemize}
-\item MCPmod on \sout{steroids} matroids
+\item Inclusion of procedures for two-sided tests in \texttt{gMCP} 
+\item Implementation of one-sided problems using polyhedral
+  programming libraries
+\item Potential applications:
+  \begin{itemize}
+  \item \texttt{MCPmod} on \sout{steroids} matroids
+  \item Model selection
+  \item Logic regression
+  \end{itemize}
+\item Tests for intersection hypotheses based on maximally independent
+  sets 
 \end{itemize}

### 2013-08-20 presentation/shaffer.tex
added animated shaffer procedure

\begin{tikzpicture}[scale=.4]
\node (H1) at (025bp,-125bp)[draw,circle,fill=white] {$\mu_1$};
\node (H2) at (225bp,-125bp)[draw,circle,fill=white] {$\mu_2$};
\node (H3) at (225bp,-325bp)[draw,circle,fill=white] {$\mu_3$};
\node (H4) at (025bp,-325bp)[draw,circle,fill=white] {$\mu_4$};
\only<-2>{\draw [line width=1pt] (H1) to (125bp, -124bp) node[fill=blue!20] {$H_{1}$} to (H2);}
\only<-2>{\draw [line width=1pt] (H1) to (075bp, -175bp) node[fill=blue!20] {$H_{2}$} to (H3);}
\only<3>{\draw [line width=1pt] (H1) to (125bp, -124bp) node[fill=red!20] {$H_{1}$} to (H2);}
\only<3>{\draw [line width=1pt] (H1) to (075bp, -175bp) node[fill=red!20] {$H_{2}$} to (H3);}
\only<-5>{\draw [line width=1pt] (H4) to (025bp, -227bp) node[fill=blue!20] {$H_{3}$} to (H1);}
\only<-4>{\draw [line width=1pt] (H2) to (225bp, -224bp) node[fill=blue!20] {$H_{4}$} to (H3);}
\only<-4>{\draw [line width=1pt] (H4) to (174bp, -176bp) node[fill=blue!20] {$H_{5}$} to (H2);}
\only<-4>{\draw [line width=1pt] (H3) to (126bp, -324bp) node[fill=blue!20] {$H_{6}$} to (H4);}
\only<4,5>{\draw [line width=2pt] (H2) to (225bp, -224bp) node[fill=green!20] {$H_{4}$} to (H3);}
\only<4,5>{\draw [line width=2pt] (H4) to (174bp, -176bp) node[fill=green!20] {$H_{5}$} to (H2);}
\only<4,5>{\draw [line width=2pt] (H3) to (126bp, -324bp) node[fill=green!20] {$H_{6}$} to (H4);}

\only<6>{\draw [line width=1pt] (H1) to (125bp, -124bp) node[fill=blue!20] {$H_{1}$} to (H2);}
\only<7>{\draw [line width=1pt] (H1) to (125bp, -124bp) node[fill=red!20] {$H_{1}$} to (H2);}
\only<6>{\draw [line width=2pt] (H3) to (126bp, -324bp) node[fill=blue!20] {$H_{6}$} to (H4);}
\only<7>{\draw [line width=2pt] (H3) to (126bp, -324bp) node[fill=red!20] {$H_{6}$} to (H4);}

\only<6,7,8,9>{\draw [line width=1pt] (H1) to (075bp, -175bp) node[fill=blue!20] {$H_{2}$} to (H3);}
\only<6,7>{\draw [line width=1pt] (H4) to (025bp, -227bp) node[fill=blue!20] {$H_{3}$} to (H1);}
\only<6,7>{\draw [line width=1pt] (H2) to (225bp, -224bp) node[fill=blue!20] {$H_{4}$} to (H3);}
\only<6,7,8,9>{\draw [line width=1pt] (H4) to (174bp, -176bp) node[fill=blue!20] {$H_{5}$} to (H2);}
\only<8>{\draw [line width=2pt] (H4) to (025bp, -227bp) node[fill=green!20] {$H_{3}$} to (H1);}
\only<8>{\draw [line width=2pt] (H2) to (225bp, -224bp) node[fill=green!20] {$H_{4}$} to (H3);}
\only<9>{\draw [line width=2pt] (H4) to (025bp, -227bp) node[fill=blue!20] {$H_{3}$} to (H1);}
\only<9>{\draw [line width=2pt] (H2) to (225bp, -224bp) node[fill=blue!20] {$H_{4}$} to (H3);}

\only<10>{\draw [line width=1pt] (H1) to (075bp, -175bp) node[fill=blue!20] {$H_{2}$} to (H3);}
\only<10,11,12>{\draw [line width=1pt] (H4) to (025bp, -227bp) node[fill=blue!20] {$H_{3}$} to (H1);}
\only<10,11,12>{\draw [line width=1pt] (H2) to (225bp, -224bp) node[fill=blue!20] {$H_{4}$} to (H3);}
\only<10>{\draw [line width=1pt] (H4) to (174bp, -176bp) node[fill=blue!20] {$H_{5}$} to (H2);}
\only<11,12>{\draw [line width=2pt] (H1) to (075bp, -175bp) node[fill=green!20] {$H_{2}$} to (H3);}
\only<11,12>{\draw [line width=2pt] (H4) to (174bp, -176bp) node[fill=green!20] {$H_{5}$} to (H2);}

\end{tikzpicture}

### 2013-08-20 paper/draft.tex
added some reference, completed proof of weakly exhaustive set theorem

@@
 
-\title{Multiple testing with logical dependencies}
+\title{Matroids \& testing in families of algebraically dependent and
+  contextually related hypotheses}
 
@@
 
+History 
+\begin{itemize}
+\item \cite{shaffer1986modified} represents pairwise comparisons
+ by set partitions
+\item \cite{bergmann1988improvements} define the concepts of an
+ exhaustive sets
+\item \cite{bernhard1991computergestuetzte} shows properties of
+ exhaustive sets that are basically {\em closure axioms} of a matroid
+ and provides algorithms to check each intersection hypotheses if is
+ a flat using rank function.
+\item \cite{westfall1997multiple} generates flats using dependence
+ arguments in vector spaces.
+\item \cite{weichert2000robuste} suggest an MTP based on clique
+ decomposition of the graph, cliques in the graph are flats
+\item \cite{westfall2007multiple} provide a branch and bound algorithm
+ to generate all flats. They note that for pairwise comparisons all
+ exhaustive hypotheses are {\em partition hypotheses}
+\end{itemize}
+
 \subsection{Pairwise comparisons of population means}
 \begin{itemize}
-\item All pairwise
-\item Gold standard designs
-\item Change point analysis
+\item All pairwise \citep{shaffer1986modified,weichert2000robuste}
+\item Gold standard designs \citep{bauer1998testing,pigeot2003assessing}
+\item Change point analysis (?)
 \item General treatment control - treatment treatment situations 
@@
 \begin{itemize}
-\item Cost effectiveness (Georg)
-\item SNP Studies (Hothorn-Hothorn)
+\item Cost effectiveness \citep{haller2011cost,gutjahr2012multiple}
+\item SNP Studies (Hothorn-Hothorn) \citep{hothorn2009order}
 \item Order restriction contrasts (Williams, ...)
@@
 \section{Two sided hypotheses}
-
-\subsection{Some observations on pairwise comparisons}
-
-\begin{definition}
-  Let $X$ be a random variable taking values in $\mathbb{R}^m$ with
-  multivariate distribution $F(\bs{\mu})$ having mean vector $\bs{\mu}
-  \in \mathbb{R}^m$. Let $R \subset \{1,...,m\}^2$, such that $i <
-  j$ for all $(i,j) \in R$. For $(i,j) \in R$ $H_{ij} = \mu_i =
-  \mu_j$. We call the family of elementary hypotheses $\dot{R}
-  = \{H_{ij}: (i,j) \in R\}$ a {\bf two-sided pairwise comparison
-    problem. (tPCP)} Let $\dot{\mathcal{T}}(m)$ denote the set of all
-  possible pairwise comparison problems of $m$ means.
-\end{definition}
-
-\begin{bsp}
-  Consider $X \sim N((\mu_1,\mu_2,\mu_3,\mu_4),\sigma \mathbb{I}_4)$
-  let $R = \{(1,2),(1,3),(1,4)\}$ and $S =
-  \{(1,4),(2,4),(3,4),(1,3),(1,2),(2,3)\}$. Then the
-  corresponding tPCP's $\dot{R},\dot{S} \in \dot{\mathcal{T}}(4)$.
-\end{bsp}
-
-\begin{definition}
-  A two-sided pairwise comparison problem $\dot{S}$ defines an
-  undirected graph $G(\dot{S}) = \{M,T\}$ with Nodes $M = \{1,...,m\}$
-  and Edges $T$. We call $G(\dot{S})$ the {\bf comparigram}. We denote
-  by $V(G(\dot{S})$, $E(G(\dot{S})$ the vertex and edge sets of the
-  corresponding comparigram. If clear from the context we may just use
-  $E(\dot{S}),V(\dot{S})$. 
-\end{definition}
-
-\begin{bsp}
-  $\dot{R}$ in above Example is represented by a claw with 3 edges,
-  $\dot{S}$ is represented by the complete graph $K_4$ with 4 vertices.
-\end{bsp}
-
-\begin{definition}
-  Let $h_{lk}:  M^2 \rightarrow \dot{\mathcal{T}}(m)$,
-  $(l,k) \mapsto \{H_{ij}: (i,j) \in T\setminus \{(l,j)\}\}$ define the
-  {\bf rejection operator}. We write $h_{lk} \circ \dot{S} =
-  \dot{S}\setminus (l,k)$. Clearly   $h_{a} \circ h_{b} \circ \dot{S}
-  = h_{b} \circ h_{a} \circ \dot{S}$.
-\end{definition}
-
-\begin{definition}
-  Consider $\dot{R},\dot{S} \in \dot{\mathcal{T}}(m)$ with associated
-  comparigrams $L = G(\dot{R}), K = G(\dot{R})$ respectively. We say
-  that $\dot{R}$ {\bf isomorphic} to $\dot{S}$, $\dot{R} \cong
-  \dot{S}$ if $L$ is isomorphic to $K$, that is there exists a
-  bijection between nodes of the graphs $f:V(K) \rightarrow V(L)$, $a
-  \mapsto u$ such that for $a,b \in V(K)$, $f(a)$ is adjacent to
-  $f(b)$, if and only if $a$ and $b$ are adjacent to each other.
-\end{definition}
-
-\begin{definition}
-  Consider $\dot{S} \in \dot{\mathcal{T}}(m)$ and $h_a,h_b$, $a,b \in
-  E(S)$ we say that $h_a \cong h_b$ are {\bf equivalent rejections} if $h_a
-  \circ \dot{S} \cong h_b \circ \dot{S}$.
-\end{definition}
-
-\begin{definition}
-  Let $h = (h_1,...,h_n)$ whith $h_i \in T$ and $n \leq |T|$ define a
-  {\bf fixed rejection sequence} of length $n$.
-\end{definition}
-
-\begin{definition}
-  Consider PCP $\dot{S}$. Two fixed rejection sequences $h, e$ are
-  called {\bf equivalent} if $h_1\circ ... \circ h_i \circ \dot{S} \cong e_1 \circ
-  ... \circ e_i \circ \dot{S}$ for all   $i \in \{1,...,n\}$
-\end{definition}
-\begin{definition}
-  Let $\dot{K}_n$ denote the $tPCP$ consisting of all pairwise
-  comparisons of $n$ different means.
-\end{definition}
-
-\begin{theorem}
-  Properties of $\circ$:
-  \begin{enumerate}
-    \item $\dot{S} \circ e_1 \circ e_2 \dot{S} \circ e_2 \circ e_1$
-    \item For $e,f,u,v \in E(\dot{S})$, such that $\dot{S} \circ e
-      \cong \dot{S} \circ f$, and $\dot{S} \circ e \circ u \cong
-      \dot{S} \circ u \circ v$ then $\exists z \in E(\dot{S} \circ f)$
-      such that $\dot{S} \circ f \circ z \cong \dot{S} \circ e \circ
-      u$. 
-  \end{enumerate}
-\end{theorem}
-
-\begin{theorem}
-  The tPCP $\dot{K}_4$ has 6 distinct (up to isomorphism) fixed
-  rejection sequences. Figure \ref{fig:k4} shows a graphical
-  representation of the distinct rejection sequences for
-  $K_4$. $\dot{K}_5$ has $657$ and $\dot{K}_6$ has $4836277$ it seems
-  that this sequence grows in the order of $\mathcal{O}(e^{n^2})$.
-\end{theorem}
-
-\begin{proof}
-  Proof by exhaustive computer search.
-\end{proof}
-
-\begin{figure}
-
-  \caption{Rejection tree for $\dot{K}_4$}
-  \label{fig:k4}
-\end{figure}
-
-\begin{theorem}
-  $\dot{K}_4$ can be represented by the complete graph $K_4$ with 4
-  vertices and 6 edges. The resulting null hypotheses are vertices of
-  the line-graph $\mathcal{L}{K_4}$. $\mathcal{L}{K_4}$ is an
-  octahedron. The exhaustive sets of $\dot{K_4}$ correspond to
-  $\mathcal{L}(K_4)$, the 3 faces of the octahedron with differently
-  labeled edges representing cliques of size 3 in $K_4$, the 3 edges
-  in the complementary graph of $\mathcal{L}(K_4)$, and the $6$
-  vertices themselves.
-\end{theorem}
 
@@
 \begin{proof}
-  Not yet done!
+  First assume that $I$ exhaustive, then for every $j \in S \setminus
+  I$ $H_{I \cup j} \subsetneq H_{I}$, that is $j \nsim I$ for all $j
+  \in S\setminus I$ and therefore $I$ is flat.
+
+  Next, consider $I \subseteq S$ and $I \in \mathscr{L}$, that is
+  $\sigma I = I$ assume that $I$ is not exhaustive, then $\exists J
+  \supersetneq I$ with $H_I \subset H_J$. Therefore, $\exists j \in J
+  \setminus I$ such that $H_I \in H_{J \setminus j}$ and therefore
+  also $H_I \subseteq H_j$. Consequently $H_{I \cup j} = H_{I}$, that
+  is $j \sim I$, which contradicts that $I$ is a flat and completes
+  the proof.
 \end{proof}
@@
 To get a result on strictly exhaustive set we need to restrict
-condition (\ref{cond:matroida}) further.
-
-\begin{theorem}
-  Consider a family of null hypotheses $\mathcal{H}$ for which above
-  the more stringent conditions hold:
-  \begin{align}
-    \label{cond:matroid2}
-    \bigcap_J \Theta_i &\nsubseteq \bigcup_{I\setminus J} \Theta_i,
-    \text{ and} \\
-    \bigcap_{J}\Theta_i &\nsupseteq     \bigcup_{I \setminus J} \Theta_i ,
-  \end{align}
-  % and additionally $\bigcap_S \Theta_i^C \neq
-  % \emptyset$ (practically that means that all alternatives may be true
-  % at the same time).
-  Then independence as defined in Definition \ref{def:indep} gives a
-  matroid for   which every flat corresponds to a strictly exhaustive
-  index set and vice versa.
-\end{theorem}
-
-\begin{proof}
-  Not yet done!
-  % We first show by contradiction that every flat is strictly
-  % exhaustive: Consider $I$ is a flat but that $I$ is not strictly
-  % exhaustive. As $I$ is a flat, we have for every $l \in S\setminus
-  % I$, $l \sim I$. Let $J \in \mathscr{I}$ be a maximally independent
-  % set in $I$. Then $J \cup l \in \mathscr{I}$ and $|J \cup l| =
-  % |J|+1$. If $I$ were not strictly exhaustive, then $\bigcap_{I}
-  % \Theta_i \cap \bigcap_{S\setminus I} \Theta^c_i = \emptyset$. 
-\end{proof}
-
-\begin{comment}
-  Above restriction (\ref{cond:matroid2}), already for the simple case
-  of $|J|=1$, is no longer a statment about the parallels of
-  $M$. 
-\end{comment}
+condition (\ref{cond:matroida}) further. {\sl It is an open question
+  whether one can find conditions that make it easier to show matroid
+  properties}.
+
+% \begin{theorem}
+%   Consider a family of null hypotheses $\mathcal{H}$ for which above
+%   the more stringent conditions hold:
+%   \begin{align}
+%     \label{cond:matroid2}
+%     \bigcap_J \Theta_i &\nsubseteq \bigcup_{I\setminus J} \Theta_i,
+%     \text{ and} \\
+%     \bigcap_{J}\Theta_i &\nsupseteq     \bigcup_{I \setminus J} \Theta_i ,
+%   \end{align}
+%   % and additionally $\bigcap_S \Theta_i^C \neq
+%   % \emptyset$ (practically that means that all alternatives may be true
+%   % at the same time).
+%   Then independence as defined in Definition \ref{def:indep} gives a
+%   matroid for   which every flat corresponds to a strictly exhaustive
+%   index set and vice versa.
+% \end{theorem}
+
+% \begin{proof}
+%   Not yet done!
+%   % We first show by contradiction that every flat is strictly
+%   % exhaustive: Consider $I$ is a flat but that $I$ is not strictly
+%   % exhaustive. As $I$ is a flat, we have for every $l \in S\setminus
+%   % I$, $l \sim I$. Let $J \in \mathscr{I}$ be a maximally independent
+%   % set in $I$. Then $J \cup l \in \mathscr{I}$ and $|J \cup l| =
+%   % |J|+1$. If $I$ were not strictly exhaustive, then $\bigcap_{I}
+%   % \Theta_i \cap \bigcap_{S\setminus I} \Theta^c_i = \emptyset$. 
+% \end{proof}
+
+% \begin{comment}
+%   Above restriction (\ref{cond:matroid2}), already for the simple case
+%   of $|J|=1$, is no longer a statment about the parallels of
+%   $M$. 
+% \end{comment}
 
@@
 for arrangement construction from polyhedral programming
-\cite{Avis-KF-92,Avis-KF-96,Ferrez-KF-Liebling-01} which are
+\citep{avis1992pivoting,avis1996reverse,ferrez2001cuts} which are
 polynomial in time and input size and compact in space!
 
+\bibliographystyle{alpha}
+\bibliography{cleanrefs}
+
 \end{document}
+
+\subsection{Some observations on pairwise comparisons}
+
+\begin{definition}
+  Let $X$ be a random variable taking values in $\mathbb{R}^m$ with
+  multivariate distribution $F(\bs{\mu})$ having mean vector $\bs{\mu}
+  \in \mathbb{R}^m$. Let $R \subset \{1,...,m\}^2$, such that $i <
+  j$ for all $(i,j) \in R$. For $(i,j) \in R$ $H_{ij} = \mu_i =
+  \mu_j$. We call the family of elementary hypotheses $\dot{R}
+  = \{H_{ij}: (i,j) \in R\}$ a {\bf two-sided pairwise comparison
+    problem. (tPCP)} Let $\dot{\mathcal{T}}(m)$ denote the set of all
+  possible pairwise comparison problems of $m$ means.
+\end{definition}
+
+\begin{bsp}
+  Consider $X \sim N((\mu_1,\mu_2,\mu_3,\mu_4),\sigma \mathbb{I}_4)$
+  let $R = \{(1,2),(1,3),(1,4)\}$ and $S =
+  \{(1,4),(2,4),(3,4),(1,3),(1,2),(2,3)\}$. Then the
+  corresponding tPCP's $\dot{R},\dot{S} \in \dot{\mathcal{T}}(4)$.
+\end{bsp}
+
+\begin{definition}
+  A two-sided pairwise comparison problem $\dot{S}$ defines an
+  undirected graph $G(\dot{S}) = \{M,T\}$ with Nodes $M = \{1,...,m\}$
+  and Edges $T$. We call $G(\dot{S})$ the {\bf comparigram}. We denote
+  by $V(G(\dot{S})$, $E(G(\dot{S})$ the vertex and edge sets of the
+  corresponding comparigram. If clear from the context we may just use
+  $E(\dot{S}),V(\dot{S})$. 
+\end{definition}
+
+\begin{bsp}
+  $\dot{R}$ in above Example is represented by a claw with 3 edges,
+  $\dot{S}$ is represented by the complete graph $K_4$ with 4 vertices.
+\end{bsp}
+
+\begin{definition}
+  Let $h_{lk}:  M^2 \rightarrow \dot{\mathcal{T}}(m)$,
+  $(l,k) \mapsto \{H_{ij}: (i,j) \in T\setminus \{(l,j)\}\}$ define the
+  {\bf rejection operator}. We write $h_{lk} \circ \dot{S} =
+  \dot{S}\setminus (l,k)$. Clearly   $h_{a} \circ h_{b} \circ \dot{S}
+  = h_{b} \circ h_{a} \circ \dot{S}$.
+\end{definition}
+
+\begin{definition}
+  Consider $\dot{R},\dot{S} \in \dot{\mathcal{T}}(m)$ with associated
+  comparigrams $L = G(\dot{R}), K = G(\dot{R})$ respectively. We say
+  that $\dot{R}$ {\bf isomorphic} to $\dot{S}$, $\dot{R} \cong
+  \dot{S}$ if $L$ is isomorphic to $K$, that is there exists a
+  bijection between nodes of the graphs $f:V(K) \rightarrow V(L)$, $a
+  \mapsto u$ such that for $a,b \in V(K)$, $f(a)$ is adjacent to
+  $f(b)$, if and only if $a$ and $b$ are adjacent to each other.
+\end{definition}
+
+\begin{definition}
+  Consider $\dot{S} \in \dot{\mathcal{T}}(m)$ and $h_a,h_b$, $a,b \in
+  E(S)$ we say that $h_a \cong h_b$ are {\bf equivalent rejections} if $h_a
+  \circ \dot{S} \cong h_b \circ \dot{S}$.
+\end{definition}
+
+\begin{definition}
+  Let $h = (h_1,...,h_n)$ whith $h_i \in T$ and $n \leq |T|$ define a
+  {\bf fixed rejection sequence} of length $n$.
+\end{definition}
+
+\begin{definition}
+  Consider PCP $\dot{S}$. Two fixed rejection sequences $h, e$ are
+  called {\bf equivalent} if $h_1\circ ... \circ h_i \circ \dot{S} \cong e_1 \circ
+  ... \circ e_i \circ \dot{S}$ for all   $i \in \{1,...,n\}$
+\end{definition}
+\begin{definition}
+  Let $\dot{K}_n$ denote the $tPCP$ consisting of all pairwise
+  comparisons of $n$ different means.
+\end{definition}
+
+\begin{theorem}
+  Properties of $\circ$:
+  \begin{enumerate}
+    \item $\dot{S} \circ e_1 \circ e_2 \dot{S} \circ e_2 \circ e_1$
+    \item For $e,f,u,v \in E(\dot{S})$, such that $\dot{S} \circ e
+      \cong \dot{S} \circ f$, and $\dot{S} \circ e \circ u \cong
+      \dot{S} \circ u \circ v$ then $\exists z \in E(\dot{S} \circ f)$
+      such that $\dot{S} \circ f \circ z \cong \dot{S} \circ e \circ
+      u$. 
+  \end{enumerate}
+\end{theorem}
+
+\begin{theorem}
+  The tPCP $\dot{K}_4$ has 6 distinct (up to isomorphism) fixed
+  rejection sequences. Figure \ref{fig:k4} shows a graphical
+  representation of the distinct rejection sequences for
+  $K_4$. $\dot{K}_5$ has $657$ and $\dot{K}_6$ has $4836277$ it seems
+  that this sequence grows in the order of $\mathcal{O}(e^{n^2})$.
+\end{theorem}
+
+\begin{proof}
+  Proof by exhaustive computer search.
+\end{proof}
+
+\begin{figure}
+
+  \caption{Rejection tree for $\dot{K}_4$}
+  \label{fig:k4}
+\end{figure}
+
+\begin{theorem}
+  $\dot{K}_4$ can be represented by the complete graph $K_4$ with 4
+  vertices and 6 edges. The resulting null hypotheses are vertices of
+  the line-graph $\mathcal{L}{K_4}$. $\mathcal{L}{K_4}$ is an
+  octahedron. The exhaustive sets of $\dot{K_4}$ correspond to
+  $\mathcal{L}(K_4)$, the 3 faces of the octahedron with differently
+  labeled edges representing cliques of size 3 in $K_4$, the 3 edges
+  in the complementary graph of $\mathcal{L}(K_4)$, and the $6$
+  vertices themselves.
+\end{theorem}

### 2013-08-20 presentation/presentation.tex
further clean up operations, mostly animations left to do

@@
                                 %intersection hypotheses since 
-        $H_{{i,j}} = H_{\{1,2,3\}}, i\neq j,$
+        $H_{\{i,j\}} = H_{\{1,2,3\}}, i\neq j,$
 
@@
 \begin{itemize}
-\item $H_i$ depends on $H_I$: $H_{I \cup i} = H_{I}$
-\item induces a dependence relation $i \sim_H I$ on index sets $I \in \mathcal{P}(S)$
-\item $I$ independent iff for all $i
+\item $H_i$ \alert{depends} on $H_I$: $H_{I \cup i} = H_{I}$
+\item induces a \alert{dependence relation} $i \alert{\sim_H} I$ on index sets $I \in \mathcal{P}(S)$
+\item $I$ \alert{independent} iff for all $i
   \in I$ $i \nsim_H I\setminus i$ (\ie $H_{I} \subsetneq H_{I\setminus i}$)
 \pause
-\item For pairwise comparisons $I$ independent iff there are no cycles in $\mathscr{G}_I$
+\item For pairwise comparisons $I$ independent iff there are no cycles
+  in the corresponding graph
 \item For general contrast tests $I$ independent iff $\bs{C_I}$ has full
@@
  structure
-\item {\em Cryptomorphism} \ie many ways to define a matroid:
+\item \alert{Cryptomorphism} \ie many ways to define a matroid:
  \begin{itemize}
@@
  \begin{minipage}{.5\textwidth}
+
+** TODO shorten animation
+
    \only<2->{\input{cycle}}
@@
  \begin{minipage}{.5\textwidth}
- \only<1>{\input{independent}}
- \only<2>{\input{cycle}}
- \invisible<1>{\alert{\footnotesize minimally dependent set!}}
+
+** TODO animate independent sets
+
+ \input{independent}
+% \only<2>{\input{cycle}}
+% \invisible<1>{\alert{\footnotesize minimally dependent set!}}
  \end{minipage} 
@@
 \begin{frame}
-  \frametitle{}
+  \frametitle{Matroid structure}
+
+** TODO add some pictures detailing general structure
+
   \begin{itemize}
-  \item Circuits are minimally dependent sets (remove any element to
+  \item \alert{Circuits} are minimally dependent sets (remove any element to
     get an independent set)
-  \item Independent sets are sets for which no subset of elements is a
+  \item \alert{Independent sets} are sets for which no subset of elements is a
     circuit
-  \item Maximally independent sets are bases (Base axioms)
-  \item Cardinality of largest independent subset is the rank of a set
+  \item Maximally independent sets are \alert{bases} (Base axioms)
+  \item Cardinality of largest independent subset is the \alert{rank} of a set
     (Rank axioms)
-  \item Maximally dependent sets are called flats (Closure axioms)
+  \item Maximally dependent sets are called \alert{flats} (Closure axioms)
   \end{itemize}
@@
 \frametitle{Dependence axioms}
-
+** TODO animate dependence
 \begin{tabular}{ll}
@@
  Relation $\sim_M$ on elements and subsets of $S$,
- if for any $x \in S$, and sets $Y,Z \subseteq S$ holds
+ if for any $j \in S$, and sets $I,K \subseteq S$ holds
  \begin{enumerate}
- \item $y \sim Y$ for all $y \in Y$
- \item if $x \sim_M Y$ and $x \nsim_M Y\setminus y$
-   then $y \sim x \cup (Y \setminus y)$
- \item if $x \sim_M Y$ and for all $y \in Y$, $y \sim_M Z$ then $x \sim_M Z$
+ \item $i \sim_M I$ for all $i \in I$
+ \item if $j \sim_M I$ and $j \nsim_M I\setminus \{i\}$
+   then $i \sim \{j\} \cup (I \setminus \{i\})$
+ \item if $j \sim_M I$ and for all $i \in I$, $i \sim_M K$ then $j \sim_M K$
  \end{enumerate}
@@
    \item<1-> $H_i$ depends on $H_I$ iff $H_{I \cup i} = H_{I}$
-   \item<2-> Does this mean $\mathscr{H}$ defines a matroid $M$, such that
-     $\sim_H = \sim_M$?
+   \item<2-> Does this mean a family of algebraically dependent
+     hypotheses defines a matroid $M$, such that $\sim_H = \sim_M$?
    \end{itemize}
@@
    hypotheses $H_I$ iff the corresponding edge $e_j$ is dependent on
-   the edge set $\{e_i:i \in I\}$ of the cycle matroid defined by
-   $\mathscr{G}_I$.
+   the edge set $\{e_i:i \in I\}$ of the \alert{cycle matroid} defined by
+   the corresponding graph.
  \end{block}
@@
    an elementary hypothesis $H_j$ depends on intersection hypotheses
-   $H_I$ iff the corresponding contrast vector $\bs{c_j}$ is
-   dependent on the column vectors of the matrix $\bs{C}_I$ in the
-   vectorial matroid defined by $\bs{C}$ on $\mathbb{R}^m$.
+   $H_I$ iff the contrast vector $\bs{c_j}$ is a linear combination of
+   column vectors of contrast matrix $\bs{C}_I$ which corresponds to
+   the \alert{vectorial matroid} defined by $\bs{C}$ on $\mathbb{R}^m$.
  \end{block}
@@
 \begin{frame}
- \frametitle{Matroids from null hypotheses}
+ \frametitle{Matroids from algebraically dependent hypotheses}
  {\Large What is it good for?\\
@@
 
-\begin{frame}
-\frametitle{Maximally independent sets}
- \begin{itemize}
- \item Bases are maximally independent sets of $S$
- \item Rank $\rho I$ is the cardinality of a maximally indepedendent set in $I$
- \item Circuits are minimally dependent sets, \ie each subset of $C$
-   with rank $\rho C - 1$ is a maximally independent set of $C$
- \end{itemize}
-\end{frame}
+% \begin{frame}
+% \frametitle{Maximally independent sets}
+%  \begin{itemize}
+%  \item Bases are maximally independent sets of $S$
+%  \item Rank $\rho I$ is the cardinality of a maximally indepedendent set in $I$
+%  \item Circuits are minimally dependent sets, \ie each subset of $C$
+%    with rank $\rho C - 1$ is a maximally independent set of $C$
+%  \end{itemize}
+% \end{frame}
 
-\begin{frame}
-\frametitle{Maximally dependent sets}
-\begin{itemize}
-\item The closure of a set $\sigma I = \{i \in S: i \sim I\}$
-\item Closed sets {\em flats}: $\mathscr{L} = \{\sigma I: I \subseteq S\}$
-\item Hyperplanes are flats of rank $\rho S - 1$
-\item All flats are given by closing hyperplanes under intersection
-\item Circuits in $M^*$ are the complements of hyperplanes in $M$
-\end{itemize} % ends low level
-\end{frame}
+% \begin{frame}
+% \frametitle{Maximally dependent sets}
+% \begin{itemize}
+% \item The closure of a set $\sigma I = \{i \in S: i \sim I\}$
+% \item Closed sets {\em flats}: $\mathscr{L} = \{\sigma I: I \subseteq S\}$
+% \item Hyperplanes are flats of rank $\rho S - 1$
+% \item All flats are given by closing hyperplanes under intersection
+% \item Circuits in $M^*$ are the complements of hyperplanes in $M$
+% \end{itemize} % ends low level
+% \end{frame}
 
-\begin{frame}{Examples}
-\framesubtitle<1>{Independent set}
-\framesubtitle<2>{Base}
-\framesubtitle<4>{Dependent set}
-\framesubtitle<3>{Cycle}
-\framesubtitle<5>{Closure}
-\only<1>{\input{independent}}
-\only<2>{\input{base}}
-\only<4>{\input{dependent}}
-\only<3>{\input{cycle}}
-\only<5>{\input{closure}}
-\end{frame}
+% \begin{frame}{Examples}
+% \framesubtitle<1>{Independent set}
+% \framesubtitle<2>{Base}
+% \framesubtitle<4>{Dependent set}
+% \framesubtitle<3>{Cycle}
+% \framesubtitle<5>{Closure}
+% \only<1>{\input{independent}}
+% \only<2>{\input{base}}
+% \only<4>{\input{dependent}}
+% \only<3>{\input{cycle}}
+% \only<5>{\input{closure}}
+% \end{frame}
 
@@
 
+\item Flats are maximally dependent sets \ie $j \nsim I \forall j \in
+  S\setminus I$
 \item The flats of a matroid are closed under intersection
-
-\item Every elementary hypothesis is a flat
-\item For each $I \subseteq S$, $H_I = H_{\sigma I}$
+\item $S$ (global null hypothesis) is a flat
+\item Every single element of $S$ (elementary hypothesis) is a flat
+\item For each $I \subseteq S$, the $H_I$ as a subset of the parameter
+  space is identical to the intersection hypotheses defined by it
+  closure $H_{\sigma I}$ (\ie flats correspond to weakly exhaustive
+  sets \cite{bergmann1988improvements})
 \item Define a stepwise procedure on the flats of $M$
-
 \end{itemize} % ends low level
@@
  \begin{block}{Multiple testing procedure for algebraically related hypotheses}
-   Assume that $\mathscr{H}$ with dependence relation $\sim_H$
-   defines a matroid on $M$ on $S$. Then a test procedure that
-   rejects elementary hypotheses $H_i$ if all $H_E$, $i \in E$, $E
-   \in \mathscr{L}$ are rejected using a local level-$\alpha$ test
-   controls the FWER in the strong sense.
+   Assume that a family of null hypotheses $\mathscr{H}$ with
+   dependence relation $\sim_H$ defines a matroid $M$ on $S$. Then a
+   test procedure that rejects elementary hypotheses $H_i$ if for all
+   flats $E \subseteq S$ with $i \in E$, $H_E$ is rejected using a
+   local level-$\alpha$ test, controls the FWER in the strong sense.
  \end{block}
 
+\end{frame}
+
+\begin{frame}
+\frametitle{Old wine in new bottles?}
+
+\begin{itemize}
+\item \cite{shaffer1986modified} represents pairwise comparisons
+ by set partitions
+\item \cite{bergmann1988improvements} define the concepts of an
+ exhaustive sets
+\item \cite{bernhard1991computergestuetzte} shows properties of
+ exhaustive sets that are basically \alert{closure axioms} of a matroid
+ and provides algorithms to check each intersection hypotheses if is
+ a flat using rank function.
+\item \cite{westfall1997multiple} generates flats using dependence
+ arguments in vector spaces.
+\item \cite{weichert2000robuste} suggest an MTP based on clique
+ decomposition of the graph, cliques in the graph are flats
+\item \cite{westfall2007multiple} provide a branch and bound algorithm
+ to generate all flats. They note that for pairwise comparisons all
+ exhaustive hypotheses are \alert{partition hypotheses}
+\item \cite{brannath2010shortcuts} provide shortcuts for stepwise
+  procedures on \eg the set of exhaustive index sets
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Added benefits}
+ \begin{itemize}
+ \item Many results about exhaustive sets follow directly from
+   corresponding results in matroid theory.
+ \item Efficient algorithms to find the flats of specific and general
+   matroids are available from matroid theory.
+ \item \alert{Scum principle} \cite{welsh2010matroid} everything
+   important happens at the top \ie it is sufficient to find the
+   \alert{hyperplanes} of the corresponding matroid.
+ \item The structure of flats provides some intuition to procedures
+   that operate on exhaustive sets. 
+ \end{itemize}
 \end{frame}
@@
 
-\begin{frame}
-\frametitle{Connections to existing procedures}
-
-\textbf{Old wine in new bottles ((c) L. Hothorn)}
-\begin{itemize}
-\item \cite{shaffer1986modified} represents pairwise comparisons
- by set partitions
-\item \cite{bergmann1988improvements} define the concepts of an
- exhaustive sets
-\item \cite{bernhard1991computergestuetzte} shows properties of
- exhaustive sets that are basically {\em closure axioms} of a matroid
- and provides algorithms to check each intersection hypotheses if is
- a flat using rank function.
-\item \cite{westfall1997multiple} generates flats using dependence
- arguments in vector spaces.
-\item \cite{weichert2000robuste} suggest an MTP based on clique
- decomposition of the graph, cliques in the graph are flats
-\item \cite{westfall2007multiple} provide a branch and bound algorithm
- to generate all flats. They note that for pairwise comparisons all
- exhaustive hypotheses are {\em partition hypotheses}
-\end{itemize}
-\end{frame}
-
-\begin{frame}
-\frametitle{Matroid representation}
- \begin{itemize}
- \item proofs are shorter (more elegant),
- \item efficient algorithms exist to find the flats of matroids,
- \item {\em Scum principle}\cite{welsh2010matroid} everything
-   important happens at the top (hyperplanes)
- \item we may define procedures using independent sets,
- \item \ldots{}
- \end{itemize}
-\end{frame}
-
 \section{Graphical approaches for algebraically dependent and
@@
    \begin{itemize}
-   \item Provides weights $w_{i,I}$ for each index $i$ and each
+   \item Provides weights $w(I)_i$ for each index $i$ and each
      non-empty subset $I \subset S$
@@
      weighted intersection hypotheses tests
-   \item We have $\sum_{i \in S} w_{i,I} \leq 1$ (conservative),
-   \item and $w_{i,I} \leq w_{i,J}$ for $J \subseteq I$ (consonance).
+   \item We have $\sum_{i \in S} w(I)_{i} \leq 1$ (conservative),
+   \item and $w(I)_{i} \leq w(J)_{i}$ for $J \subseteq I$ (consonance).
    \end{itemize}
@@
  \frametitle{Fixed sequence test of pairwise mean comparisons}
-\begin{itemize}
-\item {\bf Picture of hierarchical test hereabouts}
-\end{itemize}
+
+** TODO Picture of hierarchical test hereabouts
 
@@
 \begin{enumerate}
-\item Pretest using weighted Bonferroni or conventional graphical
- approach $R = \{i \in S: H_i \textrm{ rejected by pretest}\}$
-\item Compute all matroid hyperplanes $\mathscr{H}$ for the matroid on
- the remaining hypotheses $I$
-\item $R \gets R \cup \{i: p_i \leq \min_J w_{i,J} \alpha, J \in \mathscr{H}\}$
-\item $A \gets \{\bigcup J: \min_j (p_j / w_{j,J}) > \alpha\}$
+\item Pretest using consonant weighted closed test procedure (\eg
+  graphical approach), let $R = \{i \in S: H_i \textrm{ rejected by pretest}\}$
+\item Compute all matroid hyperplanes $\mathscr{L}$ for the matroid
+  restricted to the remaining hypotheses $I$
+\item $R \gets R \cup \{i: p_i \leq \min\{ w(J)_{i}\alpha: J \in
+  \mathscr{L} \} \}$
+\item $A \gets \{\bigcup J: \min\{p_j / w(J)_{j}: j \in J\} > \alpha,
+J \in \mathscr{L} \}$
 \item $I \gets I \setminus R$
-\item Refine: $\mathscr{H} \gets \{J\in \mathscr{L}: \rho J = \rho
- \mathscr{H} - 1\}$
-\item Prune: $\mathscr{H} \gets \{J \in \mathscr{H}: J \cap R = \emptyset\}$
-\item Repeat until $\rho \mathscr {H} = 0$
+\item Compute hyperplanes of next smaller rank: $\mathscr{L} \gets \{J\in \mathscr{L}: \rho J = \rho
+ \mathscr{L} - 1\}$
+\item Remove interval of rejected hypotheses: $\mathscr{L} \gets \{J \in \mathscr{L}: J \cap R = \emptyset\}$
+\item Reduce to interval of not accepted: $\mathscr{L} \gets \{J \in \mathscr{L}: L \nsubseteq A\}$
+\item Repeat until $\rho \mathscr{L} = 0$
 \item Reject $R$ and $I \setminus A$
@@
 \begin{itemize}
-\item We give sufficient (and probably necessary) conditions for a
- general family of hypotheses $\mathscr{H}$ to define a matroid
-\item We give sufficient (and probably necessary) conditions on
- $\mathscr{H}$ for $\mathscr{L}$ to be the set of strictly exhaustive
- sets
+\item We give sufficient necessary conditions for a general family of
+  hypotheses $\mathscr{H}$ to define a matroid
 \item Systems with one sided hypotheses, and parallel hypotheses (\eg
@@
  $\mathscr{L}$ all strictly exhaustive
+\item It is an open question under which conditions the set of
+  strictly exhaustive sets is given by the flats of a matroid
 \end{itemize}
@@
 for arrangement construction from polyhedral programming
-\cite{Avis-KF-92,Avis-KF-96,Ferrez-KF-Liebling-01} which are
+\cite{avis1992pivoting,avis1996reverse,ferrez2001cuts} which are
 polynomial in time and input size and compact in space!

### 2013-08-20 presentation/presentation.tex
incorporated willis comments

@@
 
-\item Flats are maximally dependent sets \ie $j \nsim I \forall j \in
+\item Flats are maximally dependent sets \ie $j \nsim I$ $\forall j \in
   S\setminus I$
@@
   sets \cite{bergmann1988improvements})
-\item Define a stepwise procedure on the flats of $M$
+%\item Define a stepwise procedure on the flats of $M$
 \end{itemize} % ends low level
@@
  hypotheses $H_{r+1}$ through $H_{r+s}$ can be tested simultaneously
- at full level $\alpha_ra$.
+ at full level $\alpha_r$.
 \end{block}
@@
   \mathscr{L} \} \}$
-\item $A \gets \{\bigcup J: \min\{p_j / w(J)_{j}: j \in J\} > \alpha,
-J \in \mathscr{L} \}$
+\item $A \gets \{\bigcup J: \forall j \in J,\, p_j > w(J)_{j}\alpha, J \in \mathscr{L} \}$
 \item $I \gets I \setminus R$
@@
 \item Remove interval of rejected hypotheses: $\mathscr{L} \gets \{J \in \mathscr{L}: J \cap R = \emptyset\}$
-\item Reduce to interval of not accepted: $\mathscr{L} \gets \{J \in \mathscr{L}: L \nsubseteq A\}$
+\item Reduce to interval of not accepted: $\mathscr{L} \gets \{J \in \mathscr{L}: J \nsubseteq A\}$
 \item Repeat until $\rho \mathscr{L} = 0$
@@
   hypotheses $\mathscr{H}$ to define a matroid
-\item Systems with one sided hypotheses, and parallel hypotheses (\eg
+\item Systems with one-sided hypotheses, and parallel hypotheses (\eg
  combined non-inferiority/superiority) do not give matroids with
@@
 \begin{frame}
- \frametitle{One sided hypotheses}
+ \frametitle{One-sided hypotheses}
  \begin{block}{One-sided linear contrast hypotheses}
-   A family of {\em one sided linear contrast hypotheses} is defined
+   A family of {\em one-sided linear contrast hypotheses} is defined
    by a contrast matrix $\bs{A} \in \mathbb{R}^{m\times n}$ and

### 2013-08-21 presentation/presentation.tex
streamlined shaffer procedure

@@
 \begin{document}
-\frametitle{Matroids}
-
 \maketitle
@@
 
-\section{Algebraically dependent hypotheses}
+\section{Introduction}
 
 \begin{frame}
-\frametitle{Some history:}
+\frametitle{Algebraic dependence}
 \framesubtitle{Testing hypotheses of no differences in pairwise
@@
   \begin{enumerate}
-  \item Test $H_ {\{1,2,3\}} = H_1\cap H_2 \cap H_3 :
-    \mu_1=\mu_2=\mu_3$ at level $\alpha$,
-  \item test each $H_1,H_2,H_3$ at level $\alpha$,
+  \item Test each $H_1,H_2,H_3$ at level $\alpha/3$.
+  \item If any is rejected test the remaining at level $\alpha$.
   \end{enumerate}
@@
   \begin{itemize}
-    \item  essentially closed test, since %; no need to test pairwise
-                                %intersection hypotheses since 
-        $H_{\{i,j\}} = H_{\{1,2,3\}}, i\neq j,$
+    \item  essentially closed test, since %; 
+      \begin{enumerate}
+        \item Test $H_ {\{1,2,3\}} = H_1\cap H_2 \cap H_3 :
+          \mu_1=\mu_2=\mu_3$ at level $\alpha$,
+        \item No need to test pairwise intersection hypotheses, as
+          considered as subsets of the parameter space          
+          $H_{\{i,j\}} = H_{\{1,2,3\}}, i\neq j$
+      \end{enumerate}
 
-    \item  Bonferroni: if one of the hypotheses is rejected at level $\alpha/3$ the others can be tested at level alpha.
-    \item  Hierarchical: if a predetermined primary hypothesis is rejected at level $\alpha$ the other two can be tested at level $\alpha$.
-    \item  $\Rightarrow$ improves respective Bonferroni-Holm or fixed sequence tests.
+    \item  $\Rightarrow$ improves respective Bonferroni-Holm test.
+  \end{itemize}
+
+\end{frame}
+\begin{frame}
+\frametitle{Contextual relationships}
+\framesubtitle{Testing hypotheses of no differences in pairwise
+  comparisons of 3 treatments}
+\begin{block}{$H_1:\mu_1=\mu_2,\; H_2:\mu_2=\mu_3,\;H_3:\mu_1=\mu_3$}
+  \begin{enumerate}
+  \item Test $H_1$ at level $\alpha$.
+  \item If $H_1$ is rejected, test $H_2,H_3$ at level $\alpha$ simultaneously.
+  \end{enumerate}
+  protects the FWER at level $\alpha$.
+\end{block}
+\pause
+  \begin{itemize}
+    \item  Uses a weighted Bonferroni test to reject $H_{1,2,3}$.
+    \item Requires choosing a pre-fixed ``primary'' hypotheses. This
+      depends on the context.
+    \item  $\Rightarrow$ improves respective fixed sequence tests (no
+      need to specify hierarchy for $H_2$ and $H_3$).
   \end{itemize}
@@
   \begin{minipage}{.6\textwidth}
-    \begin{block}{$m=4, n=6, \only<2->{j=2}$}
+    \begin{block}{$m=4, n=6, j=2$}
       % \begin{itemize}
@@
         \begin{description}
-        \item<3->[Static:]  $H_1$ and $H_2$ are false
+        \item<2->[Static:]  $H_1$ and $H_2$ are false
             $\Rightarrow$ $t_2 = 3$ (at most $H_4, H_5 $ and $H_6$ can be
             simultaneously true)
-        \item<7->[Dynamic:] $H_1$ and $H_6$ are false $\Rightarrow$
+        \item<5->[Dynamic:] $H_1$ and $H_6$ are false $\Rightarrow$
           $t^*_2 = 2 < t_2 =3$ (only $H_3$, $H_4$, or $H_2$, $H_5$ can be
@@
 \begin{frame}
-\frametitle{Dependent null hypotheses}
+\frametitle{Algebraically dependent null hypotheses}
 
 \begin{itemize}
-\item $H_i$ \alert{depends} on $H_I$: $H_{I \cup i} = H_{I}$
-\item induces a \alert{dependence relation} $i \alert{\sim_H} I$ on index sets $I \in \mathcal{P}(S)$
+\item $H_j$ \alert{depends} on $H_I$: $H_{I \cup j} = \bigcap_{i \in
+    I} H_i \cap H_j = \bigcap_{i \in I} H_i = H_{I}$
+\item induces a \alert{dependence relation} $j \alert{\sim_H} I$ on index sets $I \in \mathcal{P}(S)$
 \item $I$ \alert{independent} iff for all $i
@@
 
-\item Matroids generalize the structures of graphs and vector spaces,
- \eg cycles, bases, independent sets
+\item Matroids generalize the structures of graphs and vector spaces.
 \item A matroid $M$ on $S$ endows $\mathcal{P}(S)$ with an algebraic
- structure
+ structure, \eg cycles, bases, independent sets, rank
 \item \alert{Cryptomorphism} \ie many ways to define a matroid:
@@
 \begin{frame}
- \frametitle{Example: Multiple doses}
-
+ \frametitle{Example: QVA}
 \begin{tabular}{ll}
  \begin{minipage}{.4\textwidth}
-   \begin{tikzpicture}[overlay]
-     \node[yshift=70bp,xshift=-40bp] at (current page.south west){\input{lsd4.tex}};
+   \only<1,3>{   \begin{tikzpicture}[overlay]
+       \node[yshift=100bp,xshift=-40bp] at (current page.south west){\input{qvameans.tex}};
      \end{tikzpicture}
+   }   
+   \only<2>{\begin{tikzpicture}[overlay]
+       \node[yshift=-6cm,xshift=-1.4cm] at (current page.north west){\input{qvaalg.tex}};
+     \end{tikzpicture}
+   }
 \end{minipage} &
  \begin{minipage}{.6\textwidth}
-   \begin{tikzpicture}[remember picture,overlay]
-     \node[yshift=100bp,xshift=160bp] at (current page.north west){\input{lsd4t.tex}};
-   \end{tikzpicture}
+   \begin{block}{QVA Trial}
+     \begin{itemize}
+     \item QVA combination treatmend of QAB and NVA
+     \item Three treatment groups against placebo (Pl)
+     \item Fourth active comparator Tio
+     \item Some secondary endponts SE
+     \item<2-> If $H_1$ is rejected
+     \item<2-> Test $H_2$, $H_4$ and $H_3$, $H_5$ simultaneously at level $\alpha/2$
+     \item<2-> If only one is rejected no $\alpha$ may be passed on!
+     \end{itemize}
+   \end{block}
  \end{minipage}

### 2013-08-21 presentation/shaffer.tex
streamlined shaffer procedure

@@
 \node (H4) at (025bp,-325bp)[draw,circle,fill=white] {$\mu_4$};
-\only<-2>{\draw [line width=1pt] (H1) to (125bp, -124bp) node[fill=blue!20] {$H_{1}$} to (H2);}
-\only<-2>{\draw [line width=1pt] (H1) to (075bp, -175bp) node[fill=blue!20] {$H_{2}$} to (H3);}
-\only<3>{\draw [line width=1pt] (H1) to (125bp, -124bp) node[fill=red!20] {$H_{1}$} to (H2);}
-\only<3>{\draw [line width=1pt] (H1) to (075bp, -175bp) node[fill=red!20] {$H_{2}$} to (H3);}
-\only<-5>{\draw [line width=1pt] (H4) to (025bp, -227bp) node[fill=blue!20] {$H_{3}$} to (H1);}
-\only<-4>{\draw [line width=1pt] (H2) to (225bp, -224bp) node[fill=blue!20] {$H_{4}$} to (H3);}
-\only<-4>{\draw [line width=1pt] (H4) to (174bp, -176bp) node[fill=blue!20] {$H_{5}$} to (H2);}
-\only<-4>{\draw [line width=1pt] (H3) to (126bp, -324bp) node[fill=blue!20] {$H_{6}$} to (H4);}
-\only<4,5>{\draw [line width=2pt] (H2) to (225bp, -224bp) node[fill=green!20] {$H_{4}$} to (H3);}
-\only<4,5>{\draw [line width=2pt] (H4) to (174bp, -176bp) node[fill=green!20] {$H_{5}$} to (H2);}
-\only<4,5>{\draw [line width=2pt] (H3) to (126bp, -324bp) node[fill=green!20] {$H_{6}$} to (H4);}
 
-\only<6>{\draw [line width=1pt] (H1) to (125bp, -124bp) node[fill=blue!20] {$H_{1}$} to (H2);}
-\only<7>{\draw [line width=1pt] (H1) to (125bp, -124bp) node[fill=red!20] {$H_{1}$} to (H2);}
-\only<6>{\draw [line width=2pt] (H3) to (126bp, -324bp) node[fill=blue!20] {$H_{6}$} to (H4);}
-\only<7>{\draw [line width=2pt] (H3) to (126bp, -324bp) node[fill=red!20] {$H_{6}$} to (H4);}
+\only<1,4->{\draw [line width=1pt] (H1) to (125bp, -124bp) node[fill=blue!20] {$H_{1}$} to (H2);}
+\only<2-3,5->{\draw [line width=1pt] (H1) to (125bp, -124bp) node[fill=red!20] {$H_{1}$} to (H2);}
 
-\only<6,7,8,9>{\draw [line width=1pt] (H1) to (075bp, -175bp) node[fill=blue!20] {$H_{2}$} to (H3);}
-\only<6,7>{\draw [line width=1pt] (H4) to (025bp, -227bp) node[fill=blue!20] {$H_{3}$} to (H1);}
-\only<6,7>{\draw [line width=1pt] (H2) to (225bp, -224bp) node[fill=blue!20] {$H_{4}$} to (H3);}
-\only<6,7,8,9>{\draw [line width=1pt] (H4) to (174bp, -176bp) node[fill=blue!20] {$H_{5}$} to (H2);}
-\only<8>{\draw [line width=2pt] (H4) to (025bp, -227bp) node[fill=green!20] {$H_{3}$} to (H1);}
-\only<8>{\draw [line width=2pt] (H2) to (225bp, -224bp) node[fill=green!20] {$H_{4}$} to (H3);}
-\only<9>{\draw [line width=2pt] (H4) to (025bp, -227bp) node[fill=blue!20] {$H_{3}$} to (H1);}
-\only<9>{\draw [line width=2pt] (H2) to (225bp, -224bp) node[fill=blue!20] {$H_{4}$} to (H3);}
+\only<1,4->{\draw [line width=1pt] (H1) to (075bp, -175bp) node[fill=blue!20] {$H_{2}$} to (H3);}
+\only<2-3>{\draw [line width=1pt] (H1) to (075bp, -175bp) node[fill=red!20] {$H_{2}$} to (H3);}
+\only<7->{\draw [line width=2pt] (H1) to (075bp, -175bp) node[fill=green!20] {$H_{2}$} to (H3);}
 
-\only<10>{\draw [line width=1pt] (H1) to (075bp, -175bp) node[fill=blue!20] {$H_{2}$} to (H3);}
-\only<10,11,12>{\draw [line width=1pt] (H4) to (025bp, -227bp) node[fill=blue!20] {$H_{3}$} to (H1);}
-\only<10,11,12>{\draw [line width=1pt] (H2) to (225bp, -224bp) node[fill=blue!20] {$H_{4}$} to (H3);}
-\only<10>{\draw [line width=1pt] (H4) to (174bp, -176bp) node[fill=blue!20] {$H_{5}$} to (H2);}
-\only<11,12>{\draw [line width=2pt] (H1) to (075bp, -175bp) node[fill=green!20] {$H_{2}$} to (H3);}
-\only<11,12>{\draw [line width=2pt] (H4) to (174bp, -176bp) node[fill=green!20] {$H_{5}$} to (H2);}
+\only<1->{\draw [line width=1pt] (H4) to (025bp, -227bp) node[fill=blue!20] {$H_{3}$} to (H1);}
+\only<6>{\draw [line width=2pt] (H4) to (025bp, -227bp) node[fill=green!20] {$H_{3}$} to (H1);}
+
+\only<1-2,4-7>{\draw [line width=1pt] (H2) to (225bp, -224bp) node[fill=blue!20] {$H_{4}$} to (H3);}
+\only<3,6>{\draw [line width=2pt] (H2) to (225bp, -224bp)  node[fill=green!20] {$H_{4}$} to (H3);}
+\only<8>{\draw [line width=2pt,dashed] (H2) to (225bp, -224bp) node[fill=green!20] {$H_{4}$} to (H3);}
+
+\only<1-2,4->{\draw [line width=1pt] (H4) to (174bp, -176bp) node[fill=blue!20] {$H_{5}$} to (H2);}
+\only<3,7->{\draw [line width=2pt] (H4) to (174bp, -176bp) node[fill=green!20] {$H_{5}$} to (H2);}
+
+\only<1-2,4->{\draw [line width=1pt] (H3) to (126bp, -324bp) node[fill=blue!20] {$H_{6}$} to (H4);}
+\only<3>{\draw [line width=2pt] (H3) to (126bp, -324bp) node[fill=green!20] {$H_{6}$} to (H4);}
+\only<5->{\draw [line width=1pt] (H3) to (126bp, -324bp) node[fill=red!20] {$H_{6}$} to (H4);}
+
+% \only<6>{\draw [line width=1pt] (H1) to (125bp, -124bp) node[fill=blue!20] {$H_{1}$} to (H2);}
+% \only<7>{\draw [line width=1pt] (H1) to (125bp, -124bp) node[fill=red!20] {$H_{1}$} to (H2);}
+% \only<6>{\draw [line width=2pt] (H3) to (126bp, -324bp) node[fill=blue!20] {$H_{6}$} to (H4);}
+
+% \only<6,7,8,9>{\draw [line width=1pt] (H1) to (075bp, -175bp) node[fill=blue!20] {$H_{2}$} to (H3);}
+% \only<6,7>{\draw [line width=1pt] (H4) to (025bp, -227bp) node[fill=blue!20] {$H_{3}$} to (H1);}
+
+% \only<6,7,8,9>{\draw [line width=1pt] (H4) to (174bp, -176bp) node[fill=blue!20] {$H_{5}$} to (H2);}
+
+% \only<8>{\draw [line width=2pt] (H2) to (225bp, -224bp) node[fill=green!20] {$H_{4}$} to (H3);}
+% \only<9>{\draw [line width=2pt] (H4) to (025bp, -227bp) node[fill=blue!20] {$H_{3}$} to (H1);}
+% \only<9>{\draw [line width=2pt] (H2) to (225bp, -224bp) node[fill=blue!20] {$H_{4}$} to (H3);}
+
+% \only<10,11,12>{\draw [line width=1pt] (H4) to (025bp, -227bp) node[fill=blue!20] {$H_{3}$} to (H1);}
+
+% \only<10>{\draw [line width=1pt] (H4) to (174bp, -176bp) node[fill=blue!20] {$H_{5}$} to (H2);}
+% \only<11,12>{\draw [line width=2pt] (H1) to (075bp, -175bp) node[fill=green!20] {$H_{2}$} to (H3);}
+% \only<11,12>{\draw [line width=2pt] (H4) to (174bp, -176bp) node[fill=green!20] {$H_{5}$} to (H2);}
 
